---
title: "Analysing oxgyen values from BCCM for areas of interest for Pacific Hake"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "oxygen-for-hake-cache/",
  fig.path = "oxygen-for-hake-cache/",
  fig.width = 7,
  fig.height = 9
)
  # comment = "#>",

knitr::dep_auto()  # or dep_auto seems cleverer, though not always working it
                   # seems so be careful (final chunk wasn't rerun when
                   # penultimate one was changed). May be because have to have
                   # this command in there from the beginning?

options(pillar.sigfig = 6)   # show 6 sig figs when printing tibbles (so can see
                             # length resolutions, but not overwhelmed by
                             # details for bin_min and bin_max

```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("oxygen-for-hake.Rmd")
```

```{r, packages, echo = FALSE, cache = FALSE}
library(dplyr)
load_all()
#library(pacea)    # Probably need latest version
```

Curious as to whether the midwater oxygen is lower when hake are no longer
migrating north. Use the `bccm_full` outputs.

Need to decide on depth layer, seems best to do 100m to bottom.

```{r, getoxygen}
ox <- bccm_avg100mtoBot_oxygen_full()
summary(ox$`2019_9`)
```
There are some zero values of oxygen, which are obviously wrong (and then messes
up the plotting scales). Showing that month as an example.

Check if that happens for the original BCCM outputs:
```{r, getoxygenorig
ox_not_full <- bccm_avg100mtoBot_oxygen()
summary(ox_not_full$`2019_9`)
```
No negative ones there.

Okay, dig into the negative ones, at least for Sept 2019.
```{r, ox1}
ox_2019_9 <- ox$`2019_9`    # One big vector values, one for each cell.
sum(ox_2019_9 < 0)
sum(ox_2019_9 < 10)
sum(ox_2019_9 < 15)

sort(ox_2019_9)[1:33]
```

What are the lowest values for the not-full results:
```{r, ox2}
ox_not_full_2019_9 <- ox_not_full$`2019_9`
sort(ox_not_full_2019_9)[1:33]
```

Check all the months:
```{r, ox3}
ox_not_full_no_geom <- sf::st_drop_geometry(ox_not_full)

summary(unlist(ox_not_full_no_geom))
```
So there is at least one negative one in the original BCCM output. Have not done
any wrangling, so this is in pacea.

Okay, how many:
```{r, ox4}
length(unlist(ox_not_full_no_geom))
sum(unlist(ox_not_full_no_geom) < 0)

```







```{r, plotdefault}
plot(ox)
```

```{r, subset}
# subset data based on year_month column name - reduces processing time
#ox_subset <- ox %>%
#  dplyr::select(all_of(ym))

# long format for ease of filtering
#ox_long <- pacea_long(ox_subset)

#dim(ox_long)
```

Look at a smaller region, based on Travis's BCCM vignette example
```{r, smaller}
# coordinates for polygon, points 1, 2, ... specified by  lon1, lon2, ....,
# lat1, lat2, ...
area <- list(matrix(c(-124, -120, -125, -132, -124,
                        44,   44,   52,   52,   44),
                    ncol = 2))

# create polygon object with lat-lon WGS 84 projection (4326) and convert to sf object with BC Albers (3005) projection

area_sf <- sf::st_sfc(sf::st_polygon(area), crs = 4326) %>%
  sf::st_as_sf() %>%
  sf::st_transform(crs = 3005)

# mask data by indexing original sf object with new area_sf shapefile
ox_area <- ox[area_sf, ]
class(ox_area) <- c("pacea_st", class(ox_area)) %>% unique()    # put pacea_st
# first
attr(ox_area, "restrict_plotting_range") <- TRUE

ox_area

dim(ox_area)
```

Note: co-ordinates prescribe above do not seem to exactly match locations on the grid
in these figures.

Just plotting August for each survey year, over three panel plots.

```{r, surveyyears}
survey_years <- c(1995, 1998, 2001, 2003, 2005,
                  2007, 2009, 2011, 2012, 2013,
                  2015, 2017, 2019, 2021, 2023)

```{r, plotarea1}
plot(ox_area,
     months = "August",
     years = survey_years[1:5])
```



```{r, stop, cache = FALSE}
knitr::knit_exit()
```

```{r, plotarea2}
plot(ox_area,
     months = "August",
     years = survey_years[6:10])
```

```{r, plotarea3}
plot(ox_area,
     months = "August",
     years = survey_years[11:15])
```







Using `med-analysis-13.Rmd` as a template, but needing to look at the original
`med-analysis-4.Rmd` results first.

Here (`med-analysis-15.Rmd`) we determine $x_{\text min}$ as for analysis-4,
independently for each group-strata combination, but for that $x_{\text max}$ in
a few cases looked hugely influenced by a single measurement, that was not
continuous with the previous ones, even on a log scale.

Looks like this occurs for:
 - crustacea baseline
 - ceph small baseline
 - maybe ceph large ntr
 - check others.

```{r, loaddatjoined}
dat_joined <- readRDS(paste0(here::here(),
                             "/report/mediterranean-analysis-4/dat_joined.rds"))

med_4_results_table <- readRDS(paste0(here::here(),
                             "/report/mediterranean-analysis-4/res_table_agg.rds"))
```

Also need the cephalopod small/large delineation. Juliana said to use 10, but I
think it's best to stick with the value done from the analysis in `med-analysis-4`:
```{r, loadcephdelin}
ceph_small_large_delineation <-  readRDS(paste0(here::here(),
                                                "/report/mediterranean-analysis-4/ceph_small_large_delineation.rds"))

ceph_small_large_delineation
```

## Look at analysis-4 results to see which ones to remove outliers

Need some from `med-summary-4.Rmd` to load in results. Look at that pdf to
narrow down which datasets to look at more carefully.


## Load results

Load in saved results.

```{r, loadresults}
baseline_agg_list <- readRDS(paste0(here::here(),
                             "/report/mediterranean-analysis-4/baseline_agg_list.rds"))
fg_agg_list <- readRDS(paste0(here::here(),
                              "/report/mediterranean-analysis-4/fg_agg_list.rds"))
ntr_agg_list <- readRDS(paste0(here::here(),
                               "/report/mediterranean-analysis-4/ntr_agg_list.rds"))
res_table_agg <- readRDS(paste0(here::here(),
                                "/report/mediterranean-analysis-4/res_table_agg.rds"))
```

Each is a list for that strata, with 6 objects of results of class
`determine_xmin_and_fit_mlebins`. First of each is for fitting the full
community in one g.

`res_table_agg` is the summary table that has been built up during analysis (and
users should do that by using our .Rmd as a template).


## Fits to each species group

Do not really need the histogram information in each list here, so stripping out
the histogram information here, and also removing the `"full"` analyses as do
not want to include them:

```{r, striphist}
remove_strata <- "full"   # make NULL to keep everything, or "full" to remove
                          # the full analysis. Change "Full community" below
                          # also if change this.

baseline_agg <- remove_hist(baseline_agg_list,
                            remove_strata = remove_strata)

fg_agg <- remove_hist(fg_agg_list,
                      remove_strata = remove_strata)

ntr_agg <- remove_hist(ntr_agg_list,
                       remove_strata = remove_strata)

# But probably ending up showing Full and then explaining why dig deeper
res_table_agg_no_full <- filter(res_table_agg,
                            Group != "Full community")
```

\clearpage

First, need a global range for the body sizes:
```{r, globalx}
x_min_global <- min(res_table_agg_no_full$x_min)
x_max_global <- max(res_table_agg_no_full$x_max)
groups <- unique(res_table_agg_no_full$Group)    # Assume we're keeping everything in
                                         # the same order, no reason not to
strata <- unique(res_table_agg_no_full$Strata)

```

Just looking through `med-summary-4.Rmd`, look carefully into the ones that
might have the outliers at large size. Just skip those that look fine.

Approach:

### Crustacea baseline

Look at the gap between biggest and second biggest, based on `bin_min` (looked
at `bin-max` to `bin_min` but can be negative, so just stick with `bin_min`). 

```{r, crustbase}
baseline_crust_gap <- detect_outliers(baseline_agg$crust)

baseline_crust_gap %>% tail(10)   # gap_ratio is ratio of gap to second largest

plot(baseline_crust_gap$gap_ratio)
```

Clearly continuous and then a big jump, so remove the final one.

```{r, crustbase2}
baseline_crust_rem <- remove_outliers(baseline_agg$crust)

baseline_crust_rem

baseline_crust_rem$dat_keep %>% tail()

baseline_crust_rem$dat_removed

baseline_crust_rem$bin_count_removed_prop
```

Could iterate to see if should remove the next one, this is how it could be
automated. But have to decide on a cut off.

```{r, crustbase3}
baseline_crust_gap_2 <- detect_outliers(baseline_crust_rem)

baseline_crust_gap_2 %>% tail(10)

plot(baseline_crust_gap_2$gap_ratio)
```

So there is a gap that's double the next biggest, but it's not the final bin, so
no justification in removing any more.

A rule of thumb could be a gap for the final bin of five times the next biggest gap?

### Crustacea FG

```{r, crustfgremove}
fg_crust_gap <- detect_outliers(fg_agg$crust)

fg_crust_gap %>% tail(10)

max(fg_crust_gap$gap_ratio, na.rm = TRUE)

plot(fg_crust_gap$gap_ratio)
```

Not as clear cut as baseline, but seems sensible and pragmatic to keep final
one, want to remove as little data as possible, and need to be justified when
doing so.

### Crustacea NTR

```{r, crustntrremove}
ntr_crust_gap <- detect_outliers(ntr_agg$crust)

ntr_crust_gap %>% tail(10)

max(ntr_crust_gap$gap_ratio, na.rm = TRUE)

plot(ntr_crust_gap$gap_ratio)
```

Keep. Could put in a rule for whether max ratio is $> 5$ or whatever, but just
keep doing manually for now.


Now just looking at ones that might need examining, based on plots in `med-analysis-4.Rmd`.

### Actin baseline

```{r, actinbaselineremove}
baseline_actin_gap <- detect_outliers(baseline_agg$actin)

baseline_actin_gap %>% tail(10)

max(baseline_actin_gap$gap_ratio, na.rm = TRUE)

plot(baseline_actin_gap$gap_ratio)
```

Keep.

### Chondr baseline

```{r, chondrbaselineremove}
baseline_chondr_gap <- detect_outliers(baseline_agg$chondr)

baseline_chondr_gap %>% tail(10)

max(baseline_chondr_gap$gap_ratio, na.rm = TRUE)

plot(baseline_chondr_gap$gap_ratio)
```

Keep.

### Chondr ntr

```{r, chondrntrremove}
ntr_chondr_gap <- detect_outliers(ntr_agg$chondr)

ntr_chondr_gap %>% tail(10)

max(ntr_chondr_gap$gap_ratio, na.rm = TRUE)

plot(ntr_chondr_gap$gap_ratio)
```

Keep.


### Cephsmall baseline

```{r, cephsmallbaselineremove}
baseline_cephsmall_gap <- detect_outliers(baseline_agg$cephsmall)

baseline_cephsmall_gap %>% tail(10)

max(baseline_cephsmall_gap$gap_ratio, na.rm = TRUE)

plot(baseline_cephsmall_gap$gap_ratio)
```

Yes, definitely remove.

```{r, cephsmallbase2}
baseline_cephsmall_rem <- remove_outliers(baseline_agg$cephsmall)

baseline_cephsmall_rem

baseline_cephsmall_rem$dat_keep %>% tail()

baseline_cephsmall_rem$dat_removed

baseline_cephsmall_rem$bin_count_removed_prop
```

Could iterate to see if should remove the next one, this is how it could be
automated. But have to decide on a cut off.

```{r, cephsmallbase3}
baseline_cephsmall_gap_2 <- detect_outliers(baseline_cephsmall_rem)

baseline_cephsmall_gap_2 %>% tail(10)

plot(baseline_cephsmall_gap_2$gap_ratio)
```

Keep the rest then, so just removing the one, as anticipated from original figure.

### Cephlarge baseline

```{r, cephlargebaselineremove}
baseline_cephlarge_gap <- detect_outliers(baseline_agg$cephlarge)

baseline_cephlarge_gap %>% tail(10)

max(baseline_cephlarge_gap$gap_ratio, na.rm = TRUE)

plot(baseline_cephlarge_gap$gap_ratio)
```

Keep.

### Cephlarge fg

```{r, cephlargefgremove}
fg_cephlarge_gap <- detect_outliers(fg_agg$cephlarge)

fg_cephlarge_gap %>% tail(10)

max(fg_cephlarge_gap$gap_ratio, na.rm = TRUE)

plot(fg_cephlarge_gap$gap_ratio)
```
Keep.

### Cephlarge ntr

```{r, cephlargentrremove}
ntr_cephlarge_gap <- detect_outliers(ntr_agg$cephlarge)

ntr_cephlarge_gap %>% tail(10)

max(ntr_cephlarge_gap$gap_ratio, na.rm = TRUE)

plot(ntr_cephlarge_gap$gap_ratio)
```

Yes, remove that one.

```{r, cephlargebase2}
ntr_cephlarge_rem <- remove_outliers(ntr_agg$cephlarge)

ntr_cephlarge_rem

ntr_cephlarge_rem$dat_keep %>% tail()

ntr_cephlarge_rem$dat_removed

ntr_cephlarge_rem$bin_count_removed_prop
```

Could iterate to see if should remove the next one, this is how it could be
automated. But have to decide on a cut off.

```{r, cephlargebase3}
ntr_cephlarge_gap_2 <- detect_outliers(ntr_cephlarge_rem)

ntr_cephlarge_gap_2 %>% tail(10)

plot(ntr_cephlarge_gap_2$gap_ratio)
```

Keep the rest then, so just removing the one, as anticipated from original figure.

## Summary of outlier analyses

So removing the following (`$dat_keep` will be what we keep), and also the small
outlier we manually removed in `med-analysis-4.Rmd` but didn't need to for later
analyses (since xmin was fixed in those).
```{r, removingthese}
baseline_crust_rem$dat_removed

baseline_cephsmall_rem$dat_removed

ntr_cephlarge_rem$dat_removed
```

Also, in `med-analysis-4.Rmd` we removed the small outlier that had `bin_min =
0.12`, and need to here. It was Aegaeon lacazei, crustacea in ntr, can just
filter $>0.35$.


So remove those from the full analyses also, and technically only need to rerun
these ones as other results won't change, but a bit simpler to just rerun
everything, after wrangling the data for these. Gets a bit fiddly and
situation-specific to mix old and new results.

So just need to do slightly different runs to `med-analysis-4.Rmd`normal for:

- baseline full
- ntr full
- baseline crust
- baseline cephsmall
- ntr cephlarge

Estimating $x_{\text min}$ for each scenario:

```{r setxmin}
xmin_full <- NULL

xmin_crust <- NULL

xmin_actin <- NULL

xmin_chondr <- NULL

xmin_cephsmall <- NULL

xmin_cephlarge <- NULL
```

Now to tease out the $x_{\text max}$ values for this particular analysis, just
estimating each time.
```{r setxmax}
# strata_for_xmax <- "Baseline"
if(TRUE){
  # Determine xmax independently for each data set
  xmax_full <- NULL
  xmax_crust <- NULL
  xmax_actin <- NULL
  xmax_chondr <- NULL
  xmax_cephsmall <- NULL
  xmax_cephlarge <- NULL
} else {
  # Use max of xmax within each group
  xmax_full <- filter(med_4_results_table,
                      Group == "Full community")$x_max %>% max()

  xmax_crust <- filter(med_4_results_table,
                       Group == "Crustacea")$x_max %>% max()

  xmax_actin <- filter(med_4_results_table,
                       Group == "Actinopterygii")$x_max %>% max()

  xmax_chondr <- filter(med_4_results_table,
                        Group == "Chondrichthyes")$x_max %>% max()

  xmax_cephsmall <- filter(med_4_results_table,
                           Group == "Cephalopoda small")$x_max %>% max()

  xmax_cephlarge <- filter(med_4_results_table,
                           Group == "Cephalopoda large")$x_max %>% max()
}

```

\clearpage

Copying then editing the next stuff from `med-analysis-13.Rmd` (rather than
med-4) because likely more streamlined.

`mediterranean_for_mlebins()` combines duplicated bins, but don't need that
function for ones we already have above, so when doing ones from above will
combine manually and check the tibble sizes. Also need to remove relevant data
from full community analyses.

# Full community

Doing full community of values, which is okay because the sampling protocol
(trawl fishing) is the same
for each species group and the counts are standardised to be numbers per
km$^2$. For species groups, copying and pasting the analyses and then just
changing the group names (need to see everything and not worth it to functionalise
it all).

## Full Community baseline (takes a while)

```{r, fullbaseline}
# This function extracts the required data and returns what is needed for
# MLEbins; keeps species column for understanding even though not strictly
# needed for MLEbins.
data_full_baseline <- mediterranean_for_mlebins(dat_joined,
                                                strata_name = "baseline")

# Reduce:
data_full_baseline_red <- dplyr::setdiff(data_full_baseline,
                                         rbind(baseline_crust_rem$dat_removed,
                                               baseline_cephsmall_rem$dat_removed))
# Double check:
data_full_baseline
data_full_baseline_red

res_full_baseline <- determine_xmin_and_fit_mlebins(data_full_baseline_red,
                                                    x_min = xmin_full,
                                                    x_max = xmax_full)

res_full_baseline$mlebins_fit      # full list includes histogram values which
                                   # are not a tibble (am printing them all for
                                   # some later species groups).
```

```{r, fullbaselineplot}
plot(res_full_baseline,
     xlim_hist = c(0, 60),
     main = "Full Community baseline",
     seg_col = "black")

```

## Full Community fg

```{r, fullfg}
data_full_fg <- mediterranean_for_mlebins(dat_joined,
                                          strata_name = "fg")

res_full_fg <- determine_xmin_and_fit_mlebins(data_full_fg,
                                              x_min = xmin_full,
                                              x_max = xmax_full)

res_full_fg$mlebins_fit      # full list includes histogram values which
                                   # are not a tibble (am printing them all for
                                   # some later species groups).
```

```{r, fullfgplot}
plot(res_full_fg,
     xlim_hist = c(0, 60),
     main = "Full Community FG",
     seg_col = "black")
```


## Full Community ntr


Now do the analysis without the outlier:
```{r, fullntr}
data_full_ntr <- mediterranean_for_mlebins(dat_joined,
                                           strata_name = "ntr") %>%
    dplyr::filter(bin_min > 0.35)

# Reduced:
data_full_ntr_red <- dplyr::setdiff(data_full_ntr,
                                    ntr_cephlarge_rem$dat_removed)

res_full_ntr <- determine_xmin_and_fit_mlebins(data_full_ntr_red,
                                               x_min = xmin_full,
                                               x_max = xmax_full)

res_full_ntr$mlebins_fit      # full list includes histogram values which
                              # are not a tibble (am printing them all for
                              # some later species groups).
```

```{r, fullntrplot}
plot(res_full_ntr,
     xlim_hist = c(0, 60),
     main = "Full Community NTR",
     seg_col = "black")
```


## Compare strata for full community

Now to compare the three strata for full community.

Keep defined here, as fairly specific to this data set.
```{r, fullsumm}
add_row_to_summary_table <- function(summary, res, group, strata){
  rbind(summary,
        dplyr::tibble("Group" = group,
                      "Strata" = strata,
                      "x_min" = res$mlebins_fit$x_min,
                      "x_max" = res$mlebins_fit$x_max,
                      "Low b" = res$mlebins_fit$b_conf[1],
                      "MLE b" = res$mlebins_fit$b_mle,
                      "High b" = res$mlebins_fit$b_conf[2]))
}

res_table_full <- dplyr::tibble()
res_table_full <- add_row_to_summary_table(res_table_full,
                                      res_full_baseline,
                                      "Full community",
                                      "Baseline")

res_table_full <- add_row_to_summary_table(res_table_full,
                                      res_full_fg,
                                      "Full community",
                                      "FG")
res_table_full <- add_row_to_summary_table(res_table_full,
                                      res_full_ntr,
                                      "Full community",
                                      "NTR")
```

Results combined are:

```{r, resfulltable}
knitr::kable(res_table_full, digits = 2)
```

# Crustacea

## Crustacea baseline

```{r, crustbaseline}
data_crust_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Crustacea",
                                                 strata_name = "baseline")

data_crust_baseline_red <- dplyr::setdiff(data_crust_baseline,
                                          baseline_crust_rem$dat_removed)


res_crust_baseline <- determine_xmin_and_fit_mlebins(data_crust_baseline_red,
                                                     x_min = xmin_crust,
                                                     x_max = xmax_crust)

res_crust_baseline$mlebins_fit
```

```{r, crustbaselineplot}
plot(res_crust_baseline,
     xlim_hist = c(0, 60),
     main = "Crustacea baseline")
```

## Crustacea FG

```{r, crustfg}
data_crust_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Crustacea",
                                           strata_name = "fg")

res_crust_fg <- determine_xmin_and_fit_mlebins(data_crust_fg,
                                               x_min = xmin_crust,
                                               x_max = xmax_crust)


res_crust_fg$mlebins_fit
```

```{r, crustfgplot}
plot(res_crust_fg,
     xlim_hist = c(0, 60),
     main = "Crustacea FG")
```

## Crustacea NTR

As for full community NTR, need to remove Aegaeon lacazei as an outlier
(including it gives similar plot here to that earlier).

```{r, crustntr}
data_crust_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Crustacea",
                                            strata_name = "ntr") %>%
    dplyr::filter(bin_min > 0.35)

res_crust_ntr <- determine_xmin_and_fit_mlebins(data_crust_ntr,
                                               x_min = xmin_crust,
                                               x_max = xmax_crust)

res_crust_ntr$mlebins_fit
```

```{r, crustntrplot}
plot(res_crust_ntr,
     xlim_hist = c(0, 60),
     main = "Crustacea NTR")
```

Now to compare the three strata for Crustacea.

```{r, crustsumm}
res_table_crust <- dplyr::tibble()
res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_baseline,
                                      "Crustacea",
                                      "Baseline")

res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_fg,
                                      "Crustacea",
                                      "FG")
res_table_crust <- add_row_to_summary_table(res_table_crust,
                                      res_crust_ntr,
                                      "Crustacea",
                                      "NTR")
```

Results combined are:

```{r, rescrusttable}
knitr::kable(res_table_crust, digits = 2)
```

# Actinopterygii

## Actinopterygii baseline

```{r, actinbaseline}
data_actin_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Actinopterygii",
                                                 strata_name = "baseline")

res_actin_baseline <- determine_xmin_and_fit_mlebins(data_actin_baseline,
                                               x_min = xmin_actin,
                                               x_max = xmax_actin)


res_actin_baseline$mlebins_fit
```

```{r, actinbaselineplot}
plot(res_actin_baseline,
     xlim_hist = c(0, 60),
     main = "Actinopterygii baseline")
```

## Actinopterygii FG

```{r, actinfg}
data_actin_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Actinopterygii",
                                           strata_name = "fg")

res_actin_fg <- determine_xmin_and_fit_mlebins(data_actin_fg,
                                               x_min = xmin_actin,
                                               x_max = xmax_actin)

res_actin_fg$mlebins_fit
```

```{r, actinfgplot}
plot(res_actin_fg,
     xlim_hist = c(0, 60),
     main = "Actinopterygii FG")
```

## Actinopterygii NTR

```{r, actinntr}
data_actin_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Actinopterygii",
                                            strata_name = "ntr")

res_actin_ntr <- determine_xmin_and_fit_mlebins(data_actin_ntr,
                                               x_min = xmin_actin,
                                               x_max = xmax_actin)

res_actin_ntr$mlebins_fit
```

```{r, actinntrplot}
plot(res_actin_ntr,
     xlim_hist = c(0, 60),
     main = "Actinopterygii NTR")
```

Now to compare the three strata for Actinopterygii.

```{r, actinsumm}
res_table_actin <- dplyr::tibble()
res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_baseline,
                                      "Actinopterygii",
                                      "Baseline")

res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_fg,
                                      "Actinopterygii",
                                      "FG")
res_table_actin <- add_row_to_summary_table(res_table_actin,
                                      res_actin_ntr,
                                      "Actinopterygii",
                                      "NTR")
```

Results combined are:
```{r, resactintable}
knitr::kable(res_table_actin, digits = 2)
```

TODO From the baseline, the size spectrum steepens (more
negative $b$) in the fishing grounds but also for the no-take reserve but not as
much (unlike for the Crustacea above).


\clearpage

# Chondrichthyes

## Chondrichthyes baseline

```{r, chondrbaseline}
data_chondr_baseline <- mediterranean_for_mlebins(dat_joined,
                                                 group_name = "Chondrichthyes",
                                                 strata_name = "baseline")

res_chondr_baseline <- determine_xmin_and_fit_mlebins(data_chondr_baseline,
                                                      x_min = xmin_chondr,
                                                      x_max = xmax_chondr)

res_chondr_baseline$mlebins_fit
```

```{r, chondrbaselineplot}
plot(res_chondr_baseline,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes baseline")
```

## Chondrichthyes FG

```{r, chondrfg}
data_chondr_fg <- mediterranean_for_mlebins(dat_joined,
                                           group_name = "Chondrichthyes",
                                           strata_name = "fg")

res_chondr_fg <- determine_xmin_and_fit_mlebins(data_chondr_fg,
                                                x_min = xmin_chondr,
                                                x_max = xmax_chondr)


res_chondr_fg$mlebins_fit
```

```{r, chondrfgplot}
plot(res_chondr_fg,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes FG")
```

## Chondrichthyes NTR

```{r, chondrntr}
data_chondr_ntr <- mediterranean_for_mlebins(dat_joined,
                                            group_name = "Chondrichthyes",
                                            strata_name = "ntr")

res_chondr_ntr <- determine_xmin_and_fit_mlebins(data_chondr_ntr,
                                                 x_min = xmin_chondr,
                                                 x_max = xmax_chondr)

res_chondr_ntr$mlebins_fit
```

```{r, chondrntrplot}
plot(res_chondr_ntr,
     xlim_hist = c(0, 300),
     main = "Chondrichthyes NTR")
```

Now to compare the three strata for Chondrichthyes.

```{r, chondrsumm}
res_table_chondr <- dplyr::tibble()
res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_baseline,
                                      "Chondrichthyes",
                                      "Baseline")
res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_fg,
                                      "Chondrichthyes",
                                      "FG")
res_table_chondr <- add_row_to_summary_table(res_table_chondr,
                                      res_chondr_ntr,
                                      "Chondrichthyes",
                                      "NTR")
```

Results combined are:
```{r, reschondrtable}
knitr::kable(res_table_chondr, digits = 2)
```


# Cephalopoda (all) data

Need to split the Cephaloda data (see `med-analysis-4.Rmd`). First load it all.

```{r, cephload}
data_cephall_baseline <- mediterranean_for_mlebins(dat_joined,
                                                   group_name = "Cephalopoda",
                                                   strata_name = "baseline")

data_cephall_fg <- mediterranean_for_mlebins(dat_joined,
                                             group_name = "Cephalopoda",
                                             strata_name = "fg")

data_cephall_ntr <- mediterranean_for_mlebins(dat_joined,
                                              group_name = "Cephalopoda",
                                              strata_name = "ntr")

ceph_small_large_delineation
```

## Cephalopoda (small) baseline

```{r, cephsmallbaseline}
data_cephsmall_baseline <- dplyr::filter(data_cephall_baseline,
                                         bin_min < ceph_small_large_delineation)

data_cephsmall_baseline_red <- dplyr::setdiff(data_cephsmall_baseline,
                                              baseline_cephsmall_rem$dat_removed)


max(data_cephsmall_baseline_red$bin_min)

res_cephsmall_baseline <-
  determine_xmin_and_fit_mlebins(data_cephsmall_baseline_red,
                                 x_min = xmin_cephsmall,
                                 x_max = xmax_cephsmall)
```

```{r, cephsmallbaselineplot}
plot(res_cephsmall_baseline,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) baseline")
```

## Cephalopoda (small) FG

```{r, cephsmallfg}
data_cephsmall_fg <- dplyr::filter(data_cephall_fg,
                                   bin_min < ceph_small_large_delineation)

res_cephsmall_fg <- determine_xmin_and_fit_mlebins(data_cephsmall_fg,
                                                   x_min = xmin_cephsmall,
                                                   x_max = xmax_cephsmall)
```

```{r, cephsmallfgplot}
plot(res_cephsmall_fg,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) FG")
```

## Cephalopoda (small) NTR

```{r, cephsmallbaeline}
data_cephsmall_ntr <- dplyr::filter(data_cephall_ntr,
                                    bin_min < ceph_small_large_delineation)

res_cephsmall_ntr <- determine_xmin_and_fit_mlebins(data_cephsmall_ntr,
                                                    x_min = xmin_cephsmall,
                                                    x_max = xmax_cephsmall)
```

```{r, cephsmallntrplot}
plot(res_cephsmall_ntr,
     xlim_hist = c(0, 21),
     main = "Cephalopoda (small) NTR")
```

Now to compare the three strata for Cephalopoda (small):

```{r, cephsmallsumm}
res_table_cephsmall <- dplyr::tibble()
res_table_cephsmall <- add_row_to_summary_table(res_table_cephsmall,
                                      res_cephsmall_baseline,
                                      "Cephalopoda small",
                                      "Baseline")

res_table_cephsmall <- add_row_to_summary_table(res_table_cephsmall,
                                      res_cephsmall_fg,
                                      "Cephalopoda small",
                                      "FG")
res_table_cephsmall <- add_row_to_summary_table(res_table_cephsmall,
                                      res_cephsmall_ntr,
                                      "Cephalopoda small",
                                      "NTR")
```

Results combined are:

```{r, rescephsmalltable}
knitr::kable(res_table_cephsmall, digits = 2)
```

## Cephalopoda (large) baseline

```{r, cephlargebaseline}
data_cephlarge_baseline <- dplyr::filter(data_cephall_baseline,
                                         bin_min >= ceph_small_large_delineation)

res_cephlarge_baseline <-
  determine_xmin_and_fit_mlebins(data_cephlarge_baseline,
                                 x_min = xmin_cephlarge,
                                 x_max = xmax_cephlarge)
```


```{r, cephlargebaselineplot}
plot(res_cephlarge_baseline,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) baseline")
```

## Cephalopoda (large) FG

```{r, cephlargefg}
data_cephlarge_fg <- dplyr::filter(data_cephall_fg,
                                   bin_min >= ceph_small_large_delineation)

res_cephlarge_fg <- determine_xmin_and_fit_mlebins(data_cephlarge_fg,
                                                   x_min = xmin_cephlarge,
                                                   x_max = xmax_cephlarge)
```

```{r, cephlargefgplot}
plot(res_cephlarge_fg,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) FG")
```

## Cephalopoda (large) NTR

```{r, cephlargebaeline}
data_cephlarge_ntr <- dplyr::filter(data_cephall_ntr,
                                    bin_min >= ceph_small_large_delineation)

data_cephlarge_ntr_red <- dplyr::setdiff(data_cephlarge_ntr,
                                         ntr_cephlarge_rem$dat_removed)

res_cephlarge_ntr <- determine_xmin_and_fit_mlebins(data_cephlarge_ntr_red,
                                                    x_min = xmin_cephlarge,
                                                    x_max = xmax_cephlarge)
```

```{r, cephlargentrplot}
plot(res_cephlarge_ntr,
     xlim_hist = c(0, 100),
     main = "Cephalopoda (large) NTR")
```

Now to compare the three strata for Cephalopoda (large):

```{r, cephlargesummfix}
res_table_cephlarge <- dplyr::tibble()
res_table_cephlarge <- add_row_to_summary_table(res_table_cephlarge,
                                      res_cephlarge_baseline,
                                      "Cephalopoda large",
                                      "Baseline")

res_table_cephlarge <- add_row_to_summary_table(res_table_cephlarge,
                                      res_cephlarge_fg,
                                      "Cephalopoda large",
                                      "FG")
res_table_cephlarge <- add_row_to_summary_table(res_table_cephlarge,
                                      res_cephlarge_ntr,
                                      "Cephalopoda large",
                                      "NTR")
```

Results combined are:

```{r, rescephlargetable}
knitr::kable(res_table_cephlarge, digits = 2)
```
# Combining results

First, let's create a table of the main results from above that we want to move
forward with. Just from the `res_table_***` objects. Then we'll need the raw
data from each in turn. Could then functionalise this maybe, but probably best
to keep explicit and tailored for different applications. Aggregated plots done
in `med-analysis-**.Rmd`.

```{r, aggregatetable}
res_table_agg <- rbind(res_table_full,
                       res_table_crust,
                       res_table_actin,
                       res_table_chondr,
                       res_table_cephsmall,
                       res_table_cephlarge)


knitr::kable(res_table_agg,
             digits = 2)
```

## Aggregated fit to data

Now just saving the results for each strata in a list. (See end of file for what
I originally did). Can include the full results (fitting in one go). Not going
to change filenames for each analysis (`med-analysis-6.Rmd` etc.), files will be in the correct directory

Each object in each list is of class `determine_xmin_and_fit_mlebins`.

```{r, aggregatefitnew}
baseline_agg_list <- list(full = res_full_baseline,
                          crust = res_crust_baseline,
                          actin = res_actin_baseline,
                          chondr = res_chondr_baseline,
                          cephsmall = res_cephsmall_baseline,
                          cephlarge = res_cephlarge_baseline)

fg_agg_list <- list(full = res_full_fg,
                    crust = res_crust_fg,
                    actin = res_actin_fg,
                    chondr = res_chondr_fg,
                    cephsmall = res_cephsmall_fg,
                    cephlarge = res_cephlarge_fg)


ntr_agg_list <- list(full = res_full_ntr,
                     crust = res_crust_ntr,
                     actin = res_actin_ntr,
                     chondr = res_chondr_ntr,
                     cephsmall = res_cephsmall_ntr,
                     cephlarge = res_cephlarge_ntr)
```

## Saving what is needed for plots and tables, and doing in med-summary.Rmd

```{r, savingnew}
saveRDS(baseline_agg_list,
        "baseline_agg_list.rds")
saveRDS(fg_agg_list,
        "fg_agg_list.rds")
saveRDS(ntr_agg_list,
        "ntr_agg_list.rds")

saveRDS(res_table_agg,
        "res_table_agg.rds")

```
