---
title: "Buoy Sea Surface Temperature Anomalies"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Buoy Sea Surface Temperature Anomalies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("buoys-from-vignette.Rmd")
or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 14,   #5.7,
  fig.height = 7  #5
)
```

## Daily average sea surface temperature from buoys - climatologies and anomalies

```{r setup}
load_all()
# library(pacea)
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In pacea we include daily average sea surface temperature (SST) that we have calculated from
measurement from `r nrow(buoy_metadata)` buoys in
Canadian Pacific waters, yielding over 200,000 daily means.
Data are from Environment and Climate Change Canada, and Fisheries and
Oceans Canada. The earliest data are from September 1987, and 18 buoys were
still providing data as of August 2026.

See the
[buoys_vignette](https://pbs-assess.github.io/pacea/vignettes/buoys.html) and
the help file `?buoy_sst` 
for background concerning the calculated daily mean sea surface temperatures
and associated simple plotting functions in pacea. Here we focus on
constructing climatologies and subsequent anomalies, to plot in 
a standardised way.

TODO rerun that vignette with the new order.

The locations of the buoys are:
```{r buoysmap, echo=FALSE, fig.width = 5.7, fig.height = 5, out.width = "90%"}
knitr::include_graphics(paste0(here::here(),
                               "/man/figures/buoys_map.png"))
```

If you want to quickly get a sense of current temperatures (in the wider
  northeast Pacific also), check out Andrea Hilborn et al.'s
  [Pacific_SST_Monitoring website](https://ios-osd-dpg.github.io/Pacific_SST_Monitoring_Site/)
  page. The extracted buoy SST in pacea were wrangled with the help of
  Andrea. pacea is more suited for doing your own tailored plots and analyses.

## Calculations and plotting

### Default plot

We can easily calculate a climatology (by month) and subsequent anomalies for
each buoy, and plot results:
```{r, buoyanomalies}
all_buoys_anomalies <- calculate_anomaly(buoy_sst)

plot(all_buoys_anomalies)
```

The plot's title describes that this default plot is for the anomalies **in April** for all
buoys, and the climatology was 1991-2020 where
available. Buoys are arranged in a roughly north-to-south order (see map). So each square
gives the SST anomaly for that buoy for April of that year. Empty (white)
squares indicate not enough data available for that particular calculation.

The figure clearly shows that April 2015 and April 2016 were warmer than usual
for April, for each buoy. The bottom three buoys clearly have limited data. 

### Absolute SST's

To show the actual monthly average SST's (rather than anomalies), again for the
default month of April:
```{r, buoyanomalies3}
plot(all_buoys_anomalies,
     sst_plot = "mean")
```
Here you can roughly see the water getting warmer rom north to south, noting
that the three "NOMAD" buoys are much further offshore than the others, and
look slightly cooler.

### Full results for a specific buoy

If you specify a buoy, you get the monthly anomalies across the years (again
with an automated title):
```{r, buoyanomalies2}
plot(all_buoys_anomalies,
     stn_id = "Halibut Bank")   # In the Strait of Georgia
```

You can check that the April 2016 high anomaly matches that shown in the first
anomaly plot shown above.

### Anomalies for all buoys for different months

You can specify a single month (recall the default is 4 for April):
```{r, buoyanomaliesmonth}
plot(all_buoys_anomalies,
     months = 12)
```

Interestingly, in 2023 the December anomalies are all positive, whereas in the
first anomaly plot above the April anomalies are all negative (for the available
buoys in each case).

If you specify multiple consecutive months (e.g. `4:6`), then the plot shows the anomalies
averaged over those months (and the title is automatically correct):
```{r, buoyanomaliesmonths}
plot(all_buoys_anomalies,
     months = 4:6)
```

TODO HERE (when doing 4:6) Look at South Nomad and plot it's full data. 2014 only has April (out
of Apr-Jun) and 2016 has no April, hence the anomalies are a bit
misleading. Suggest we do insist that all months have values, or at least give
an option to require a certian proportion, default 100%. 
Do an example with months 1:12.

And if you want a wintertime index that includes December, that can be done:
```{r, buoyanomaliesmonths2}
plot(all_buoys_anomalies,
     months = c(12, 1, 2, 3))
```

As stated in the title, the year on the x-axis is the year of the January.

IMPORTANT: Note that for these plots with multiple months, we have not included
a criteria that each of the requested months has data (or at least some
proportion of them do). It is a bit tricky to come up with rules for all
scenarios. Each plotted anomaly is just the mean of the appropriate available monthly
anomalies.

TODO move to later after have talked about the outputs.

Example of no climatology for some months for S. Georgia Strait buoy, so when
do:
```{r, buoyanomaliesmonths46}
plot(all_buoys_anomalies,
     months = 4:6)
```

we lose that buoy in the plot because it does not have a climatology for Jun
therefore we cannot do a mean climatology for Apr-Jun. 

The climatology is:

```{r, noclimatology}
all_buoys_anomalies$climatology %>%
  filter(stn_id == "C46303") %>%
  a()
```
and the anomalies for 2022 (as an example) are:
```{r, noclimatology2}
all_buoys_anomalies$anomaly %>%
  filter(year == 2022,
         stn_id == "C46303") %>%
  a()
```
where you can see that Jun and Jul have SST values, but no anomaly because there
is no climatology for those months for this buoy (due to limited data, and the
default climatology years of 1991-2020). 

However, if we specify that we only require 2 months to be available to make an anomaly for Apr-Jun
```{r, buoyanomaliesmonths46needless}
plot(all_buoys_anomalies,
     months = 4:6,
     require_requested_months = 2)
```





TODO may have to move:

If you specify `months` and a single buoy, the anomalies for those months are
all shown:
```{r, buoyanomaliesmonths3}
plot(all_buoys_anomalies,
     months = c(12, 1, 2, 3),
     stn_id = "Halibut Bank")  
```

This plot can be used as a check to see what anomalies are available (if you are
only interested in a particular buoy). Email Andy or make an Issue if you would
like some automated criteria regarding averaging anomalies over multiple
months). From that plot it looks like removing years that do not have all four
the requested months would not lose too many years. But other buoys and/or month
sequences may not be as forgiving.

### Returning the values shown in the plot

Setting `return_results = TRUE` returns a list containing the plot and then the
results used for the plot. For the previous plot:

```{r, buoyanomaliesmonths4}
res <- plot(all_buoys_anomalies,
     months = c(12, 1, 2, 3),
     stn_id = "Halibut Bank", return_results = TRUE)  
res
```

So `res$plot` will show the above plot again, and `res$results` shows the a
tibble that includes the values
in the plot, as `sst_plot_value` and then `label` for the rounded values that are shown
numerically. You can then use these for analyses if desired.

That tibble contains results that depend on options used in the plotting. But we
will now examine the full results generated at the very beginning in `all_buoys_anomalies`.

### Examining the climatology and anomalies

HERE talk about results.

```{r, buoyanomaliesdetails}
all_buoys_anomalies
```
This is a list object, with `$climatology` giving, for each buoy and month, a mean SST
(`clim_sst_mean`) and standard deviation (`clim_sst_sd`) and sample size
(`clim_sst_n`) of the values used for that buoy-month combination.

The default for `calculate_anomaly()` creates a monthly climatology for each
buoy based on the years 1991-2020 (details of other arguments will be
investigated later).

The `$anomaly` tibble gives, for each buoy-year-month combination, the anomaly
for that month compared to the climatology. The first buoy shown is "C46184"
which is the `r buoy_metadata$name[match("C46184", buoy_metadata$stn_id)]`
buoy. For October 1987 the mean SST was 10.1 degC (`sst_mean`), which was
calculated from 29 days (`sst_n`) in that month. The resultig anomaly from the
climatology is -1.16 (`sst_anomaly`). This can be manually checked -- the
climatology for buoy C46184 in October is a mean SST of 11.3 (the top tibble),
the mean SST in October 1987 was 10.1, which is -1.16 less (`sst_anomaly`). This
should help you understand how calculations are done. 

So the same calculations are done for all buoy-year-month combinations. There
happen to be a few NA's early on in the data for this buoy (more on that later).

The vector `$climatology_years` gives the years prescribed for the
climatology in the `calculate_anomaly()` call, for which the default is `1991:2020`.

The attribute given just tells R to use our customised plotting
function on these results, which is automatically called with `plot()`:


TODO confirm understanding - average the Apr anomalies for ECOBUOY, should be 0 as
they comprise the full climatology.


```{r, stop33, cache = FALSE}
knitr::knit_exit()
```



```{r, saveplotmaybe, echo = FALSE}
# ggsave(filename = "Annual_heatmap_SST_anomaly_SOPOregions_allyears.png", width = 9, height = 3., dpi = 300,
#       scale = 1.4)
```

How about the average anomaly for April to June for all buoys:
```{r, buoyanomalyplot2}
plot(all_buoys_anomalies,
     months = 4:6)
```

Other options, to add to the help and the testing:
```{r, buoyanomalyplot4}
plot(all_buoys_anomalies,
     sst_plot = "mean",
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     months = c(12, 1, 2))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184")

plot(all_buoys_anomalies,
     sst_plot = "count")  # check what this is doing, mean
# count over all years?? Ah no, it's april counts each year each buoy.

plot(all_buoys_anomalies,
     sst_plot = "count",
     count_breaks = c(0, 20, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184",
     count_breaks = c(0, 15, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8,
     stn_id = "C46184")
```


All buoys, but no 15-day requirement in the climatology or anomaly, instead use
all days:
```{r, usealldays
all_buoys_anomalies_all_days <- calculate_anomaly(buoy_sst,
                                                  min_days_per_month = 0,
                                                  max_consecutive_missing_days =
                                                  32)
plot(all_buoys_anomalies_all_days)

# expect_equal(all_buoys_anomalies, all_buoys_anomalies_all_days) worked when
# first arg used the original function before we added this in
```

And here is the default with the cut off of requiring 15 days:
```{r, need15days}
plot(all_buoys_anomalies)
```

For example, Central Dixon Entrance has a few years that appear when using all
days but not when using the 15-day-per-month cut off (remember above two plots
are for April). Can look at that in
detail, by using the `min_days_per_month = 0` results:
```{r, cde}
plot(all_buoys_anomalies_all_days,
     stn_id = "C46145",
     sst_plot = "count",
     count_breaks = c(0, 15, 31))
```

Now doing the max consecutive missing days:
```{r, maxmissing}
all_buoys_anomalies_all_gaps_allowed <- calculate_anomaly(buoy_sst,
                                                  max_consecutive_missing_days =
                                                    32)
# expect_equal(all_buoys_anomalies, all_buoys_anomalies_all_gaps_allowed) worked when
# first arg used the original function before we added this in
```

Need to confirm that does what we want.
Buoy C46004 has NA's from 1989-11-06 to 1990-07-11, also a run of 7 in August 1990. Let's force 1989 to be in
climatology and check it those end months get properly ignored.

```{r, climtest}
test_buoy <- filter(buoy_sst,
                    stn_id == "C46004")
test_buoy_anomalies <- calculate_anomaly(test_buoy,
                                         climatology_years = 1989:1995)
plot(test_buoy_anomalies)

test_buoy %>% filter(date < "1992-10-01") %>% a()

test_buoy_anomalies$anomaly[1:50,] %>% a()
```
That shows some that are wrong. For example
```{r, wrong}
test_buoy_march_april_1991 <-
  filter(test_buoy,
         date >= "1991-03-01" & date <= "1991-04-30")

# Can see a streak of 3 NA's and one of 4 NA's, with 24 non-NA's in March, but
# April just has the one NA ...
test_buoy_march_april_1991 %>% a()  

# ... so the default settings do work, but if we get stricter then they do not
# for March:
calculate_anomaly(test_buoy_march_april_1991)
calculate_anomaly(test_buoy_march_april_1991,
                  min_days_per_month = 25)
calculate_anomaly(test_buoy_march_april_1991,
                  max_consecutive_missing_days = 4) # 4 NA streak allowed
calculate_anomaly(test_buoy_march_april_1991,
                  max_consecutive_missing_days = 3) # no climatology

    
```

Also maybe do some of (this may be okay now, was while figuring out gaps)
Why is C46004 March 1991 getting excluded? Should not be. Also:
 C46004 1992     1       NaN     0          NaN  
 C46004 1992     7       NaN     0          NaN   Though if max gaps is 7 then
 it should not be excluded, check.


Workflow should be:
- do the calculations with the defaults
- plot the counts for that buoy, with count_breaks as 0,15,31 (could make the
  default)
- check if that looks satisfactory, both in terms of calculating the climatology
  and for doing the anomalies. Hard to automate, so just have to inspect
  manually (e.g. not too many big gaps).
- change `min_days_per_month` value if possible. Try 0 at first as an extreme to see if that helps enough.

Now add max consecutive number of NA's in a row, within a month-year
combination. And suggest that check also.

Do single buoys also.

We can exclude the three buoys that do not have a long time series (this might
change when we refine the analyses):

don't need I think:

```{r, buoyanomalyplot3}

all_buoys_plot_3 <- plot(all_buoys_anomalies,
                         months = 4:6)
all_buoys_plot_3
```

Can do various options. 

Have averaged the anomalies for multiple months.


### Single buoy -- show each month

And if just one stn_id then plot months on the y-axis, January at the top:

```{r, singlebuoy}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206")
```
Hopefully self explanatory. 

Thinking of quick look at a relevant buoy for hake migration north with respect
to hake catches in Canada (which aren't in pacea but could be):
```{r, singlebuoy2}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206",
     months = 4:6)
```
If hake catches were added on then would be easy to spot any pattern.

Get the plots working when `climatology_time = "week"`.



### Refining the analyses

Important to be aware of:

Above all assumes our defaults in the calculations as specified in
`calculate_anomaly.pacea_buoy()`. These are somewhat subjective and can be
changed by the user. Specifically:

- the number of days needed within a calendar month to use to give a monthly
  average for calculating climatoloties and anomalies (default is 15 days)
- the maximum consecutive number of missing days allowed within a calendar
  month, above which that month is not used to give a monthly
  average for calculating climatoloties and anomalies (default is 6 days)
  
We pragmatically decided on requiring 15 days (roughly half) of daily SST within
a month, and a maximum consecutive number of missing days of 6. These are the
defaults, but can be changed by the user (as described below) to `NULL` to allow
all data and then examine it for plotting.

Similarly, if a user wants to look at average SST for a buoy over the months
Jan-Jun, a decision has to be made as to how many months should be
available. Could say a default of >75%, with no details about requiring
consecutive ones. Maybe they should just look at the data and decide.

Given the huge number of options that users may wish to explore (many buoys,
many years, many possible combinations of months), users may wish to relax or
tighten the defaults, and have the power to do so.



### Daily swings

Issue #26 regards what to do with big daily swings that might be air temps
instead, had thought these would only come after some NA's or at the start of a
buoy's time series. Was going to take a look, but looking at the actual swings
we have in the daily data (grouping by buoy):

```{r, dailyswings}
max_daily_swing_by_buoy <- buoy_sst %>%
  group_by(stn_id) %>%
  mutate(daily_increase = sst - lag(sst)) %>%
  group_by(stn_id) %>%
  summarise(max_daily_increase = max(daily_increase,
                                     na.rm = TRUE),
            max_daily_decrease = min(daily_increase,
                                     na.rm = TRUE))

max_daily_swing_by_buoy
summary(max_daily_swing_by_buoy)
```
it looks like we have nothing unreasonable. Asked Andrea on #26 so see what she
says, but think we're all good and can close that.


```{r, stop, cache = FALSE}
knitr::knit_exit()
```



### Rough notes, being adopted above somewhat


This works, but not for plotting yet, and may need to add options for how many
days needed in a week.

all_buoys_anomalies_weekly <- calculate_anomaly(buoy_sst, climatology_time = "week")

Need to remove data-poor months. Have some defaults that users can change, and
tell users to look at data. Elise thinks at least half the month, with no gap of
more than 5-6 days (no consecutive NA's longer than that).
Example we had was

```{r ex}
> yy <- table(all_buoys_anomalies$anomaly$sst_n)
> yy

   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14   15 
 777   32   19   29   26   24   29   21   24   26   26   26   36   19   34   34 

  16   17   18   19   20   21   22   23   24   25   26   27   28   29   30   31 
  57   68   72   59   76   69   94  113  149  171  346  301  727  682 1341 1798 

sum(as.data.frame(yy)$Freq)
[1] 7305
> sum(as.data.frame(yy)[21:32, ]$Freq)   # How many overall months have >=20
days of values.
[1] 5867
> 
```

The one showed up when we were playing around was:
filter(buoy_sst, stn_id == "C46183", year(date) == 2019, month(date) == 6) %>% a()
Has 13 missing days but they are kind of spread out, so looks usable. Hence go
with half a month needed, but no big gaps.

plot(all_buoys_anomalies, months = 1:12)  # North Hecate St 2019 has a big
anomaly, but it's only from that one month of data, as seen clearly in:
plot(all_buoys_anomalies, stn_id = "C46183")

Also need to think about climatologies. And whether to put the thinking part in
the climatology calculations and/or the plotting. Have to be in the climatology
calculations, as anomalies will change when we change the rules. And then all
resulting monthly climatologies/anomalies will be valid.

In the plotting we need a condition - how many months out of the requested ones
need to have anomaly calculations? Since monthly anomalies will have been calculated.

Could do the heatmap type plot on weekly or even daily values (for a single
buoy) to get an idea of the amount of missing data. Daily would be interesting.


Need to think about what then needs to be done in the anomaly calculation
function. Think have to do go back and forth a bit, therefore need a decent vignette.

- also look at biggest daily swings in temperature, and check if they're always
  at the start of the time series for a buoy. Implies turned on while on boat,
  bit like the albatrosses.

## Weekly anomalies - delete?

Travis's early code can do weekly, it just will not have all the extra
  functionality we have shown here (but we can modify  
