---
title: "Buoy Sea Surface Temperature Anomalies"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Buoy Sea Surface Temperature Anomalies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("buoys-from-vignette.Rmd")
or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 5
)
```

## Daily average sea surface temperature from buoys - do anomalies

Copying the buoys vignette to here to do some example code for anomalies, then
move that into a new vignette. Currently not really a vignette, more a train of thought.

```{r setup}
load_all()
# library(pacea)
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In pacea we include daily average sea surface temperature (SST) that we have calculated from
measurement from `r nrow(buoy_metadata)` buoys in
Canadian Pacific waters, yielding over 200,000 daily means.
Data are from Environment and Climate Change Canada, and Fisheries and
Oceans Canada. The earliest data are from September 1987, and 14 buoys were still providing data as of May 2023.

See the [buoys
  vignette](https://pbs-assess.github.io/pacea/vignettes/buoys.html) for
  background concerning the calculated daily mean sea surface temperatures and
  associated plotting functions in pacea. Also check out Andrea Hilborn et al.'s
  [Pacific_SST_Monitoring website](https://ios-osd-dpg.github.io/Pacific_SST_Monitoring_Site/)
  page for up-to-date graphics (pacea extractions were based on Andrea's code,
  but we do not have all her great figures as part of pacea). Adapt some of her
  code/ideas to do time series type plots. Here we are working towards a grid
  type plot that can be combined with other types of data (e.g. zooplankton anomalies).

Here we focus on constructing climatologies and subsequent anomalies, to plot in
a standardised way.

Let's choose Halibut Bank buoy C46146 in the Strait of Georgia as our example
one. It has a good timespan of data (though not back to 1990, which is a common
start year for a climatology).

```{r buoy}
one_stn_id_example <- "C46146"
two_stn_id_example <- c("C46146",
                        "C46185")
buoy_example <- buoy_sst %>%
  # filter(stn_id == one_stn_id_example)
  filter(stn_id %in% two_stn_id_example)
summary(buoy_example)
tail(buoy_example)
```

The default pacea plot is
```{r, plot}
plot(buoy_sst,
     stn_id = one_stn_id_example)
```

TODO Aside. We have not fully refined how to deal with some of the data
issues (e.g. the odd buoy had large fluctations in early days, but it's hard to
automate removing those). Acutally the anomaly code might help deal with
this. But some outliers might still need to be removed manually.



Next steps:

1. Create a climatology. Nice to have options for daily/weekly/monthly/annual?

2. Do anomalies. 

3. Generalise.

## Climatologies

Andrea does a daily climatology, so let's go with that. For the final plots I
have in mind, may then need to roll up into weekly/monthly/annual, or should you
really do those straight away?

The aim is to do ecosystem summaries for stock assessment and management
purposes. So likely want to be able to specify a time period of interest for
that particular stock, and then do a climatology/anomaly for that time period.

So do you do each day separately, then average them together? Or take a mean
over the dates and calculate a climatology and anomaly? Or are they equivalent?
Think I've thought about this before (see email of the math, but that ignore
sd).

Let's do climatology for each day.

Travis had `calc_clim()`, let's make `calculate_anomaly()` to do climatology and
anomaly, and then
generalise it. 

See Issue 96

Currently gives anomaly for every day, but makes more sense to use
`climatology_time`. And add a daily option.

This works, need to decide what we really want in the output.
```{r, anomaly}
xx <- calculate_anomaly(buoy_example,
                            climatology_time = "month")
xx

yy <- calculate_anomaly(buoy_example,
                            climatology_time = "week")
yy

```

Next need to adapt Andrea's ggplot code to do the anomaly plot; can do for
multiple stn_id's. 

Start with the example in the anomaly function. 

## Working out the plotting function

COPYING TO `plot.pacea_buoy_anomaly_list()` in  `plot-pacea-buoy-anomaly-list.R`
in this folder, to tidy up.
```{r newplot}
anomaly_plot <- 
xx$anomaly %>%
ggplot(aes(x = year, y = stn_id)) +
  geom_tile(aes(fill = sst_anomaly),
            colour = "black") +
  scale_fill_gradientn(colours = pals::ocean.balance(20)[3:18],
                       limits = c(-1.7, 1.7),
                       name = bquote("Annual SST anomaly ("*degree*C*")")) +
  scale_x_continuous(expand = c(0,0), name = NULL,
                     breaks = 1992:2026) +
  scale_y_discrete(expand = c(0,0), name = NULL) +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text = element_text(face = "bold",
                                  size = 12),
        panel.spacing.y = unit(0.1,
                               "lines")
  ) +
    geom_text(aes(label = round(sst_anomaly, 1)),
              size = 3.5) +
    guides(fill = guide_colorbar(barwidth = 15, title.position = "top",
                                 frame.colour = "black",
                                 ticks.colour = "black")) +
    labs(caption = "Ooh look at me")
  # Baseline period: 1991-2020\nDataset: CoralTemp 5km SST")

anomaly_plot

# ggsave(filename = "Annual_heatmap_SST_anomaly_SOPOregions_allyears.png", width = 9, height = 3., dpi = 300,
#       scale = 1.4)

```


HERE!



