---
title: "Buoy Sea Surface Temperature Anomalies"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Buoy Sea Surface Temperature Anomalies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("buoys-from-vignette.Rmd")
or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 5
)
```

## Daily average sea surface temperature from buoys - do anomalies

Copying the buoys vignette to here to do some example code for anomalies, then
move that into a new vignette. Currently not really a vignette, more a train of thought.

```{r setup}
load_all()
# library(pacea)
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In pacea we include daily average sea surface temperature (SST) that we have calculated from
measurement from `r nrow(buoy_metadata)` buoys in
Canadian Pacific waters, yielding over 200,000 daily means.
Data are from Environment and Climate Change Canada, and Fisheries and
Oceans Canada. The earliest data are from September 1987, and 14 buoys were still providing data as of May 2023.

See the [buoys
  vignette](https://pbs-assess.github.io/pacea/vignettes/buoys.html) for
  background concerning the calculated daily mean sea surface temperatures and
  associated plotting functions in pacea. Also check out Andrea Hilborn et al.'s
  [Pacific_SST_Monitoring website](https://ios-osd-dpg.github.io/Pacific_SST_Monitoring_Site/)
  page for up-to-date graphics (pacea extractions were based on Andrea's code,
  but we do not have all her great figures as part of pacea). Adapt some of her
  code/ideas to do time series type plots. Here we are working towards a grid
  type plot that can be combined with other types of data (e.g. zooplankton anomalies).

Here we focus on constructing climatologies and subsequent anomalies, to plot in
a standardised way.

Let's choose Halibut Bank buoy C46146 in the Strait of Georgia as our example
one. It has a good timespan of data (though not back to 1990, which is a common
start year for a climatology).

```{r buoy}
one_stn_id_example <- "C46146"
two_stn_id_example <- c("C46146",
                        "C46185")
buoy_example <- buoy_sst %>%
  # filter(stn_id == one_stn_id_example)
  filter(stn_id %in% two_stn_id_example)
summary(buoy_example)
tail(buoy_example)
```

The default pacea plot is
```{r, plot}
plot(buoy_sst,
     stn_id = one_stn_id_example)
```

### TODO Warning and caveats. 

We have not fully refined how to deal with some of the data
issues (e.g. the odd buoy had large fluctations in early days, but it's hard to
automate removing those). Acutally the anomaly code might help deal with
this. But some outliers might still need to be removed manually. And how many
daily values do we need a month to produce a climatology/anomaly? Easier to play
around with these once we have the plots. 

So don't treat the results as exact yet, this is a proof of concept.


Next steps:

1. Create a climatology. Nice to have options for daily/weekly/monthly/annual?

2. Do anomalies. 

3. Generalise.

## Climatologies

Andrea does a daily climatology, so had thought to go with that. For the final plots I
have in mind, may then need to roll up into weekly/monthly/annual, or should you
really do those straight away?

But, the aim is to do ecosystem summaries for stock assessment and management
purposes. So likely want to be able to specify a time period of interest for
that particular stock, and then do a climatology/anomaly for that time period.

So do you do each day separately, then average them together? Or take a mean
over the dates and calculate a climatology and anomaly? Or are they equivalent?
Think I've thought about this before (see email of the math, but that ignore
sd).

Let's do climatology for each day.

Travis had `calc_clim()`, let's make `calculate_anomaly()` to do climatology and
anomaly, and then
generalise it. 

See Issue 96

Currently gives anomaly for every day, but makes more sense to use
`climatology_time`. And add a daily option.

This works, need to decide what we really want in the output.
```{r, anomaly}
xx <- calculate_anomaly(buoy_example,
                            climatology_time = "month")
xx

yy <- calculate_anomaly(buoy_example,
                            climatology_time = "week")
yy

```



## Plotting function

Can calculate the anomalies for all buoys at once! And then tailor the plots.

The default plot is for the anomalies in April for all buoys:

```{r, buoyanomalyplot}
all_buoys_anomalies <- calculate_anomaly(buoy_sst)
plot(all_buoys_anomalies)
```

```{r, saveplotmaybe, echo = FALSE}
# ggsave(filename = "Annual_heatmap_SST_anomaly_SOPOregions_allyears.png", width = 9, height = 3., dpi = 300,
#       scale = 1.4)
```

How about the average anomaly for April to June for all buoys:
```{r, buoyanomalyplot2}
plot(all_buoys_anomalies,
     months = 4:6)
```

Other options, to add to the help and the testing:
```{r, buoyanomalyplot4}
plot(all_buoys_anomalies,
     sst_plot = "mean",
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     months = c(12, 1, 2))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184")

plot(all_buoys_anomalies,
     sst_plot = "count")  # check what this is doing, mean
# count over all years?? Ah no, it's april counts each year each buoy.

plot(all_buoys_anomalies,
     sst_plot = "count",
     count_breaks = c(0, 20, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184",
     count_breaks = c(0, 15, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8,
     stn_id = "C46184")
```


All buoys, but no 15-day requirement in the climatology or anomaly, instead use
all days:
```{r, usealldays
all_buoys_anomalies_all_days <- calculate_anomaly(buoy_sst,
                                                  min_days_per_month = 0)
plot(all_buoys_anomalies_all_days)

# expect_equal(all_buoys_anomalies, all_buoys_anomalies_all_days) worked when
# first arg used the original function before we added this in
```


Do single buoys also.

We can exclude the three buoys that do not have a long time series (this might
change when we refine the analyses):

don't need I think:

```{r, buoyanomalyplot3}

all_buoys_plot_3 <- plot(all_buoys_anomalies,
                         months = 4:6)
all_buoys_plot_3
```

Can do various options. 

Have averaged the anomalies for multiple months.


### Single buoy -- show each month

And if just one stn_id then plot months on the y-axis, January at the top:

```{r, singlebuoy}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206")
```
Hopefully self explanatory. 

Thinking of quick look at a relevant buoy for hake migration north with respect
to hake catches in Canada (which aren't in pacea but could be):
```{r, singlebuoy2}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206",
     months = 4:6)
```
If hake catches were added on then would be easy to spot any pattern.

Get the plots working when `climatology_time = "week"`.



### Refining the analyses

Important to be aware of:

Above all assumes our defaults in the calculations as specified in
`calculate_anomaly.pacea_buoy()`. These are somewhat subjective and can be
changed by the user. Specifically:

- the number of days needed within a calendar month to use to give a monthly
  average for calculating climatoloties and anomalies (default is 15 days)
- the maximum consecutive number of missing days allowed within a calendar
  month, above which that month is not used to give a monthly
  average for calculating climatoloties and anomalies (default is 6 days)
  
We pragmatically decided on requiring 15 days (roughly half) of daily SST within
a month, and a maximum consecutive number of missing days of 6. These are the
defaults, but can be changed by the user (as described below) to `NULL` to allow
all data and then examine it for plotting.

Similarly, if a user wants to look at average SST for a buoy over the months
Jan-Jun, a decision has to be made as to how many months should be
available. Could say a default of >75%, with no details about requiring
consecutive ones.

Given the huge number of options that users may wish to explore (many buoys,
many years, many possible combinations of months), users may wish to relax or
tighten the defaults, and have the power to do so.

Okay:

- add options to calculate function, with NULL being what we normally do, but
  defaults of 15 days and 6 days.


```{r, stop, cache = FALSE}
knitr::knit_exit()
```



### Rough notes, being adopted above somewhat


This works, but not for plotting yet, and may need to add options for how many
days needed in a week.

all_buoys_anomalies_weekly <- calculate_anomaly(buoy_sst, climatology_time = "week")

Need to remove data-poor months. Have some defaults that users can change, and
tell users to look at data. Elise thinks at least half the month, with no gap of
more than 5-6 days (no consecutive NA's longer than that).
Example we had was

```{r ex}
> yy <- table(all_buoys_anomalies$anomaly$sst_n)
> yy

   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14   15 
 777   32   19   29   26   24   29   21   24   26   26   26   36   19   34   34 

  16   17   18   19   20   21   22   23   24   25   26   27   28   29   30   31 
  57   68   72   59   76   69   94  113  149  171  346  301  727  682 1341 1798 

sum(as.data.frame(yy)$Freq)
[1] 7305
> sum(as.data.frame(yy)[21:32, ]$Freq)   # How many overall months have >=20
days of values.
[1] 5867
> 
```

The one showed up when we were playing around was:
filter(buoy_sst, stn_id == "C46183", year(date) == 2019, month(date) == 6) %>% a()
Has 13 missing days but they are kind of spread out, so looks usable. Hence go
with half a month needed, but no big gaps.

plot(all_buoys_anomalies, months = 1:12)  # North Hecate St 2019 has a big
anomaly, but it's only from that one month of data, as seen clearly in:
plot(all_buoys_anomalies, stn_id = "C46183")

Also need to think about climatologies. And whether to put the thinking part in
the climatology calculations and/or the plotting. Have to be in the climatology
calculations, as anomalies will change when we change the rules. And then all
resulting monthly climatologies/anomalies will be valid.

In the plotting we need a condition - how many months out of the requested ones
need to have anomaly calculations? Since monthly anomalies will have been calculated.

Could do the heatmap type plot on weekly or even daily values (for a single
buoy) to get an idea of the amount of missing data. Daily would be interesting.


Need to think about what then needs to be done in the anomaly calculation
function. Think have to do go back and forth a bit, therefore need a decent vignette.

- also look at biggest daily swings in temperature, and check if they're always
  at the start of the time series for a buoy. Implies turned on while on boat,
  bit like the albatrosses.
