---
title: "Create_BC_Partition"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Script for Creating the spatial Partition of BC

This markdown document creates the regular grid of BC's coastline. This regular
spatial grid defines how the data are stored internally within PACea. To obtain
covariates on other spatial resolutions, the function `ffff` is used to
aggregate/project the values from the grid onto the other regions (typically
with polygons of a larger area).

All other regions available for use in PACea are also defined here, with a
mapping matrix created for projecting the covariate values from the grid to the
regions.

First, we read in all the different regions.

```{r}
library(dplyr)

make_grid <- F

# Shapefile of the entire BC Coastline for plotting
Coastline <- sf::st_as_sf(PACea::Coastline)

Coastline_Simp <- sf::st_simplify(sf::st_buffer(Coastline, dist=20), dTolerance=25)

# Coarse polygons defining 4 Major regions: QCS, WCVI, HS, WCHG
BC_Major_Area_Boundaries <- sf::st_as_sf(
  PACea::BC_Major_Area_Boundaries)
BC_Major_Area_Boundaries <- sf::st_transform(
  BC_Major_Area_Boundaries,
  sf::st_crs(Coastline)
)
# Name the individual polygons
BC_Major_Area_Boundaries$Poly_Name <-
  c('HS','QCS','WCHG','WCVI')

# Coarse polygons defining the Pacific Herring Sections (from SpawnIndex package)
Pacific_Herring_Sections <- sf::st_transform(
  PACea::Pacific_Herring_Sections,
  sf::st_crs(Coastline))
# Convert Section to Statistical Area (first two digits)
Pacific_Herring_Sections$Statistical_Area <-
  factor(
  substr(Pacific_Herring_Sections$Section,1,2)
  )
# Merge by Statistical Area
Pacific_Herring_Sections <-
Pacific_Herring_Sections %>%
  dplyr::group_by(Statistical_Area) %>%
  dplyr::summarize()

# Name the individual polygons
Pacific_Herring_Sections$Poly_Name <-
  paste0('Herring_SA_',
         Pacific_Herring_Sections$Statistical_Area
)

which.max(diff(sf::st_coordinates(Coastline)[,2]))
sf::st_coordinates(Coastline)[1333:1337,]
sf::st_bbox(Coastline)
# expand the polygon by 50km Westwards
poly_expansion <-
  sf::st_polygon(list(matrix(c(390,1036.5509,
                      339,1036.5509,
                      1030,253.6858,
                      1080,253.6858,
                      390,1036.5509),
                      ncol=2,
                    byrow = T)))
poly_expansion <-
  sf::st_sf(sf::st_sfc(poly_expansion), crs = sf::st_crs(Coastline))

Coastline <- sf::st_union(poly_expansion, Coastline)
plot(Coastline)

if(make_grid)
{
# create large pixel grid over the simplified coastline
# Rotate the grid to maximize the spatial overlap
rotang = 318.5
rot = function(a) matrix(c(cos(a), sin(a), -sin(a), cos(a)), 2, 2)
tran = function(geo, ang, center) (geo - center) * rot(ang * pi / 180) + center

center <- sf::st_centroid(sf::st_union(Coastline))
# This is the function call which actually creates the master grid
# To increase the resolution (almost certainly desired), increase the
# n argument (e.g. n=c(100,100)).
Coastline_Grid <- sf::st_sf(
  sf::st_make_grid(tran(sf::st_geometry(Coastline), -rotang, center),
                   n=c(10,10)
                   )
)
Coastline_Grid <- tran(sf::st_geometry(Coastline_Grid), rotang, center)

# Keep only the polygons that fall within or touch the original Coastline to save storage space
Coastline_Grid <- sf::st_as_sf(sf::as_Spatial(Coastline_Grid))
sf::st_crs(Coastline_Grid) <- sf::st_crs(Coastline)

# Name the individual polygons by row-column method
# Note that sf works from bottom-left to top-right
# The first grid cell is the bottom-left grid cell (1,1)
# The final grid cell is the top-right grid cell (n,n)
# NEED TO CHANGE THE VALUES 10 TO MATCH THE CHOICE OF n=c(,) above!!
Coastline_Grid$Poly_Name <-
  paste0('Grid_Index_(',
         rep(1:10, each=10), # repeat each value of 1:cols, nrows
         ',',
         rep(1:10, times=10), # repeat sequence 1:nrows, ncols times
         ')'
)

Coastline_Grid  <-
  Coastline_Grid[Coastline,]

plot(Coastline_Grid)
plot(Coastline, add=T, col='red')

usethis::use_data(Coastline_Grid, overwrite = T)
}
if(!make_grid)
{
  Coastline_Grid <- PACea::Coastline_Grid
}


```

Next, we bind the competing region definitions into a single spatial object
called BC_Partition. We name the BC grid, which will eventually store the data,
as 'BC Grid' within the variable `Regions_Name`. We then define a matrix to
perform the aggregations from the BC Grid to the other regions' polygons

```{r}
# If additional regions are added, simply add them to the bind_rows
# function call. Note that BC Grid MUST GO FIRST
All_regions <-
  dplyr::bind_rows(
        Coastline_Grid %>% dplyr::mutate(Regions_Name = 'BC Grid'),
        BC_Major_Area_Boundaries %>% dplyr::mutate(Regions_Name = 'BC Major Areas'),
        Pacific_Herring_Sections %>% dplyr::mutate(Regions_Name = 'Pacific Herring Spawn Statistical Areas'))


#BC_Partition <-
#  sf::st_intersection(All_regions)

# Convert the GEOMETRYCOLLECTION to multipolygon
#BC_Partition <-  sf::st_collection_extract(BC_Partition, type = c("POLYGON"))

# Create a Poly_ID variable
BC_Partition <-
  All_regions %>%
  mutate(Poly_ID = row_number()) %>%
  select(Poly_Name, Regions_Name, Poly_ID)

# Compute the areas of each region
region_Areas <-
sf::st_area(BC_Partition)

# Compute the area of intersection between the (approximate) ocean polygon and each of the regions
Ocean_Intersection_Areas <- rep(0,dim(BC_Partition)[1])
for(i in 1:dim(BC_Partition)[1])
{
  Ocean_Intersection_Areas[i] <-
    ifelse(is.null(sf::st_area(sf::st_intersection(BC_Partition[i,], Coastline, drop=F))),
           0,
           sf::st_area(sf::st_intersection(BC_Partition[i,], Coastline, drop=F)))

}

# Check that all the regions are POLYGON or MULTIPOLYGON type
sf::st_geometry_type(BC_Partition)

plot(BC_Partition)

# Define the mapping between each region with the BC Grid
  # Note - the matrix will also be a useful concept in future, when
  # users wish to define new regions which are union of existing regions
Mapping_Matrix <- #diag(nrow=dim(BC_Partition)[1])
   matrix(0,
          nrow = dim(Coastline_Grid)[1],#dim(BC_Partition)[1],
          ncol = dim(BC_Partition)[1])

for(i in 1:dim(BC_Partition)[1])
 {
  # loop through the partitioned regions and compute the area of
  # intersection with Coastline Grid.
  Mapping_Matrix[sf::st_intersects(Coastline_Grid, BC_Partition[i,],
                                   sparse = F),
                 i] <-
    as.numeric(sf::st_area(sf::st_intersection(Coastline_Grid, BC_Partition[i,])))
  # Divide by rowSums to normalize to 1 (to create spatially-weighted mean)
  if(sum(Mapping_Matrix[,i])>0)
  {
    Mapping_Matrix[,i] <- Mapping_Matrix[,i] / sum(Mapping_Matrix[,i])
  }
  if(sum(Mapping_Matrix[,i])==0)
  {
    Mapping_Matrix[,i] <- 0
  }
  }

# Check the mapping has been made correctly
Mapping_Matrix
colSums(Mapping_Matrix)

# create a named list with each entry equal to a index vector
# each index vector points the set of regions to the correct rows
# Mapping Matrix  (e.g. WCHG, WCVI, HS, QCS)
index_vectors <-
  list(Coastline = which(BC_Partition$Regions_Name=="BC Grid"),
       BC_Major_Area_Boundaries = which(BC_Partition$Regions_Name=="BC Major Areas"),
       Pacific_Herring_Sections = which(BC_Partition$Regions_Name=="Pacific Herring Spawn Statistical Areas"))


# save the results for future use
BC_Partition_Objects <-
  list(BC_Partition=BC_Partition,
       region_Areas=region_Areas,
       Ocean_Intersection_Areas=Ocean_Intersection_Areas,
       Mapping_Matrix=Mapping_Matrix,
       index_vectors=index_vectors)

usethis::use_data(BC_Partition_Objects, overwrite = T)

```

The basic idea of the mapping matrix is as follows. The number of rows of the
matrix equals the number of cells of the BC grid (in which the data is
stored). The number of columns of the matrix equals the number of polygons
across all the regions stored in PACea (e.g. Herring Sections). Then, a given
column of the matrix has entries equal to the fraction of its total area which
is contained within each grid cell. These fractions sum to 1. Thus, to evaluate
the average value of a covariate (e.g. SST) in a specific region, simply sum the
average values within each grid cell, multiplied by the fraction of the total
area of intersection. Because these sum to 1, the quantity is a
spatially-weighted average. Of course, this only works if the data are not
missing in any grid cell, else the fractions need to be
re-weighted. Furthermore, if a sum or total of a quantity over a region is
desired instead of a mean concentration or density, then we simply scale by the
area of the region - stored as `Ocean_Intersection_Areas`, assuming the quantity
can only exist in the water.
