---
title: "BCCM Results Over Full Domain"
author: "Andrew Edwards and Travis Tai"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BCCM Results Over Full Domain}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r run, echo = FALSE, eval = FALSE}
rmarkdown::render("bccm_full.Rmd")
# to build, or click the knit button in RStudio
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  cache = TRUE,
  cache_path = "bccm-full-cache/",
  fig.path = "bccm-full-figs-cache/",
  fig.width = 7.7,
  fig.height = 7
)
```

# Introduction

The British Columbia Continental Margin (BCCM) model is a
physical biogeochemical oceanographic model, implemented using the Regional Ocean Modeling
System (ROMS; Peña et al., 2019). Results restricted to Canada's Exclusive Economic Zone and mapped
to a 2 km x 2 km grid onshore and 6 x 6 km grid offshore are
discussed in the [BCCM Results vignette](bccm.html), which should be looked at
to understand more about the model. The restrictions and
varying of the gridsize were done to reduce file sizes.

Due to several requests from users we have now also mapped the BCCM results for its full domain (extending
north and south into US waters) onto a 2 km x 2km grid. These requests included:

- [Issue #49](https://github.com/pbs-assess/pacea/issues/49): extend the BCCM
  results into US waters
- [Issue #58](https://github.com/pbs-assess/pacea/issues/58): add the results from the
HOTSSea physical oceanographic model for the Salish Sea (a region not covered by
the BCCM results, ideally on a grid that overlaps as that used for the BCCM
- [Issue #48](https://github.com/pbs-assess/pacea/issues/48) and [Issue
  #62](https://github.com/pbs-assess/pacea/issues/62): have an easily exported grid that covers a larger domain
- [Issue #60](https://github.com/pbs-assess/pacea/issues/60): projecting the Optimally Interpolated Sea Surface Temperature values onto the same grid as the
BCCM results.

These requests are now mostly fulfilled, and include the BCCM results on the
full domain. The results are in the same format as those for the restricted
domain described in the [BCCM Results vignette](bccm.html) vignette. The
file sizes are much larger (each variable is about 120 Mb rather than about 30
Mb), but only need to be downloaded once.

We have kept consistent terminology and functionality as for the restricted BCCM
results. Basically:
- variable names are appended with `_full`
- plotting and analysis functions should still work on the `_full` variables.

Deciding if you need the `_full` outputs or just the original over Canada's EEZ:

- use the original if you want OISST data on its original grid (unrelated to
BCCM grid) over the EEZ, and/or BCCM over EEZ, and don't want to do
spatiotemporal modelling involve both of those or also involving HOTSSea output
(it is okay for you if the BCCM and HOTSSea grids do not overlap) -- use existing variables
- if you want anything else, like OISST or BCCM extended to a larger domain,
  HOTSSea and/or OISST and/or BCCM on the same grid (especially for
  spatiotemporal analyses) - use the new BCCM `_full` variables described here.

## Available variables

```{r packages}
library(pacea)
library(dplyr)
library(sf)
library(ggplot2)
```

For the restricted domain, the available variables are given in the data object
`bccm_data`. For the full domain, the variables are therefore given in
`bccm_data_full`:
```{r bccm_data-Full}
bccm_data_full
```
The `bccm_data` object gives the same `r nrow(bccm_data_full)` names, but without
the `_full`.

## Downloading bccm full results

The 22 BCCM full model results are stored on Zenodo at https://zenodo.org/records/14031460. Each can be downloaded
individually (if you only want certain variables) as described below. It is
easiest to just
just simply download them all to your computer in one go, using
```{r download, eval = FALSE}
bccm_all_variables_full()
```
This places them into a cache directory on your local machine, given by
by `paste0(pacea_cache(), "/bccm_full")`. It only took about 10 minutes to
download them all (the code uses parallel cores on Windows) on a home network,
but could be longer on a work network. You can check progress in your local
cache directory, where files should gradually appear.
Any files that are already
present in your cache directory are not re-downloaded. See
`?bccm_all_variables_full` if you have problems.

### Using the bccm full results

As an example, let's look at the estimates of primary production, calling the
object `pp` locally:
```{r pp}
pp <- bccm_primaryproduction_full()
```
Note the `()`. The `bccm_primaryproduction_full()` function assigns the
output to our designated variable simply named `pp`.

If you have not already previously downloaded the relevant results it
will first download them into your cache folder (i.e. it will get just the
specified model results, rather than all of them as in the
`bccm_all_variables_full()` described above). If you have previously downloaded
the results the function will simply load them in from your cached folder.

View the data help for more information
(e.g. `?bccm_primaryproduction_full`); similar help functions all point to a
common help file for all `bccm_full` objects.

So, what does the object look like?
```{r pp1}
pp
head(pp[, 1:5])   # note that the geometry column is included when selecting columns from
                  # an `sf` object.
```

The object is in wide format, with each column representing a unique year-month
combination, and is a 'simple features' (`sf`) R object, and so contains a
`geometry` column. The data also have various attributes, such as the units for the
data values, and some extra ones to automate some plotting. (This is useful but
not essential for users to know).
```{r}
names(attributes(pp))
attributes(pp)$units
class(pp)
```
## Visualising the bccm full results

We give some examples here, and then refer users to the [BCCM Results
vignette](bccm.html) since the examples shown there should work for the `_full`
results. (Note we have not tested every potential idea you might have, but by
keeping the structure of the outputs the same, our custom plotting and analysis
functions should work).


### Plotting with `plot()`

Results can be plotted using the `plot()` function (that we have
customised). The default settings are to plot results for April, 2018. However,
users can specify any time period(s) available in the results. Examples are:
```{r plot1}
plot(pp)
```

To see the estimated primary production for each month of 2018
```{r, ppplot, fig.width = 16, fig.height = 12}
plot(pp, months = 1:12)
```

This shows the production ramping up through the spring.

### Climatologies and anomalies

Let's have a quick look at anomalies, say for oxygen at the bottom of the water
column, adapting the examples from the [BCCM
vignette](http://htmlpreview.github.io/?https://github.com/pbs-assess/pacea/blob/main/vignettes/bccm.html). Again,
see that vignette for extra details.

So let's ask: how do the anomalies of bottom oxygen for
April and September in 2010 and 2018, look compared to a climatology (for each
month) from 1993 to 2010 (the full range of years currently available).

We can easily load the model results and calculate the anomalies for a
climatology (see `?calc_clim` to obtain the climatology also):
```{r, anomalies}
bottom_oxygen <- bccm_bottom_oxygen_full()
anom <- calc_anom(bottom_oxygen,
                  clim_years = 1993:2010,
                  time_period_return = c("Apr", "Sep"),
                  years_return = c(2010, 2018))
anom
```
```{r, anomaliesplot, fig.width = 8, fig.height = 8}
plot(anom,
     months.plot = c("Apr", "Sep"),
     years.plot = c(2010, 2018),
     eez = FALSE)
```

So some areas are relatively oxygen rich (green) in these months compared to the
same areas from 1993 to 2010, whereas others are lower in oxygen (red). This
varies spatially and by month. Along the coast there looks to be more oxygen
than normal in April 2010, generally a bit less in September 2010, and even less
in April and September 2018.

See the [BCCM
vignette](http://htmlpreview.github.io/?https://github.com/pbs-assess/pacea/blob/main/vignettes/bccm.html)
for further example calculations, which hopefully stimulate your own ideas.

## Reference

Peña, M.A., Fine, I. and Callendar, W. 2019. Interannual variability in primary
production and shelf-offshore transport of nutrients along the northeast Pacific
Ocean margin. Deep-Sea Research II, doi:10.1016/j.dsr2.2019.104637.
