---
title: "Buoy Sea Surface Temperature Data"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Buoy Sea Surface Temperature Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 5
)
```

## Daily average sea surface temperature from buoys

```{r setup}
library(pacea)
library(dplyr)
library(tibble)  # Else prints all of a tibble
```

In pacea we include daily average sea surface temperature (SST) from 19 buoys in
Canadian Pacific waters, yielding over 170,000
values. Data are from Environment and Climate Change Canada, and Fisheries and
Oceans Canada. The earliest data are from September 1987, and 14 buoys were still providing data as of May 2023.

Metadata for the buoys is given by
```{r buoymeta}
buoy_metadata
```
which includes each buoy's World Meteorological Organisation's weather station id
(`wmo_id`), its 'common name' (`name`), location, and depth of the water in which the buoy resides. See
`?buoy_metadata` for descriptions of all columns. The full names of all buoys
are
```{r names}
buoy_metadata$name
```

The buoys are located at

TODO plot of buoy locations, once have written it.

Some of the data are shown at
https://github.com/IOS-OSD-DPG/Pacific_SST_Monitoring/ by Andrea Hilborn,
Charles Hannah and Lu Guan, which is automatically updated roughly every
week (so look there for a quick glance at recent conditions). For pacea we have,
with Andrea, adapted some of that code to include the data and create plotting
functions in our package.

The wrangling of data is taken care of within pacea, and includes using
protocols to remove certain flagged data, dealing with timezones and pesky
daylight savings time changes, and averaging over a day (the original raw data are even
higher resolution).

The sst value are saved in a tibble:   TODO prob give it it's own class also
```{r buoysst}
buoy_sst
tail(buoy_sst)
```
with `stn_id` specifying each buoy as described above in `buoy_metadata`, and
`date` based on UTC -8 hours (i.e. Pacific Standard Time, not changing due to
daylight savings), and `sst` is the mean SST (deg C) for that day at that
station; see `?buoy_sst` for full details.


To see the ranges of dates for each buoy, use `dplyr` functions in the usual
way:
```{r buoydates}
buoy_ranges <- buoy_sst %>%
  group_by(stn_id) %>%
  summarise(start = min(date),
            end = max(date))
buoy_ranges
```

TODO plots

Plot data from a single buoy for all years, the default is for buoy C46146:
```{r plot}
# Play with one to get the hang of ggplot2, adding in Andrea's options maybe.
# also see https://www.r-bloggers.com/2020/04/lubridate-ggplot-date-helpers/
#  for making extra columns early on.
plot(buoy_sst)
```

Crude way of doing each one in turn (easier to see details on big plots, but
will adapt Andrea's multi-panel plots also TODO):
```{r multiloop}
for(i in 1:nrow(buoy_metadata)){
  plot(buoy_sst,
       stn_id = buoy_metadata[i, ]$stn_id) %>% print()
}
```

HERE

Might needs some of this for multipanel plot
```{r eval=FALSE}
# Andrea's:
ggplot(rbind(dat1,dat2), aes(month(date, label=TRUE, abbr=TRUE),
                value, group=factor(year(date)), colour=factor(year(date)))) +



p <- buoy_sst %>%      # then filter as necessary given function options. Try
                       # and just do all for now.
  arrange(stn_id) %>%      # TODO maybe not needed, or do up front in data-raw
  ggplot() +
  facet_wrap(~stn_id, ncol = 3) +
  geom_ribbon(data = buoy_sst,
              aes(x = date,
                  y = sst,
                  ymin = 0,
                  ymax = 20                  ))
                  )) # ymin = SSTP_10perc_day, ymax = SSTP_90perc_day), fill = "grey70", colour = NA, alpha = 0.5) +
  ## geom_path(data = sstclim, aes(x = fakedate, y = SSTP_clim_mean_day), colour = "grey30", linewidth = 0.8) +
  ## # Heat dome - for 2021-2022 plots
  ## # geom_vline(xintercept = as.Date("2021-06-26"), linetype = "dashed") +
  ## geom_path(data = prevyr %>% filter((year(date) == plot_yrs[1]) | is.na(SSTP)), aes(x = fakedate, y = SSTP, colour = col_key), size = 0.6) +

  ## geom_path(data = currentyr %>% filter((year(date) == plot_yrs[2]) | is.na(SSTP)),aes(x = fakedate, y = SSTP), colour = "black", size = 1.4, lineend = "round") +
  ## geom_path(data = currentyr %>% filter(year(date) == plot_yrs[2] | is.na(SSTP)), aes(x = fakedate, y = SSTP), colour = "white", size = 0.5, lineend = "round") +
  ## scale_colour_manual(values = (glasbey_mod[1:(nrow(buoyplot))])) +
  ## scale_colour_identity(breaks = buoyplot$col_key) +
  ## xlab(NULL) +
  ## scale_x_date(labels = scales::date_format("%b"), breaks = "2 months") +
  ## theme(legend.position = "none",
  ##       strip.background = element_rect(colour = "grey70", fill = "grey95"))

# Time series plot ####
s <- prevyr %>%
  arrange(STN_ID, fakedate) %>%          # STN_ID %in% c("C46206","C46146")) %>%
  ggplot() +
  facet_wrap(~name_key, ncol = 3) +
  geom_ribbon(data = sstclim, aes(x = fakedate, ymin = SSTP_10perc_day, ymax = SSTP_90perc_day), fill = "grey70", colour = NA, alpha = 0.5) +
  geom_path(data = sstclim, aes(x = fakedate, y = SSTP_clim_mean_day), colour = "grey30", linewidth = 0.8) +
  # Heat dome - for 2021-2022 plots
  # geom_vline(xintercept = as.Date("2021-06-26"), linetype = "dashed") +
  geom_path(data = prevyr %>% filter((year(date) == plot_yrs[1]) | is.na(SSTP)), aes(x = fakedate, y = SSTP, colour = col_key), size = 0.6) +

  geom_path(data = currentyr %>% filter((year(date) == plot_yrs[2]) | is.na(SSTP)),aes(x = fakedate, y = SSTP), colour = "black", size = 1.4, lineend = "round") +
  geom_path(data = currentyr %>% filter(year(date) == plot_yrs[2] | is.na(SSTP)), aes(x = fakedate, y = SSTP), colour = "white", size = 0.5, lineend = "round") +
  scale_colour_manual(values = (glasbey_mod[1:(nrow(buoyplot))])) +
  scale_colour_identity(breaks = buoyplot$col_key) +
  ylab(expression("Mean Daily SST " ( degree*C))) +
  xlab(NULL) +
  scale_x_date(labels = scales::date_format("%b"), breaks = "2 months") +
  theme(legend.position = "none",
        strip.background = element_rect(colour = "grey70", fill = "grey95"))
```
