---
title: "Ecosystem Summary Heatmaps"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ecosystem Summary Heatmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("buoy_anomalies.Rmd")
or click the knit button in RStudio -->


```{r, start, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 14,   #5.7,
  fig.height = 7  #5
)
```

## Customisable ecosystem summary heatmaps

```{r setup}
load_all()    # for developer
# library(pacea)  # for user
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In the
[buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we showed how to make customisable heatmaps of the buoy sea-surface
temperature data that are saved in pacea. They are in a format famililar to
oceanographers, using prescribed climatologies and subsequent anomalies.

Here we show how to produce customisable heatmaps that include disparate types
of ecosystem information, thus bringing together the various data objects in
pacea. Importantly the user can choose:

- the types of ecosystem information to be included, tailored to their
  particular stock
- the area over which spatial information is summarised

The aim of pacea is to
facilitate inclusion of ecosystem information in advice to fisheries management,
such as stock assessments. Given the multitude of stocks in DFO's Pacific
Region, both the diversity of species and the many stock-specific spatial areas, we decided
to create functions for user-customisable plots that easily summarise the relevant ecosystem
information.

Expert knowledge can be used to suggest relevant ecosystem information to be
used for a particular stock. Futhermore, the ease of generating the heatmaps can
allow exploration, formulation of ideas, and quick hypothesis testing.

For comparing disparate types of ecosystem information, we standardise each time
series (by subtracting the mean and dividing by the standard deviation across
the time series, which might not be all fully plotted). 
This is so that indices
are all dimensionless and can be compared on equivalent scales, and shown on a
single heatmap style plot.

As an example, let's use various ecosystem information for the Strait of Georgia
(SOG). We create an index for each type of ecosystem information in turn and
then make the plot. We just need to use `create_index()` multiple times to
create the indices, and then `plot_indices()` to make the function (we will
explain them in more detail later):
```{r, examples}
# Average sea surface temperature in April from a buoy in the SOG
sst_buoy_halibut_bank_apr <- create_index(buoy_sst,
                                          months = 4,
                                          stn_id = "Halibut Bank")

# Large calanoid copepods in the SOG 
calanoid_copepods_large_sog_index <- create_index(zooplankton_sog,
                                                  zooplankton_species_group =
                                              "calanoid_copepods_large")

# Total zooplankton in the SOG
total_zooplankton_sog_index <- create_index(zooplankton_sog,
                                            zooplankton_species_group = "total_biomass")

# Annual age-2 herring recruitment in the SOG
herring_recruitment_sog_index <- create_index(herring_recruitment,
                                              herring_region = "SOG")

# Annual herring spawning biomass in the SOG
herring_biomass_sog_index <- create_index(herring_spawning_biomass,
                                          herring_region = "SOG")

# Annual population estimates of Harbour Seals in the SOG
harbour_seals_sog_index <- create_index(harbour_seals,
                                        seal_region = "SOG")
```

Now make the plot, manually basing the start year on the shortest
of those (zooplankton data):
```{r, sogplot, fig.height = 4}
plot_indices(sst_buoy_halibut_bank_apr,
             calanoid_copepods_large_sog_index,
             total_zooplankton_sog_index,
             herring_recruitment_sog_index,
             herring_biomass_sog_index,             
             harbour_seals_sog_index,
             years = 1996:2026,
             gap_after = c("sst_buoy_halibut_bank_apr",
                           "total_zooplankton_sog_index",
                           "herring_biomass_sog_index"))
```

The first six arguments are the indices that we created above, and `years`
restricts the plot to the years specified.
To delineate different types of ecosystem information in the plot
(e.g. here we have temperature, zooplankton, fish, mammals), we used the optional
`gap_after` argument to add a small gap after the specified indices.

The restriction of the years being shown is why the seal index is always
positive. By definition the mean of the index 
should be 0, and so it is negative earlier in the time series. A quick check of
the absolute seal populations numbers confirms the population has indeed
increased:

```{r, sealplot, fig.width = 5.7, fig.height = 5}
plot(harbour_seals,
     region = "SOG")
```


TODO Add an option in the plot function to restandardise across the years specified?

(see the
relevant vignettes and help files for details of the pacea objects used here as
the first argument). The first is from the buoy sea surface temperature, looking at the [buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we will use the Halibut Bank buoy in the SoG, which has good timespans of
data, and just pick April. We then create indices for other types of ecosystem
information (we will do into details 

This is done in the
`create_index()` function, to create an object of class
`pacea_standardised_index`, no matter what type the ecosystem information is. 


Buoy SST getting working, need to prove if normalised index of mean anomalies is same
as normalised index of mean of means. Think so since both relative to the same
climatology. TODO. Also it's still based on climatology; does mean of actual
values just give the same?

```{r, buoysst}
sst_west_dixon_apr_using_anomalies <- create_index(buoy_sst)
sst_west_dixon_apr_using_mean <- create_index(buoy_sst, buoy_statistic = "mean")
expect_equal(sst_west_dixon_apr_using_anomalies,
             sst_west_dixon_apr_using_mean)
# Not convinced that will always hold though, I think only difference would be
# if you did not have a climatology for a particular month, which does not
# happen with the default years. So probably best to keep the generality. TODO
```

Some examples of different choices of months, and how `index` and `index_label`
get automatically prescribed:
```{r, createindmonths}
# Different examples:
create_index(buoy_sst,
             months = 5)
create_index(buoy_sst,
             months = 5:6)
create_index(buoy_sst,
             months = c(11, 12, 1, 2, 3))
```

Note that in the first example we have a value for 1992, but not in the second
one. This must be because June 1992 does not have enough daily values for this
particular buoy, which we can confirm from the raw data:
```{r, confirm}
filter(buoy_sst,
       stn_id == "C46205",   # default in create_index.pacea_buoy()
       date >= "1992-06-01",
       date <= "1992-06-30",
       !is.na(sst))
```
So there are only 6 daily SST values for this buoy in June 1992, and so an
anomaly is not calculated for that month. We could force an anomaly to be
calculated by changing the arguments `min_days_per_month` (default 15) and 
`max_consecutive_missing_days` (default 6) that get passed onto
`calculate_anomalies.pacea_buoy()`:
```{r, confirmchange}
create_index(buoy_sst,
             months = 5:6,
             min_days_per_month = 6,
             max_consecutive_missing_days = 24)
```
thus giving an index value for 1992, given these definitions (which we do not
recommend, but are showing as examples to help understanding, and to show that
arguments get correctly passed onto further functions).


Further examples (originally had above while generating the code), also check
the help file of `plot_indices()` and add to that TODO
```{r, furthrexamples}
herring_recruitment_hg_index <- create_index(herring_recruitment,
                                             herring_region= "HG")
herring_recruitment_prd_index <- create_index(herring_recruitment,
                                              herring_region = "PRD")
herring_recruitment_cc_index <- create_index(herring_recruitment,
                                             herring_region = "CC")
herring_recruitment_wcvi_index <- create_index(herring_recruitment,
                                              herring_region = "WCVI")
herring_biomass_wcvi_index <- create_index(herring_spawning_biomass,
                                           herring_region = "WCVI")

hake_recruitment_index <- create_index(hake_recruitment)

hake_biomass_index <- create_index(hake_biomass)

harbour_seals_wcvi_index <- create_index(harbour_seals)  

# Many examples (adding as go along), likely not really needed:
plot_indices(herring_recruitment_hg_index,
             herring_recruitment_prd_index,
             herring_recruitment_cc_index,
             herring_recruitment_sog_index,
             herring_recruitment_wcvi_index,
             herring_biomass_wcvi_index,             
             hake_recruitment_index,
             hake_biomass_index,
             harbour_seals_wcvi_index,
             harbour_seals_sog_index,
             total_zooplankton_sog_index)
```

See the test for more examples. Summarise here maybe. 


TODO should the years argument then require indices to be recalculated across
just those years? Question for the audience maybe..., show the consequences of
that. Historically good/bad versus short-term good/bad. Will need both options,
decide on default.


## OISST data

Just doing examples here for now (might plot strange with so many plots), may
move to new vignette. You might want to do the anomalies first to look at them
spatially, and can then calculate an index directly from the anomalies list, or
directly from `create_index(oisst_month)` with options, which will then call `calculate_anomalies()` and then
`create_index.pacea_oisst_anomalies_list()` , but only return the index.

```{r, oisst1}
oisst_anomalies <- calculate_anomalies(oisst_month)

plot(oisst_anomalies)

oisst_anomalies_area_126 <- calculate_anomalies(oisst_month,
                                                area = "pfma_area_126")  # hardwired in at the moment TODO

plot(oisst_anomalies_area_126)

# Examples with the options:
plot(oisst_anomalies_area_126, months = 3)

plot(oisst_anomalies, months = 3)


# Shows MHW coming on. Then suggests an index to create, e.g. average of the
# whole area for Jan to Jun (TODO)
plot(oisst_anomalies,
     months.plot = 1:12,
     years.plot = 2014:2016)


plot(oisst_anomalies_area_126,
     year = c(2015, 2016, 2026),
     month = 3:6)

```

TODO need options for total area to show

TODO HERE go through these, checking can do the anomalies and create an index,
and get the same index if just creating them directly from `oisst_month` with the
area option.

Having calcualted the full anomalies, can create a variety of indices.

Now do the index, playing around for now:
```{r, oisstindex}
oisst_anomalies_area_126
xx <- create_index(oisst_anomalies_area_126)
# Taking area 126 then making a slightly smaller polygon
## area_as_list_of_matrix <-
##         list(matrix(c(-127.1506, -128.2331, -129.3492, -127.9167, -127.1847,
##                       -126.8200, -127.1506,
##                       49.85766, 49.00000, 48.99991, 50.11915, 50.40183,
##                       50.24466, 49.85766),
##                     ncol = 2))    # From Travis's OISST vignette
area_smaller_than_126_as_list <-
  list(matrix(c(-128, -129.3492, -127.9167, -127.1847,
                      -126.8200, -128,
                      49.85766, 48.99991, 50.11915, 50.40183,
                      50.24466, 49.85766),
               ncol = 2))

area_smaller_than_126_polygon <-
  sf::st_sfc(sf::st_polygon(area_smaller_than_126_as_list),
             crs = 4326) %>%
  sf::st_as_sf()

# Check we have less values:
oisst_anomalies_area_smaller_than_126 <-
  oisst_anomalies_area_126$anomalies[area_smaller_than_126_polygon, ]

# This has just 5 entries, area_126 has 20 (shown next). So fine for a test run.
oisst_anomalies_area_smaller_than_126 %>% filter(year == 1981, month == 9)

oisst_anomalies_area_126$anomalies %>% filter(year == 1981, month == 9)

# TODO try area option in create_index, see below

# Get it working for area 126, then for the defined subset of area26

oisst_area_126_apr <-
  create_index(oisst_anomalies_area_126)
oisst_area_126_apr

# Average sea surface temperature in April from a buoy within area 126
sst_buoy_south_brooks_apr <- create_index(buoy_sst,
                                          months = 4,
                                          stn_id = "South Brooks")
plot_indices(oisst_area_126_apr,
             sst_buoy_south_brooks_apr)

# Then restrict to 1995 onwards as that's when the buoy starts
oisst_area_126_apr_from_1995 <-
  create_index(oisst_anomalies_area_126,
               years = 1995:2026)

plot_indices(oisst_area_126_apr_from_1995,
             sst_buoy_south_brooks_apr)

# Try smaller area option
oisst_area_smaller_than_126_apr_from_1995 <-
  create_index(oisst_anomalies_area_126,
               area = area_smaller_than_126_polygon,
               years = 1995:2026,
               index_name = "smaller_area",
               index_label = "Smaller area than 126")

oisst_area_smaller_than_126_apr_from_1995
plot_indices(oisst_area_126_apr_from_1995,
             oisst_area_smaller_than_126_apr_from_1995,
             sst_buoy_south_brooks_apr)

```

You can create an index directly from the `oisst_month` data, for which the
anomalies are calculated internally. Though it might be a good idea to calculate
full spatial anomalies for your area to take a look at them first, as we did
above. But `create_index()` will work directly on `oisst_month`, with various options.
```{r, oisstindex}
oisst_area_smaller_than_126_apr_from_1995_directly <- create_index(oisst_month,
                                                                   area = area_smaller_than_126_polygon,
                                                                   years = 1995:2026,
                                                                   index_name = "small_direct",
                                                                   index_label =
                                                                     "Small area directly calculated")

oisst_area_smaller_than_126_apr_from_1995_directly   # gives same answer as earlier

expect_equal(oisst_area_smaller_than_126_apr_from_1995_directly$value,
             oisst_area_smaller_than_126_apr_from_1995$value)

# As expected, these are the same:
plot_indices(oisst_area_smaller_than_126_apr_from_1995_directly,
             oisst_area_smaller_than_126_apr_from_1995)

# Keeping plot.pacea_oisst_anomalies_list.Rd as an option, so explain that.


```



```{r, stop123, cache = FALSE}
# Do example index with:
# return_results = TRUE)

knitr::knit_exit()
```



In pacea we include daily average sea surface temperature (SST) that we have calculated from
two-minute measurements from `r nrow(buoy_metadata)` buoys in
Canadian Pacific waters, yielding over 200,000 daily means.
Data are from Environment and Climate Change Canada, and Fisheries and
Oceans Canada. The earliest data are from September 1987, and 18 buoys were
still providing data as of August 2026.

See the
[buoys_vignette](https://pbs-assess.github.io/pacea/vignettes/buoys.html) and
the help file `?buoy_sst` 
for background concerning the calculated daily mean sea surface temperatures
and associated simple plotting functions in pacea. Here we focus on
constructing climatologies and subsequent anomalies, to plot in 
a standardised way.

The locations of the buoys are:
```{r buoysmap, echo=FALSE, fig.width = 5.7, fig.height = 5, out.width = "90%"}
knitr::include_graphics(paste0(here::here(),
                               "/man/figures/buoys_map.png"))
```

If you want to quickly get a sense of current temperatures (in the wider
  northeast Pacific also), check out Andrea Hilborn et al.'s
  [Pacific_SST_Monitoring website](https://ios-osd-dpg.github.io/Pacific_SST_Monitoring_Site/)
  page. The pacea package is more suited for doing more speficic plots and
  analyses (because the data are easily accessible).

## Calculations and plotting

### Default plot

We can easily calculate a climatology (by month) and subsequent anomalies for
each buoy, and plot results:
```{r, buoyanomalies}
all_buoys_anomalies <- calculate_anomalies(buoy_sst)

plot(all_buoys_anomalies)
```

The plot's title describes that this plot is for the anomalies **in April** (the
default) for all
buoys, and the climatology was 1991-2020 where
available. Buoys are arranged in a roughly north-to-south order (see map). So each square
gives the SST anomaly for that buoy for April of that year. Empty (white)
squares indicate not enough data available for that particular calculation.

The figure clearly shows that April 2015 and April 2016 were warmer than usual
for April, for every buoy (with those years having the warmest two Aprils for
almost every buoy). The bottom three buoys clearly have limited data. 

### Absolute SST's

To show the actual monthly average SST's (rather than anomalies), again for the
default month of April:
```{r, buoyanomalies3}
plot(all_buoys_anomalies,
     sst_plot = "mean")
```
Here you can roughly see the water getting warmer from north to south, noting
that the three "NOMAD" buoys are much further offshore than the others, and
look slightly cooler.

### Full results for a specific buoy

If you specify a buoy, you get the monthly anomalies across the years (again
with an automated title), with months running from top to bottom:
```{r, buoyanomalies2}
plot(all_buoys_anomalies,
     stn_id = "Halibut Bank")   # In the Strait of Georgia
```

You can check that the April 2016 high anomaly matches that shown in the first
anomaly plot shown above.

### Anomalies for all buoys for different months

You can specify a single month (the default being "4" for April):
```{r, buoyanomaliesmonth}
plot(all_buoys_anomalies,
     months = 12)
```

Interestingly, in 2023 the December anomalies are all positive, whereas in the
first anomaly plot above the April anomalies are all negative (for the available
buoys in each case). Discuss. (We may use this example in future vignettes for
future plots; stay tuned).

If you specify multiple consecutive months (e.g. `4:6`), then the plot shows the anomalies
averaged over those months (and the title is automatically correct):
```{r, buoyanomaliesmonths}
plot(all_buoys_anomalies,
     months = 4:6)
```

The eagle-eyed amongst you will have spotted that the S. Georgia Strait buoy
does not appear on this plot (it was second from bottom in the previous
plots). We will come back to this later (briefly, the plot requires every
specified month in `months` to have an anomaly available, else it shows a blank).

And if you want a wintertime index that includes December, that can be done:
```{r, buoyanomaliesmonths2}
plot(all_buoys_anomalies,
     months = c(12, 1, 2, 3))
```

As stated in the title, the year on the x-axis is the year of the January.

If you specify `months` and a single buoy, the anomalies for those months are
all shown:
```{r, buoyanomaliesmonths3}
plot(all_buoys_anomalies,
     months = c(12, 1, 2, 3),
     stn_id = "Halibut Bank")  
```

This plot can be used as a check to see what anomalies are available (if you are
only interested in a particular buoy). It also shows explicitly that Dec-Mar is
missing at least one month in 1993 (i.e. Dec 1992 - Mar 1993) and 1994, but has
all four months available for 1995. That is why in the plot above that one,
"Halibut Bank" has no anomaly shown for 1993 or 1994, but does for 1995. 

### Returning the values shown in the plot

Setting `return_results = TRUE` returns a list containing the plot and then the
results used for the plot (and also automatically shows the plot). Repeating the
previous plot with this setting:

```{r, buoyanomaliesmonths4}
res <- plot(all_buoys_anomalies,
            months = c(12, 1, 2, 3),
            stn_id = "Halibut Bank",
            return_results = TRUE)  
res$results
```

So `res` is a list (`res$plot` will show the above plot again), and
`res$results` is a tibble that includes the values
in the plot, as `sst_plot_value` and then `label` for the rounded values that are shown
numerically. You can then use these for analyses if desired.

That tibble contains results that depend on options used in the plotting. But we
will now examine the full results generated at the very beginning in
`all_buoys_anomalies`.

### Examining the climatology and anomalies

At the beginning we ran
```{r, alreadyrun, eval = FALSE}
all_buoys_anomalies <- calculate_anomalies(buoy_sst)
```

which we have plotted using various options, and now examine
`all_buoys_anomalies` in detail:

```{r, buoyanomaliesdetails}
all_buoys_anomalies
```
This is a list object, with `$climatology` giving, for each buoy and month, a mean SST
(`clim_sst_mean`), standard deviation (`clim_sst_sd`), and sample size
(`clim_sst_n`) of the values used for that buoy-month combination.
The default for `calculate_anomalies()` creates the monthly climatology for each
buoy based on the years 1991-2020 (details of other arguments will be
investigated later).

The `$anomalies` tibble gives, for each buoy-year-month combination, the anomaly
for that month compared to the climatology. The first buoy shown is "C46184"
which is the `r buoy_metadata$name[match("C46184", buoy_metadata$stn_id)]`
buoy. For October 1987 the mean SST was 10.1 degC (`sst_mean`), which was
calculated from 29 days (`sst_n`) in that month. The resulting anomaly from the
climatology is -1.16 (`sst_anomaly`). This can be manually checked -- the
climatology for buoy C46184 in October is a mean SST of 11.3 (the top tibble),
the mean SST in October 1987 was 10.1, which is -1.16 less (`sst_anomaly`). This
should help you understand how calculations are done. 

So the same calculations are done for all buoy-year-month combinations. There
happen to be a few NA's early on in the data for this buoy (more on that later).

The vector `$climatology_years` gives the years prescribed for the
climatology in the `calculate_anomalies()` call, for which the default is `1991:2020`.

The attribute shown (under `attr(, "class")`) given just tells R to use our customised plotting
function on these results, which is automatically called with `plot()`.

There is more functionality already in the functions, which I will add to
soon (I have examples further down in this .Rmd file, but they are not rendered). Hopefully this is enough for users to start playing around with.

```{r, stop, cache = FALSE}
knitr::knit_exit()
```

More ideas I have kept track of and want to tidy up to include properly in the
vignette.

HERE

- Check each plotting option has been shown, at least mention the other ones.

- 
TODO confirm understanding - average the Apr anomalies for ECOBUOY, should be 0 as
they comprise the full climatology.


TODO move to later after have talked about the outputs.

TODO HERE (when doing 4:6) Look at South Nomad and plot it's full data. 2014 only has April (out
of Apr-Jun) and 2016 has no April, hence the anomalies are a bit
misleading. Suggest we do insist that all months have values, or at least give
an option to require a certian proportion, default 100%. 
Do an example with months 1:12.

Example of no climatology for some months for S. Georgia Strait buoy, so when
do:
```{r, buoyanomaliesmonths46}
plot(all_buoys_anomalies,
     months = 4:6)
```

we lose that buoy in the plot because it does not have a climatology for Jun
therefore we cannot do a mean climatology for Apr-Jun. 

The climatology is:

```{r, noclimatology}
all_buoys_anomalies$climatology %>%
  filter(stn_id == "C46303") %>%
  a()
```
and the anomalies for 2022 (as an example) are:
```{r, noclimatology2}
all_buoys_anomalies$anomalies %>%
  filter(year == 2022,
         stn_id == "C46303") %>%
  a()
```
where you can see that Jun and Jul have SST values, but no anomaly because there
is no climatology for those months for this buoy (due to limited data, and the
default climatology years of 1991-2020). 

However, if we specify that we only require 2 months to be available to make an anomaly for Apr-Jun
```{r, buoyanomaliesmonths46needless}
plot(all_buoys_anomalies,
     months = 4:6,
     require_requested_months = 2)
```




```{r, stop33, cache = FALSE}
knitr::knit_exit()
```



```{r, saveplotmaybe, echo = FALSE}
# ggsave(filename = "Annual_heatmap_SST_anomaly_SOPOregions_allyears.png", width = 9, height = 3., dpi = 300,
#       scale = 1.4)
```

How about the average anomaly for April to June for all buoys:
```{r, buoyanomalyplot2}
plot(all_buoys_anomalies,
     months = 4:6)
```

Other options, to add to the help and the testing:
```{r, buoyanomalyplot4}
plot(all_buoys_anomalies,
     sst_plot = "mean",
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     months = c(12, 1, 2))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184")

plot(all_buoys_anomalies,
     sst_plot = "count")  # check what this is doing, mean
# count over all years?? Ah no, it's april counts each year each buoy.

plot(all_buoys_anomalies,
     sst_plot = "count",
     count_breaks = c(0, 20, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     stn_id = "C46184",
     count_breaks = c(0, 15, 31))

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8)

plot(all_buoys_anomalies,
     sst_plot = "count",
     return_results = TRUE,
     months = 6:8,
     stn_id = "C46184")
```


All buoys, but no 15-day requirement in the climatology or anomalies, instead use
all days:
```{r, usealldays
all_buoys_anomalies_all_days <- calculate_anomalies(buoy_sst,
                                                  min_days_per_month = 0,
                                                  max_consecutive_missing_days =
                                                  32)
plot(all_buoys_anomalies_all_days)

# expect_equal(all_buoys_anomalies, all_buoys_anomalies_all_days) worked when
# first arg used the original function before we added this in
```

And here is the default with the cut off of requiring 15 days:
```{r, need15days}
plot(all_buoys_anomalies)
```

For example, Central Dixon Entrance has a few years that appear when using all
days but not when using the 15-day-per-month cut off (remember above two plots
are for April). Can look at that in
detail, by using the `min_days_per_month = 0` results:
```{r, cde}
plot(all_buoys_anomalies_all_days,
     stn_id = "C46145",
     sst_plot = "count",
     count_breaks = c(0, 15, 31))
```

Now doing the max consecutive missing days:
```{r, maxmissing}
all_buoys_anomalies_all_gaps_allowed <- calculate_anomalies(buoy_sst,
                                                  max_consecutive_missing_days =
                                                    32)
# expect_equal(all_buoys_anomalies, all_buoys_anomalies_all_gaps_allowed) worked when
# first arg used the original function before we added this in
```

Need to confirm that does what we want.
Buoy C46004 has NA's from 1989-11-06 to 1990-07-11, also a run of 7 in August 1990. Let's force 1989 to be in
climatology and check it those end months get properly ignored.

```{r, climtest}
test_buoy <- filter(buoy_sst,
                    stn_id == "C46004")
test_buoy_anomalies <- calculate_anomalies(test_buoy,
                                         climatology_years = 1989:1995)
plot(test_buoy_anomalies)

test_buoy %>% filter(date < "1992-10-01") %>% a()

test_buoy_anomalies$anomalies[1:50,] %>% a()
```
That shows some that are wrong. For example
```{r, wrong}
test_buoy_march_april_1991 <-
  filter(test_buoy,
         date >= "1991-03-01" & date <= "1991-04-30")

# Can see a streak of 3 NA's and one of 4 NA's, with 24 non-NA's in March, but
# April just has the one NA ...
test_buoy_march_april_1991 %>% a()  

# ... so the default settings do work, but if we get stricter then they do not
# for March:
calculate_anomalies(test_buoy_march_april_1991)
calculate_anomalies(test_buoy_march_april_1991,
                  min_days_per_month = 25)
calculate_anomalies(test_buoy_march_april_1991,
                  max_consecutive_missing_days = 4) # 4 NA streak allowed
calculate_anomalies(test_buoy_march_april_1991,
                  max_consecutive_missing_days = 3) # no climatology

    
```

Also maybe do some of (this may be okay now, was while figuring out gaps)
Why is C46004 March 1991 getting excluded? Should not be. Also:
 C46004 1992     1       NaN     0          NaN  
 C46004 1992     7       NaN     0          NaN   Though if max gaps is 7 then
 it should not be excluded, check.


Workflow should be:
- do the calculations with the defaults
- plot the counts for that buoy, with count_breaks as 0,15,31 (could make the
  default)
- check if that looks satisfactory, both in terms of calculating the climatology
  and for doing the anomalies. Hard to automate, so just have to inspect
  manually (e.g. not too many big gaps).
- change `min_days_per_month` value if possible. Try 0 at first as an extreme to see if that helps enough.

Now add max consecutive number of NA's in a row, within a month-year
combination. And suggest that check also.

Do single buoys also.

We can exclude the three buoys that do not have a long time series (this might
change when we refine the analyses):

don't need I think:

```{r, buoyanomalyplot3}

all_buoys_plot_3 <- plot(all_buoys_anomalies,
                         months = 4:6)
all_buoys_plot_3
```

Can do various options. 

Have averaged the anomalies for multiple months.


### Single buoy -- show each month

And if just one stn_id then plot months on the y-axis, January at the top:

```{r, singlebuoy}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206")
```
Hopefully self explanatory. 

Thinking of quick look at a relevant buoy for hake migration north with respect
to hake catches in Canada (which aren't in pacea but could be):
```{r, singlebuoy2}
plot(all_buoys_anomalies,
     stn_id_to_plot = "C46206",
     months = 4:6)
```
If hake catches were added on then would be easy to spot any pattern.

Get the plots working when `climatology_time = "week"`.



### Refining the analyses

Important to be aware of:

Above all assumes our defaults in the calculations as specified in
`calculate_anomalies.pacea_buoy()`. These are somewhat subjective and can be
changed by the user. Specifically:

- the number of days needed within a calendar month to use to give a monthly
  average for calculating climatoloties and anomalies (default is 15 days)
- the maximum consecutive number of missing days allowed within a calendar
  month, above which that month is not used to give a monthly
  average for calculating climatoloties and anomalies (default is 6 days)
  
We pragmatically decided on requiring 15 days (roughly half) of daily SST within
a month, and a maximum consecutive number of missing days of 6. These are the
defaults, but can be changed by the user (as described below) to `NULL` to allow
all data and then examine it for plotting.

Similarly, if a user wants to look at average SST for a buoy over the months
Jan-Jun, a decision has to be made as to how many months should be
available. Could say a default of >75%, with no details about requiring
consecutive ones. Maybe they should just look at the data and decide.

Given the huge number of options that users may wish to explore (many buoys,
many years, many possible combinations of months), users may wish to relax or
tighten the defaults, and have the power to do so.



### Daily swings

Issue #26 regards what to do with big daily swings that might be air temps
instead, had thought these would only come after some NA's or at the start of a
buoy's time series. Was going to take a look, but looking at the actual swings
we have in the daily data (grouping by buoy):

```{r, dailyswings}
max_daily_swing_by_buoy <- buoy_sst %>%
  group_by(stn_id) %>%
  mutate(daily_increase = sst - lag(sst)) %>%
  group_by(stn_id) %>%
  summarise(max_daily_increase = max(daily_increase,
                                     na.rm = TRUE),
            max_daily_decrease = min(daily_increase,
                                     na.rm = TRUE))

max_daily_swing_by_buoy
summary(max_daily_swing_by_buoy)
```
it looks like we have nothing unreasonable. Asked Andrea on #26 so see what she
says, but think we're all good and can close that.




### Rough notes, being adopted above somewhat


This works, but not for plotting yet, and may need to add options for how many
days needed in a week.

all_buoys_anomalies_weekly <- calculate_anomalies(buoy_sst, climatology_time = "week")

Need to remove data-poor months. Have some defaults that users can change, and
tell users to look at data. Elise thinks at least half the month, with no gap of
more than 5-6 days (no consecutive NA's longer than that).
Example we had was

```{r ex}
> yy <- table(all_buoys_anomalies$anomalies$sst_n)
> yy

   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14   15 
 777   32   19   29   26   24   29   21   24   26   26   26   36   19   34   34 

  16   17   18   19   20   21   22   23   24   25   26   27   28   29   30   31 
  57   68   72   59   76   69   94  113  149  171  346  301  727  682 1341 1798 

sum(as.data.frame(yy)$Freq)
[1] 7305
> sum(as.data.frame(yy)[21:32, ]$Freq)   # How many overall months have >=20
days of values.
[1] 5867
> 
```

The one showed up when we were playing around was:
filter(buoy_sst, stn_id == "C46183", year(date) == 2019, month(date) == 6) %>% a()
Has 13 missing days but they are kind of spread out, so looks usable. Hence go
with half a month needed, but no big gaps.

plot(all_buoys_anomalies, months = 1:12)  # North Hecate St 2019 has a big
anomaly, but it's only from that one month of data, as seen clearly in:
plot(all_buoys_anomalies, stn_id = "C46183")

Also need to think about climatologies. And whether to put the thinking part in
the climatology calculations and/or the plotting. Have to be in the climatology
calculations, as anomalies will change when we change the rules. And then all
resulting monthly climatologies/anomalies will be valid.

In the plotting we need a condition - how many months out of the requested ones
need to have anomaly calculations? Since monthly anomalies will have been calculated.

Could do the heatmap type plot on weekly or even daily values (for a single
buoy) to get an idea of the amount of missing data. Daily would be interesting.


Need to think about what then needs to be done in the anomaly calculation
function. Think have to do go back and forth a bit, therefore need a decent vignette.

- also look at biggest daily swings in temperature, and check if they're always
  at the start of the time series for a buoy. Implies turned on while on boat,
  bit like the albatrosses.

## Weekly anomalies - delete?

Travis's early code can do weekly, it just will not have all the extra
  functionality we have shown here (but we can modify  
