---
title: "Ecosystem Summary Heatmaps"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ecosystem Summary Heatmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("ecosystem_summary_heatmaps.Rmd")
or click the knit button in RStudio -->


```{r, start, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 14,   # 14 #5.7,
  fig.height = 5.25    # 7  #5  
)
fig_height_two_row_heatmap <- 2.7    
fig_height_three_row_heatmap <- 3  #
fig_height_five_row_heatmap <- 4   # Looks good, scale others to this kind of
fig_height_six_row_heatmap <- 4.33 
fig_height_ten_row_heatmap <- 6.2   

fig_height_one_map <- 4
fig_width_one_map <- 4
fig_height_two_map <- 4
fig_height_three_map <- 4       # width depends how many in each row, so do manually



```

## Customisable ecosystem summary heatmaps

```{r setup}
load_all()    # for developer
# library(pacea)  # for user
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In the
[buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we showed how to make customisable heatmaps of the buoy sea-surface
temperature data that are saved in pacea. They are in a format famililar to
oceanographers, using prescribed climatologies and subsequent anomalies.

Here we show how to produce customisable heatmaps that include disparate types
of ecosystem information, thus bringing together the various data objects in
pacea. Importantly the user can choose:

- the types of ecosystem information to be included, tailored to their
  particular stock;
- the area over which spatial information is summarised.

The aim of pacea is to
facilitate inclusion of ecosystem information in advice to fisheries management,
such as stock assessments. Given the multitude of stocks in DFO's Pacific
Region, both in terms of the diversity of species and the many stock-specific spatial areas, we decided
to create functions for user-customisable plots that easily summarise the relevant ecosystem
information.

Expert knowledge can be used to suggest relevant ecosystem information to be
used for a particular stock. Futhermore, the ease of generating the heatmaps can
allow exploration, formulation of ideas, and quick hypothesis testing.

For comparing disparate types of ecosystem information, we calculate an annual
index which is standardised (by subtracting the mean and dividing by the standard deviation across
the time series; note that all years might not get shown in a plot).
This is so that indices
are all dimensionless and can be compared on equivalent scales, and shown on a
single heatmap style plot.

As an example, let's use some example ecosystem information for the Strait of Georgia
(SOG) that is in pacea. We create an index for each type of ecosystem information in turn and
then make the plot. We just need to use `create_index()` multiple times to
create the indices, and then `plot_indices()` to make the function (we will
explain the functions in more detail later):
```{r, examples}
# Average sea surface temperature in April from a buoy in the SOG
sst_buoy_halibut_bank_apr_index <- create_index(buoy_sst,
                                                months = 4,
                                                stn_id = "Halibut Bank")

# Large calanoid copepods in the SOG 
calanoid_copepods_large_sog_index <- create_index(zooplankton_sog,
                                                  zooplankton_species_group =
                                                    "calanoid_copepods_large")

# Total zooplankton in the SOG
total_zooplankton_sog_index <- create_index(zooplankton_sog,
                                            zooplankton_species_group = "total_biomass")

# Annual age-2 herring recruitment in the SOG
herring_recruitment_sog_index <- create_index(herring_recruitment,
                                              herring_region = "SOG")

# Annual herring spawning biomass in the SOG
herring_biomass_sog_index <- create_index(herring_spawning_biomass,
                                          herring_region = "SOG")

# Annual population estimates of Harbour Seals in the SOG
harbour_seals_sog_index <- create_index(harbour_seals,
                                        seal_region = "SOG")
```
See the
relevant vignettes and help files for details of the pacea objects used for each
index. The first index is from the buoy sea surface temperature. Looking at the [buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we have used the Halibut Bank buoy in the SoG, which has good timespans of
data, and just pick April. 

The `create_index()` function automatically creates an object of class
`pacea_standardised_index`, no matter what type the ecosystem information is. 

Now make the plot, manually basing the start year on the shortest
of those (zooplankton data):
```{r, sogplot, fig.height = fig_height_five_row_heatmap}
plot_indices(sst_buoy_halibut_bank_apr_index,
             calanoid_copepods_large_sog_index,
             total_zooplankton_sog_index,
             herring_recruitment_sog_index,
             herring_biomass_sog_index,             
             harbour_seals_sog_index,
             years = 1996:2026,
             gap_after = c("sst_buoy_halibut_bank_apr_index",
                           "total_zooplankton_sog_index",
                           "herring_biomass_sog_index"))
```

The first six arguments are the indices that we created above, and `years`
restricts the plot to the years specified.
To delineate different types of ecosystem information in the plot
(e.g. here we have temperature, zooplankton, fish, mammals), we used the optional
`gap_after` argument to add a small gap after the specified indices.

Each index is standardised over its full time series, and red for a year indicates
above-average conditions, blue indicates below-average, with the black numbers
giving the exact standardised anomaly (number of standard deviations above the
mean), as used for the colour coding. 

The restriction of the years being shown is why the seal index is always
positive. By definition the mean of an index across its full time series 
is 0, and so the seal index must be negative earlier in the time series. A quick check of
the absolute seal populations numbers confirms this, in that the  population has
indeed 
increased from the early years:

```{r, sealplot, fig.width = 5, fig.height = 4}
plot(harbour_seals,
     region = "SOG")
```

Here are the details of what is in the buoy SST index object created above:
```{r, indexdetails}
sst_buoy_halibut_bank_apr_index
```
It is a tibble, with a value for each year, plus an `index` to refer to in code,
and an `index_label` to use for the labels in the plot (these are repeated in
each row, which just makes things easier at the plotting stage). 

We can chose multiple months. Note that `index` and `index_label`
get automatically prescribed;
examine the arguments and results for these three indices (for the default buoy,
namely that 
in West Dixon Entrance):
```{r, createindmonths}
create_index(buoy_sst,
             months = 5)
create_index(buoy_sst,
             months = 5:6)
create_index(buoy_sst,
             months = c(11, 12, 1, 2, 3))
```
In the third example note how the months are specified from November to March -
the year indicated is the year of the January.

Note that in the first example we have a value for 1992, but not in the second
one. This must be because June 1992 does not have enough daily values for this
particular buoy, which we can confirm from the raw data:
```{r, confirm}
filter(buoy_sst,
       stn_id == "C46205",   # default in create_index.pacea_buoy()
       date >= "1992-06-01",
       date <= "1992-06-30",
       !is.na(sst))
```
So there are only 6 daily SST values for this buoy in June 1992, and so an
anomaly is not calculated for that month because the default is to require 15
daily SST values in a month. We could force an anomaly to be
calculated by changing the arguments `min_days_per_month` (default 15) and 
`max_consecutive_missing_days` (default 6) that get passed onto
`calculate_anomalies.pacea_buoy()`:
```{r, confirmchange}
create_index(buoy_sst,
             months = 5:6,
             min_days_per_month = 6,
             max_consecutive_missing_days = 24)
```
thus giving an index value for 1992, given these definitions (which we do not
recommend, but are showing as examples to help understanding, and to show that
arguments get correctly passed onto further functions). 

For any type of pacea object, check the exact calculations and further options
by checking `class(object)` and then `?create_index.<class>`. For example
`?create_index.pacea_buoy` for the `buoy_sst` data.


## OISST spatiotemporal data

We now create various indices from the Optimally Interpolated Sea Surface
Temperature (OISST) spatiotemporal data. An index is a single standardised value
for each year, and for SST would be based on the mean anomalies over a given
area.

It is first best to plot the spatial anomalies for the area of interest, to
verify the area and understand what is going on. We wil then calculate an index
directly from the anomalies list, or 
directly from `create_index(oisst_month)` with various options.

<!-- which will then call `calculate_anomalies()` and then
`create_index.pacea_oisst_anomalies_list()` , but only return the index.
-->

First, let's look at the full domain of the OISST data and calculate annual
anomalies and do the default plot (for the most recent month of data):
```{r, oisst1, fig.height = fig_height_one_map}
oisst_anomalies <- calculate_anomalies(oisst_month)

plot(oisst_anomalies)
```

The `oisst_anomalies` list object (of class `pacea_oisst_anomalies_list`) contains the climatology for each month-location
combination, plus the anomalies relative to those for each year-month-location
combination, plus the `climatology_years` used:
```{r, oisst3}
oisst_anomalies
```

We can specify months and years in a plot, chosen here to show the marine heatwave forming
in 2014:
```{r, oisst2, fig.height = fig_height_three_map}
plot(oisst_anomalies,
     months = 1:12,
     years = 2014:2016,
     x_axis_labels = c(-135, -125)) # specify the x-axis labels to be shown     
```

We can create an index of the mean anomaly each year averaged over, say, Apr-Jun, over the
full domain shown (which is Canada's Exclusive Economic Zone):

```{r, fulldomain}
oisst_eez_apr_jun_index <- create_index(oisst_anomalies,
                                        months = 4:6)
oisst_eez_apr_jun_index
```
The index tibble has the same format as the other types shown above, with one
value per day. We can simply add this to our earlier heatmap summary plot:
```{r, sogplot2, fig.height = fig_height_six_row_heatmap}
plot_indices(oisst_eez_apr_jun_index,
             sst_buoy_halibut_bank_apr_index,
             calanoid_copepods_large_sog_index,
             total_zooplankton_sog_index,
             herring_recruitment_sog_index,
             herring_biomass_sog_index,             
             harbour_seals_sog_index,
             years = 1996:2026,
             gap_after = c("sst_buoy_halibut_bank_apr_index",
                           "total_zooplankton_sog_index",
                           "herring_biomass_sog_index"))
```
This is showing the proof of concept, in that, we can easily compare the
mean SST anomaly across the whole EEZ over April to June with the buoy SST mean
anomaly for April, at one location (the point being you can define an index
however you like and easily compare it). 

In practice, we likely want to restrict an index (and annual summary figure) to a
specific spatial region, and also specific months.

### Restricting an index over a spatial area

To create anomalies for a specific area, such as Pacific Marine
Fisheries Area 126, again showing
the heatwave progressing
```{r, area126plot, fig.height = fig_height_three_map}
oisst_anomalies_area_126 <- calculate_anomalies(oisst_month,
                                                area = pfma_area_126_temp)

plot(oisst_anomalies_area_126,
     months = 1:12,
     years = 2014:2016,
     restrict_plot = TRUE,          # restrict the plot to the area of the anomalies
     buffer_proportion = 0.6,       # add a buffer proportion to zoom out a bit
     x_axis_labels = c(-130, -128)) # specify the x-axis labels to be shown
```
(Some of those settings are hard to automate with multiple panels, so users may need to
tweak the options; also `pfma_area_126_temp` is a temporary data object in pacea
and will change names when we import many fisheries management areas).

For investigating plots (rather than doing analyses), we can also simply use the `area`
argument when plotting the full `oisst_anomalies`, giving the same plot as above
(so we have not shown it again here):
```{r, area126inplot, eval = FALSE}
plot(oisst_anomalies,
     area = pfma_area_126_temp,
     months = 1:12,
     years = 2014:2016,
     restrict_plot = TRUE,
     buffer_proportion = 0.6,
     x_axis_labels = c(-130, -128))
```

To calculate the index for Area 126:
```{r, oisstindex}
oisst_area_126_apr_index <- create_index(oisst_anomalies_area_126)   # April is the default

oisst_area_126_apr_index
```
For comparison we can also calculate the mean sea surface temperature for April from a buoy
within area 126
```{r, buoy126, fig.height = fig_height_two_row_heatmap}
sst_buoy_south_brooks_apr_index <- create_index(buoy_sst,
                                                months = 4,
                                                stn_id = "South Brooks")
plot_indices(oisst_area_126_apr_index,
             sst_buoy_south_brooks_apr_index)
```
This shows that the two different indices representing SST are fairly correlated
but not exactly the same, as expected.

<!-- TODO phase plot when done -->

It makes sense to restrict the plot to 1995 onwards when we have both types of
data. You can either restrict the years in the plot, or recalculate the OISST
anomalies from 1995 onwards. Note how these give slightly different figures;
restricting the plot:
```{r, buoy126b, fig.height = fig_height_two_row_heatmap}
plot_indices(oisst_area_126_apr_index,
             sst_buoy_south_brooks_apr_index,
             years = 1995:2026)
```
or recalculate a restricted index:
```{r, buoy126c, fig.height = fig_height_two_row_heatmap}
oisst_area_126_apr_from_1995_index <-
  create_index(oisst_anomalies_area_126,
               years = 1995:2026)

plot_indices(oisst_area_126_apr_from_1995_index,
             sst_buoy_south_brooks_apr_index)
```

You can also create an index directly from the `oisst_month` data, for which the
anomalies are calculated internally. Though it seems better to calculate
the full spatial anomalies for your area to take a look at them first, as we did
above. Nevertheless, `create_index()` will work directly on `oisst_month`, with
the `area` option, giving the same result as above
```{r, oisstindex2}
oisst_area_126_apr_index_directly <- create_index(oisst_month,
                                                  area = pfma_area_126_temp)

expect_equal(oisst_area_126_apr_index,
             oisst_area_126_apr_index_directly)
```

We can also show every month of a spatially averaged anomaly in a heatmap style
plot, like in the
[buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette. This nicely summarises the many spatial maps (which can't all easily
digested by a reader) into one plot. Note that these are NOT annual indices:

```{r, oisstheatmap}
plot(oisst_anomalies_area_126,
     heatmap_style = TRUE,
     area_name = "Area 126")
```

You can just do a few months of interest:
```{r, oisstheatmap2, fig.height = fig_height_three_row_heatmap}
plot(oisst_anomalies_area_126,
     heatmap_style = TRUE,
     area_name = "Area 126",
     months = 3:5)
```

See the code of the vignette (specifically the chunk headers) to see how to get
the figure sizes looking good; these are not automated in the functions.
