---
title: "Ecosystem Summary Heatmaps"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ecosystem Summary Heatmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("buoy_anomalies.Rmd")
or click the knit button in RStudio -->


```{r, start, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 14,   #5.7,
  fig.height = 7  #5
)
```

## Customisable ecosystem summary heatmaps

```{r setup}
load_all()    # for developer
# library(pacea)  # for user
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

In the
[buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we showed how to make customisable heatmaps of the buoy sea-surface
temperature data that are saved in pacea. They are in a format famililar to
oceanographers, using prescribed climatologies and subsequent anomalies.

Here we show how to produce customisable heatmaps that include disparate types
of ecosystem information, thus bringing together the various data objects in
pacea. Importantly the user can choose:

- the types of ecosystem information to be included, tailored to their
  particular stock
- the area over which spatial information is summarised

The aim of pacea is to
facilitate inclusion of ecosystem information in advice to fisheries management,
such as stock assessments. Given the multitude of stocks in DFO's Pacific
Region, both the diversity of species and the many stock-specific spatial areas, we decided
to create functions for user-customisable plots that easily summarise the relevant ecosystem
information.

Expert knowledge can be used to suggest relevant ecosystem information to be
used for a particular stock. Futhermore, the ease of generating the heatmaps can
allow exploration, formulation of ideas, and quick hypothesis testing.

For comparing disparate types of ecosystem information, we calculate an annual
index which is standardised (by subtracting the mean and dividing by the standard deviation across
the time series; all values might not get fully plotted). 
This is so that indices
are all dimensionless and can be compared on equivalent scales, and shown on a
single heatmap style plot.

As an example, let's use various ecosystem information for the Strait of Georgia
(SOG). We create an index for each type of ecosystem information in turn and
then make the plot. We just need to use `create_index()` multiple times to
create the indices, and then `plot_indices()` to make the function (we will
explain them in more detail later):
```{r, examples}
# Average sea surface temperature in April from a buoy in the SOG
sst_buoy_halibut_bank_apr <- create_index(buoy_sst,
                                          months = 4,
                                          stn_id = "Halibut Bank")

# Large calanoid copepods in the SOG 
calanoid_copepods_large_sog_index <- create_index(zooplankton_sog,
                                                  zooplankton_species_group =
                                              "calanoid_copepods_large")

# Total zooplankton in the SOG
total_zooplankton_sog_index <- create_index(zooplankton_sog,
                                            zooplankton_species_group = "total_biomass")

# Annual age-2 herring recruitment in the SOG
herring_recruitment_sog_index <- create_index(herring_recruitment,
                                              herring_region = "SOG")

# Annual herring spawning biomass in the SOG
herring_biomass_sog_index <- create_index(herring_spawning_biomass,
                                          herring_region = "SOG")

# Annual population estimates of Harbour Seals in the SOG
harbour_seals_sog_index <- create_index(harbour_seals,
                                        seal_region = "SOG")
```

Now make the plot, manually basing the start year on the shortest
of those (zooplankton data):
```{r, sogplot, fig.height = 4}
plot_indices(sst_buoy_halibut_bank_apr,
             calanoid_copepods_large_sog_index,
             total_zooplankton_sog_index,
             herring_recruitment_sog_index,
             herring_biomass_sog_index,             
             harbour_seals_sog_index,
             years = 1996:2026,
             gap_after = c("sst_buoy_halibut_bank_apr",
                           "total_zooplankton_sog_index",
                           "herring_biomass_sog_index"))
```

The first six arguments are the indices that we created above, and `years`
restricts the plot to the years specified.
To delineate different types of ecosystem information in the plot
(e.g. here we have temperature, zooplankton, fish, mammals), we used the optional
`gap_after` argument to add a small gap after the specified indices.

The restriction of the years being shown is why the seal index is always
positive. By definition the mean of the index 
should be 0, and so it is negative earlier in the time series. A quick check of
the absolute seal populations numbers confirms the population has indeed
increased:

```{r, sealplot, fig.width = 5.7, fig.height = 5}
plot(harbour_seals,
     region = "SOG")
```


TODO Add an option in the plot function to restandardise across the years specified?

(see the
relevant vignettes and help files for details of the pacea objects used here as
the first argument). The first is from the buoy sea surface temperature, looking at the [buoy_anomalies](https://pbs-assess.github.io/pacea/vignettes/buoy_anomalies.html)
vignette we will use the Halibut Bank buoy in the SoG, which has good timespans of
data, and just pick April. We then create indices for other types of ecosystem
information (we will do into details 

This is done in the
`create_index()` function, to create an object of class
`pacea_standardised_index`, no matter what type the ecosystem information is. 


Buoy SST getting working, need to prove if normalised index of mean anomalies is same
as normalised index of mean of means. Think so since both relative to the same
climatology. TODO. Also it's still based on climatology; does mean of actual
values just give the same?

```{r, buoysst}
sst_west_dixon_apr_using_anomalies <- create_index(buoy_sst)
sst_west_dixon_apr_using_mean <- create_index(buoy_sst, buoy_statistic = "mean")
expect_equal(sst_west_dixon_apr_using_anomalies,
             sst_west_dixon_apr_using_mean)
# Not convinced that will always hold though, I think only difference would be
# if you did not have a climatology for a particular month, which does not
# happen with the default years. So probably best to keep the generality. TODO
```

Some examples of different choices of months, and how `index` and `index_label`
get automatically prescribed:
```{r, createindmonths}
# Different examples:
create_index(buoy_sst,
             months = 5)
create_index(buoy_sst,
             months = 5:6)
create_index(buoy_sst,
             months = c(11, 12, 1, 2, 3))
```

Note that in the first example we have a value for 1992, but not in the second
one. This must be because June 1992 does not have enough daily values for this
particular buoy, which we can confirm from the raw data:
```{r, confirm}
filter(buoy_sst,
       stn_id == "C46205",   # default in create_index.pacea_buoy()
       date >= "1992-06-01",
       date <= "1992-06-30",
       !is.na(sst))
```
So there are only 6 daily SST values for this buoy in June 1992, and so an
anomaly is not calculated for that month. We could force an anomaly to be
calculated by changing the arguments `min_days_per_month` (default 15) and 
`max_consecutive_missing_days` (default 6) that get passed onto
`calculate_anomalies.pacea_buoy()`:
```{r, confirmchange}
create_index(buoy_sst,
             months = 5:6,
             min_days_per_month = 6,
             max_consecutive_missing_days = 24)
```
thus giving an index value for 1992, given these definitions (which we do not
recommend, but are showing as examples to help understanding, and to show that
arguments get correctly passed onto further functions).


Further examples (originally had above while generating the code), also check
the help file of `plot_indices()` and add to that TODO
```{r, furthrexamples}
herring_recruitment_hg_index <- create_index(herring_recruitment,
                                             herring_region= "HG")
herring_recruitment_prd_index <- create_index(herring_recruitment,
                                              herring_region = "PRD")
herring_recruitment_cc_index <- create_index(herring_recruitment,
                                             herring_region = "CC")
herring_recruitment_wcvi_index <- create_index(herring_recruitment,
                                              herring_region = "WCVI")
herring_biomass_wcvi_index <- create_index(herring_spawning_biomass,
                                           herring_region = "WCVI")

hake_recruitment_index <- create_index(hake_recruitment)

hake_biomass_index <- create_index(hake_biomass)

harbour_seals_wcvi_index <- create_index(harbour_seals)  

# Many examples (adding as go along), likely not really needed:
plot_indices(herring_recruitment_hg_index,
             herring_recruitment_prd_index,
             herring_recruitment_cc_index,
             herring_recruitment_sog_index,
             herring_recruitment_wcvi_index,
             herring_biomass_wcvi_index,             
             hake_recruitment_index,
             hake_biomass_index,
             harbour_seals_wcvi_index,
             harbour_seals_sog_index,
             total_zooplankton_sog_index)
```

See the test for more examples. Summarise here maybe. 


TODO should the years argument then require indices to be recalculated across
just those years? Question for the audience maybe..., show the consequences of
that. Historically good/bad versus short-term good/bad. Will need both options,
decide on default.


## OISST spatiotemporal data

Just doing examples here for now (might plot strange with so many plots), may
move to new vignette. You might want to do the anomalies first to look at them
spatially, and can then calculate an index directly from the anomalies list, or
directly from `create_index(oisst_month)` with options, which will then call `calculate_anomalies()` and then
`create_index.pacea_oisst_anomalies_list()` , but only return the index.

First, let's look at the full domain of the OISST data:
```{r, oisst1}
oisst_anomalies <- calculate_anomalies(oisst_month)

plot(oisst_anomalies)
```

The default above is for the most recent month of data. We can specify months
and years:
```{r, oisst2}
plot(oisst_anomalies,
     months = 1:12,
     years = 2014:2016)
```
That shows the marine heatwave forming. 

In practice, we likely want to restrict an index (and annual summary) to a
specific spatial region, and also specific months.

We can restrict the calculations to a specific area, such as Pacific Marine
Fisheries Area 126: (TODO we use a temporary data object for that is currently in
pacea, pending many areas being imported):
```{r, area126plot}
oisst_anomalies_area_126 <- calculate_anomalies(oisst_month,
                                                area = pfma_area_126_temp)

# Default plot still shows the full coast for the latest month of data
plot(oisst_anomalies_area_126)
```

We can show multiple months and years, and add some options to improve the figure:
```{r, oisst126plots}
plot(oisst_anomalies_area_126,
     months = 1:12,
     years = 2014:2016,
     restrict_plot = TRUE,          # restrict the plot to the area of the anomalies
     buffer_proportion = 0.6,       # add a buffer proportion to zoom out a bit
     x_axis_labels = c(-130, -128)) # specify the x-axis labels to be shown
```
(Some of those are hard to automate with multiple panels, so users may need to
tweak the options).

For investigating plots (rather than doing analyses), we can also use the `area`
argument when plotting the full `oisst_anomalies`, giving the same plot as above
(not shown here):
```{r, area126inplot, eval = FALSE}
plot(oisst_anomalies,
     area = pfma_area_126_temp,
     months = 1:12,
     years = 2014:2016,
     restrict_plot = TRUE,
     buffer_proportion = 0.6,
     x_axis_labels = c(-130, -128))
```



TODO HERE go through these, checking can do the anomalies and create an index,
and get the same index if just creating them directly from `oisst_month` with the
area option.

Having calculated the full anomalies, can create a variety of indices.

Now do the index, playing around for now:
```{r, oisstindex}
oisst_anomalies_area_126
xx <- create_index(oisst_anomalies_area_126)
# Taking area 126 then making a slightly smaller polygon
## area_as_list_of_matrix <-
##         list(matrix(c(-127.1506, -128.2331, -129.3492, -127.9167, -127.1847,
##                       -126.8200, -127.1506,
##                       49.85766, 49.00000, 48.99991, 50.11915, 50.40183,
##                       50.24466, 49.85766),
##                     ncol = 2))    # From Travis's OISST vignette
area_smaller_than_126_as_list <-
  list(matrix(c(-128, -129.3492, -127.9167, -127.1847,
                      -126.8200, -128,
                      49.85766, 48.99991, 50.11915, 50.40183,
                      50.24466, 49.85766),
               ncol = 2))

area_smaller_than_126_polygon <-
  sf::st_sfc(sf::st_polygon(area_smaller_than_126_as_list),
             crs = 4326) %>%
  sf::st_as_sf()

# Check we have less values:
oisst_anomalies_area_smaller_than_126 <-
  oisst_anomalies_area_126$anomalies[area_smaller_than_126_polygon, ]

# This has just 5 entries, area_126 has 20 (shown next). So fine for a test run.
oisst_anomalies_area_smaller_than_126 %>% filter(year == 1981, month == 9)

oisst_anomalies_area_126$anomalies %>% filter(year == 1981, month == 9)

# TODO try area option in create_index, see below

# Get it working for area 126, then for the defined subset of area26

oisst_area_126_apr <-
  create_index(oisst_anomalies_area_126)
oisst_area_126_apr

# Average sea surface temperature in April from a buoy within area 126
sst_buoy_south_brooks_apr <- create_index(buoy_sst,
                                          months = 4,
                                          stn_id = "South Brooks")
plot_indices(oisst_area_126_apr,
             sst_buoy_south_brooks_apr)

# TODO phase plot example when I've written it

# Then restrict to 1995 onwards as that's when the buoy starts
oisst_area_126_apr_from_1995 <-
  create_index(oisst_anomalies_area_126,
               years = 1995:2026)

plot_indices(oisst_area_126_apr_from_1995,
             sst_buoy_south_brooks_apr)

# Try smaller area option
oisst_area_smaller_than_126_apr_from_1995 <-
  create_index(oisst_anomalies_area_126,
               area = area_smaller_than_126_polygon,
               years = 1995:2026,
               index_name = "smaller_area",
               index_label = "Smaller area than 126")

oisst_area_smaller_than_126_apr_from_1995
plot_indices(oisst_area_126_apr_from_1995,
             oisst_area_smaller_than_126_apr_from_1995,
             sst_buoy_south_brooks_apr)

```

You can also create an index directly from the `oisst_month` data, for which the
anomalies are calculated internally. Though it might be a good idea to calculate
full spatial anomalies for your area to take a look at them first, as we did
above. But `create_index()` will work directly on `oisst_month`, with various options.
```{r, oisstindex}
oisst_area_smaller_than_126_apr_from_1995_directly <- create_index(oisst_month,
                                                                   area = area_smaller_than_126_polygon,
                                                                   years = 1995:2026,
                                                                   index_name = "small_direct",
                                                                   index_label =
                                                                     "Small area directly calculated")

oisst_area_smaller_than_126_apr_from_1995_directly   # gives same answer as earlier

expect_equal(oisst_area_smaller_than_126_apr_from_1995_directly$value,
             oisst_area_smaller_than_126_apr_from_1995$value)

# As expected, these are the same:
# plot_indices(oisst_area_smaller_than_126_apr_from_1995_directly,
#              oisst_area_smaller_than_126_apr_from_1995)

# Keeping plot.pacea_oisst_anomalies_list.Rd as an option, so explain that. TODO


```

How about showing every month of a spatially averaged anomaly in a heatmap style
plot - nicely abstracts many spatial maps into one plot. Can't show all the
spatial ones. (These are NOT annual indices).

```{r, oisstheatmap}
# Averages over the whole domain (default):

# TODO can add area as an argument
# Should be able to replace with
plot(oisst_anomalies_area_126,
     heatmap_style = TRUE,
     area_name = "Area 126")

# Just do a few months of interest:
plot(oisst_anomalies_area_126,
     heatmap_style = TRUE,
     area_name = "Area 126",
     months = 3:5)

```

## For new Area vignette


For new Areas vignette. TODO mention here.

We will have many fisheries management areas imported into pacea soon, explain
 those.
 
TODO:
update with new vignette when done. Copy the next bit to that to show how to
create your own area.

If necessary, you can create your own area (as an sf object) from a matrix or tibble of latitude
and longitude values, using our `create_area_sf_polygon()` function (if the
first and last rows are different then the function repeats the first row as the
last to make an enclosed polygon). Using a matrix or tibble gives the same result:
```{r, myarea}
my_area_matrix <- matrix(c(-130, -132, -130, -128,
                           48, 50, 52, 50),
                         ncol = 2)
my_area_matrix
my_area_1 <- create_area_sf_polygon(my_area_matrix)
my_area_1

my_area_tibble <- tibble::tibble(
  lon = c(-130, -132, -130, -128),
  lat = c(48, 50, 52, 50))
my_area_tibble
my_area_2 <- create_area_sf_polygon(my_area_tibble)
expect_equal(my_area_1,
             my_area_2)
```

Travis: 
Hmmm. Looking at the oisst data, the geometry coordinates are stored as points (centre of each cell). However with BCCM, they were staored as polygons (each corner of the cell). I'm not 100% sure why I did this but it's likely because of the different 2x2 and 6x6 grid we initially went with for bccm. If we just stored the centre of coordinates for the bccm data, it would reduce the size a bit, but I don't think by that much. 
 
st_intersects is TRUE if any part of the geometry is within your defined area. 
 
For POINTS data, st_intersects and st_within act the same as each 'cell' is just one coordinate point. Therefore the centre of the cell has to be within your defined area. 
 
If we had POLYGON geometries, then st_intersects and st_within act different. st_intersects is any one of the coordinates (corners) for each cell can be in the defined area for it to be TRUE. Whereas st_within requires all coordinates of the polygon (cell) to be in the area
 
Me: That makes sense, thanks. (The Sasquatch was super useful in noticing this). So doing the same thing for BCCM  will look a bit different.
Don't think I've downloaded BCCM data yet but plan to soon, had to clear up my full hard drive a bit.
 
Suggest we don't mess with the current storage of the data. I guess I can't make an option to make them agree (by switching between st_within and st_intersects, as neither will exactly match up). Okay, just have to mention this somewhere. In reality, it shouldn't make such a difference provided areas are not too small. And areas will usually be less topologically interesting than a sasquatch.
 


Restrict spatiotemporal data to an area:
```{r, areaplots}
oisst_sasquatch <- restrict_to_area(oisst_month, sasquatch)
plot(oisst_sasquatch,
     years = 2015:2016,
     months = 1:12)

oisst_area_126 <- restrict_to_area(oisst_month,
                                   pfma_area_126_temp)
plot(oisst_area_126,
     years = 2015:2016,
     months = 1:12)
```

```{r, anomaliessqtch}
oisst_anomalies_sasquatch <- calculate_anomalies(oisst_month,
                                                 area = sasquatch)
# TODO replace with plot, think NAMESPACE just temporarily messed up
plot.pacea_oisst_anomalies_list(oisst_anomalies_sasquatch)

plot.pacea_oisst_anomalies_list(oisst_anomalies_sasquatch, years = 2014:2026)

# In average years he hides away, so only comes out when it's warm or
# cold. Probably hibernating.

plot.pacea_oisst_anomalies_list(oisst_anomalies_sasquatch,
                                months = 9,
                                years = 2014:2025)


plot.pacea_oisst_anomalies_list(oisst_anomalies_sasquatch, heatmap_style = TRUE)


```

Now create an index
```{r, sasindex}
sasquatch_index <- create_index(oisst_anomalies_sasquatch,
                                months = 4,
                                index_label = "OISST in Sasquatch for Apr")
# TODO HERE redo the area126 stuff, can simplify lots of things above
oisst_area_126_anomalies <- calculate_anomalies(oisst_month,
                                                 area = pfma_area_126_temp)
# TODO see if can do it one go, without calculate_anomalies part.
oisst_area_126_index <- create_index(oisst_area_126_anomalies,
                                     months = 4,
                                     index_label = "OISST in Area 126 for Apr")

plot_indices(sasquatch_index,
             oisst_area_126_index,
             sst_buoy_halibut_bank_apr)
```


