---
title: "Fraser River Discharge Data"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fraser River Discharge Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("fraser_discharge.Rmd")
or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
)
```

```{r setup}
library(pacea)
library(dplyr)
library(tibble)  # Else prints all of a tibble
library(ggplot2)
```

## Fraser River discharge data sets

The Fraser River is the longest river in British Columbia and one of the most
important waterways in western Canada. Fraser River discharge is particularly
significant for salmon populations, as it is an important determinant of
survival for Fraser River Sockeye Salmon during juvenile migration from
freshwater to the ocean, and during return migration back to the spawning
grounds.

In pacea we include two Fraser River discharge data sets, both measured at Hope,
British Columbia:

1. **`fraser_discharge_mean`** - monthly mean discharge data
2. **`fraser_discharge_peak`** - monthly peak discharge data

Both data sets are derived from measurements
at the [Hope station](https://wateroffice.ec.gc.ca/map/index_e.html?type=historical&target_station=08MF005) and provide monthly time series from 1912 onwards,
downloaded from Environment and Climate Change Canada. The
data are of class `pacea_index` and contain measurements in cubic meters per
second (m³/s). See `?fraser_discharge_mean` for further details.

### Mean discharge data

The mean discharge values are the monthly averages:

```{r mean-data}
fraser_discharge_mean
tail(fraser_discharge_mean)
```

The columns are `year` and `month` (1-12), and `value`, which is the mean over
the month of the daily means (calculated from measurements throughout the day), in m³/s.

```{r mean-summary}
summary(fraser_discharge_mean)
```

Built-in plotting functions are included in pacea, with automatic axis labelling:
```{r plot-mean}
plot(fraser_discharge_mean)
```

This plot shows the long-term patterns in monthly mean discharge, revealing both seasonal cycles and inter-annual variability.

### Peak discharge data

The peak discharge data are monthly peaks calculated from daily mean values:

```{r peak-data}
fraser_discharge_peak
tail(fraser_discharge_peak)
```

The columns are `year` and `month` (1-12), and `value`, which is the peak over
the month of the daily means (calculated from measurements thoughout the day),
in m³/s.

```{r peak-summary}
summary(fraser_discharge_peak)
plot(fraser_discharge_peak)
```

The peak discharge plot shows the variability in monthly maximum flows, which is important for understanding flood conditions and extreme hydrological events.

## Example analyses

That desribes the two data objects. The rest of this vignette provides example
analyses that can be conducted.

### Seasonal patterns

The Fraser River exhibits strong seasonal discharge patterns due to its
snowmelt-dominated hydrological regime:

Calculate long-term statistics of monthly means and then monthly peaks:
```{r seasonal-patterns}
monthly_mean_stats <- fraser_discharge_mean %>%
  group_by(month) %>%
  summarise(
    avg = mean(value),
    min = min(value),
    max = max(value)) %>%
  ungroup() %>%
  mutate(month_name = month.name[month])

monthly_mean_stats

monthly_peak_stats <- fraser_discharge_peak %>%
  group_by(month) %>%
  summarise(
    avg = mean(value),
    min = min(value),
    max = max(value)) %>%
  ungroup() %>%
  mutate(month_name = month.name[month])

monthly_peak_stats
```

Now plot the average seasonal cycle
```{r seasonal-cycle, fig.height = 9}
par(mfrow = c(2, 1))

y_lim = c(0,
          max(c(monthly_mean_stats$avg, monthly_peak_stats$avg)))

plot(monthly_mean_stats$month,
     monthly_mean_stats$avg,
     type = "l",
     lwd = 2,
     xlab = "Month",
     ylab = "Average daily mean discharge (m³/s)",
     ylim = y_lim,
     main = "Average seasonal cycle - mean discharge",
     xaxt = "n")
axis(1, at = 1:12, labels = month.abb)

plot(monthly_peak_stats$month,
     monthly_peak_stats$avg,
     type = "l",
     lwd = 2,
     col = "red",
     xlab = "Month",
     ylab = "Average peak discharge (m³/s)",
     main = "Average seasonal cycle - peak discharge",
     ylim = y_lim,
     xaxt = "n")
axis(1, at = 1:12, labels = month.abb)
```

### Months with highest mean or highest peak discharge:

```{r extreme-events}
fraser_discharge_mean %>%
  arrange(desc(value))

fraser_discharge_peak %>%
  arrange(desc(value))
```

### Months with lowest mean or lowest peak discharge:

```{r low-flows}
fraser_discharge_mean %>%
  arrange(value)

fraser_discharge_peak %>%
  arrange(value)
```

### Annual statistics

Annual statistics based on the monthly means, and also the annual peak (being the
maximum monthly peak each year):

```{r annual-patterns}
annual_mean_stats <- fraser_discharge_mean %>%
  group_by(year) %>%
  summarise(
    annual_mean = mean(value),
    annual_min = min(value),
    annual_max = max(value)) %>%
  ungroup()

# Calculate annual peaks
annual_peak <- fraser_discharge_peak %>%
  group_by(year) %>%
  summarise(
    annual_peak = max(value)) %>%
  ungroup()

annual_summary <- annual_mean_stats %>%
  left_join(annual_peak,
            by = "year")

annual_summary
annual_summary %>% tail()
```
