---
title: "Population Estimates"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("populations.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

## Estimates of animal populations

```{r setup}
library(pacea)
library(dplyr)
```

In pacea we include estimates of populations, including uncertainty, as calculated from recent
stock assessments.

Currently, we have Pacific Hake, a major component of the ecosystem, and Pacific Harbour Seal, the most
abundant pinniped species in the Northeast Pacific.

NEW (since original release of pacea): we have updated the Pacific Hake
estimates with those from the 2024 assessment, but retained the 2023
assessment's estimates within the package (as `hake_biomass_2023` etc.). See
below and the help files for more details.

## Pacific Hake

The hake biomass time series is saved as a `tibble` in `pacea`, as simply:
```{r hakebiomass}
hake_biomass
tail(hake_biomass)
```
with the annual estimates given as median, low and high ends of the 95% credible
intervals. Units are millions of tonnes of female spawning biomass at the start
of the year, as documented in `?hake_biomass` which also refers to
`?hake_recruitment` for full details and further background. For example, it is important to
realise that the assessment considers a coastwide stock off the west coast of
Canada and the United States.

To plot the time series of estimated coastwide biomass,
simply do
```{r hakebiomassplot}
plot(hake_biomass)
```

where the solid circles are the annual medians, with credible intervals given by
the shaded regions (mostly reproducing Figure d in the latest assessment).

Note that the `hake_biomass` object also has class `pacea_biomass`:
```{r hakebiomassclass}
class(hake_biomass)
```
such that the above command `plot(hake_biomass)` first looks to use our tailored function
`plot.pacea_biomass()`, which gives the style of plot shown above. We use this
approach for `hake_recruitment` and various other objects in `pacea`, to enable
easy default plotting.

The `hake_recruitment` object contains annual estimates of recruitment of age-0
fish
```{r hakerec}
hake_recruitment
tail(hake_recruitment)
```
for which values are billions of age-0 fish, and our default plot is
```{r hakerecplot}
plot(hake_recruitment)
```

The dots represent the medians, with blue bars showing the 95% credible intervals.
This reproduces Figure f in the latest stock assessment (without some extra
details).

Note that, as mentioned in `?hake_recruitment` the most recent few years are
very uncertain as they are not informed by data. In particular, the final year
represents recruitment at the start of the year but has no data to inform it,
and so should definitely not be used for any analyses (we include it as it is
shown in the stock assessment; it also essentially represents the prior distribution used for recruitment).

The estimated uncertainty of recruitments is large, and so we also include
scaled values that make it easier to compare years. The 2010 recruitment is the
second largest in the time series (below only 1980's) and as such people have an
intuition that it is large, and we can compare other years by looking at
recruitment scaled by the 2010 recruitment:
```{r hake2010}
hake_recruitment_over_2010
tail(hake_recruitment_over_2010)
plot(hake_recruitment_over_2010)
```

This shows the medians and 95% credible intervals (red lines) of the recruitment
divided by that in 2010, which clearly shows that, for example, the 2014 is
clearly smaller than the 2010 recruitment (highlighted by the green dashed
line), which is not intuitive in the above `plot(hake_recruitment)` plot (see
`?hake_recruitment_over_2010` for more details and motivation).

This approach
somewhat scales out some of the uncertainty in the absolute scale of $R_0$, the
mean unfished equilibrium recruitment. This is similarly seen by instead scaling
recruitments by $R_0$, as
```{r hakeR0}
hake_recruitment_over_R0
tail(hake_recruitment_over_R0)
plot(hake_recruitment_over_R0)
```

### Updates due to new stock assessments

Note that the Pacific Hake stock assessment is conducted annually at the start of each year,
and these estimates will be updated in `pacea` by April each year. So
estimates will change (particularly for the most recent years) as more data
inform the stock assessment model. So `hake_biomass`, `hake_recruitment`
etc. will be based on the most recent stock assessment (see
`?hake_recruitment`).

To ensure that estimates from previous assessments are
still available in `pacea` (so you can ensure fully repeatable analyses that do
not change when `pacea` is updated with the new estimates), we also save each
assessment's estimates with the year appended; for example, `hake_biomass_2023`
and `hake_biomass_2024` for the spawning biomass estimates from the 2023 and
2024 assessments. Similarly, we also save `hake_recruitment_2023`,
`hake_recruitment_2024`, `hake_recruitment_over_2010_2023`,
`hake_recruitment_over_2010_2024`, `hake_recruitment_over_R0_2023`, and
`hake_recruitment_over_R0_2024`, plus later years once they become
available. See `?hake_recruitment` for details and references. The above
examples will work on all these objects, for example, the values and plot shown
in the previous code chunk, but based on the 2023 assessment results, are:
```{r hakeR02023}
hake_recruitment_over_R0_2023
tail(hake_recruitment_over_R0_2023)
plot(hake_recruitment_over_R0_2023)
```

Note how `hake_recruitment_over_R0_2023` contains no estimate for the 2024
recruitment, but `hake_recruitment` does (although it is not actually informed
by data, so should not be used in any analysis). And see how the (highly
uncertain) 2020 and 2021
recruitment estimates changed somewhat from 2023 assessment to the 2024
assessment, due to updated data:
```{r hake2020}
filter(hake_recruitment_2023, year %in% c(2020, 2021))
filter(hake_recruitment_2024, year %in% c(2020, 2021))
```
See page 13 of the 2024 assessment document for explanation.

## Pacific Harbour Seals

Estimated abundances for seven regions (and coastwide) were calculated by DFO
(2022; SAR 2022/034), and are included in `pacea` as:
```{r seals}
harbour_seals

tail(harbour_seals)
```
Estimates are for each region and for the coastwide population, reproducing Figure 3 of DFO (2022):

```{r sealplot}
plot(harbour_seals)
```

Or just plot estimates for a single region:
```{r sealplot2}
plot(harbour_seals,
     region = "SOG")
```

For full details and reference see
```{r sealhelp, eval = FALSE}
?harbour_seals
```

We anticipate adding assessment output for other species soon, so stay tuned
(check the NEWS). For zooplankton anomalies see the new zooplankton vignette:
[zooplankton.html](http://htmlpreview.github.io/?https://github.com/pbs-assess/pacea/blob/main/vignettes/zooplankton.html).
