---
title: "Zooplankton anomalies in the Strait of Georgia"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Zooplankton anomalies in the Strait of Georgia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "Last rendered on `r format(Sys.time(), '%d %B, %Y')`"
---

<!-- To build either run
rmarkdown::render("zooplankton.Rmd")
 or click the knit button in RStudio -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)
```

## Zooplankton anomalies in the Strait of Georgia

```{r setup}
library(pacea)
library(dplyr)
```

Pacea now includes anomalies of zooplankton biomass in the Strait of Georgia,
BC, provided by Kelly Young (Institute of Ocean Sciences, DFO). These update the calculations
presented by Perry et al. (2021) and in the annual State of the Pacific Ocean
report (Young et al., 2023); see `?zooplankton` for full details and
latest citations. Here we explain what is included, and update the figures presented
by Perry et al. (2021) for total biomass and 20 species groups.

The data are stored as a `tibble` in `pacea`, as simply:
```{r zoo}
zooplankton_sog
```
with the final years being
```{r zoofinal}
tail(zooplankton_sog)
```
The first three columns are `year` (the year of the calculated anomalies),
`number_samples` (total number of samples used for the annual anomalies), and
`volume_filtered` (total volume of water filtered for all tows, in m$^3$). The
`number_samples` column matches the `Total` column in Table 3 of Perry et al. (2021),
though only up to 2013; from 2014-2018 the numbers here are higher due to further
samples being processed. Data from 2019 onwards were not available for the
Perry at al. (2021) study.

The remaining columns correspond to anomalies for `total_biomass` (the total
biomass of all zooplankton) or various species groups, given as annual anomalies
in units of $\log10~\mbox{g m}^{-2}$. See `?zooplankton_sog` for more details,
and Perry et al. (2021) for the full calculations. (As ever in `pacea`, please
read through the help files for any data objects you are using).

The full available species
groups (plus the first three columns) are:
```{r zoonames}
names(zooplankton_sog)
```
The descriptions are fairly self explanatory, and are explained in more detail in Perry et
al. (2021).

`pacea` has a built-in plotting function, so users can simply plot the total
biomass (the default) with
```{r zooplot}
plot(zooplankton_sog)
```
where the bars are automatically coloured red/blue for positive/negative
anomalies. The plot defaults to showing the total biomass anomaly, and so
updates Figure 2D of Perry et al. (2021), which included data only to 2018. Note
that anomalies are now calculated compared to a climatology for the full time
series (Young et al. 2023), whereas Perry et al. (2021) used the years 1996-2010
for their climatology, and extra historical samples have been processed (see
`?zooplankton_sog`).  Hence, numbers can differ when comparing new plots to
older ones.

So, to plot one of the other species groups (`total_biomass` is the default, and
so was plotted above):
```{r zooplot2}
plot(zooplankton_sog,
     species_group = "cladocera")
```

Supplementary Figure S1 of Perry et al. (2021) showed the anomalies for 20
species groups (almost all those in our data). We produce an updated version
here, arranging the plots in the same way as Perry et al. (2021) for ease of
comparison. For simplicity we have included a column `in_perry_fig_s1` in the
data object `zooplankton_sog_axis_names`, which we an use to filter the required
species names.
```{r zooplot3}
sp_groups <- filter(zooplankton_sog_axis_names,
                    in_perry_fig_s1)$species_group_name
lwd_val <- 7      # Thickness of the bars, tweak depending on output
```
Now plot in three panels:
```{r zooplot4, fig.height = 14}
par(mfrow = c(4, 2))
for(i in 1:8){
  plot(zooplankton_sog,
       species_group = sp_groups[i],
       lwd = lwd_val)
}
```

```{r zooplot5, fig.height = 14}
par(mfrow = c(4, 2))
for(i in 9:16){
  plot(zooplankton_sog,
       species_group = sp_groups[i],
       lwd = lwd_val)
}
```

```{r zooplot6, fig.height = 7}
par(mfrow = c(2, 2))
for(i in 17:20){
  plot(zooplankton_sog,
       species_group = sp_groups[i],
       lwd = lwd_val)
}
```

All species groups show positive anomalies in recent years, continuing the
pattern shown in Figure S1 of Perry et al. (2021). The only group in that figure
that ended with a negative anomaly (in 2018) was Chaetognatha; the run of negative
anomalies from 2016-2018 show up here, but are followed ever since by positive anomalies.

The remaining species groups not shown (including Total biomass, even
though it was already shown above), are:
```{r zooplot7, fig.height = 10.5}
sp_groups_not <- filter(zooplankton_sog_axis_names,
                        !in_perry_fig_s1)$species_group_name
par(mfrow = c(3, 2))
for(i in 1:5){
  plot(zooplankton_sog,
       species_group = sp_groups_not[i],
       lwd = lwd_val)
}
```
These also show recent positive anomalies, except for Scyphozoa.

## Technial notes

Note that the `zooplankton_sog` object has class `pacea_zooplankton`:
```{r hakebiomassclass}
class(hake_biomass)
```
such that the above command `plot(zooplankton_sog)` first looks to use our tailored function
`plot.pacea_zooplankton()`, which in turn calls `plot.pacea_index()`, to give
the style of plot shown above. The `lwd` argument may need to be tailored to
control the widths of the bars, as in the final plots here, depending on how big your plot is (see
`?plot.pacea_zooplankton`).

Also note how the y-axis labels are automatically created with sensible text, including
the anomaly units. Users should not need to worry too much, but this
is done internally by `plot.pacea_zooplankton()`; see example in
`?zooplankton_sog_axis_names` to adapt one of the existing ones, or just define
your own and specify it as the usual `ylab` argument.
