---
title: "Examples of querying GLORYS data"
author: "Andrew Edwards, Eric Ward, and Megan Feddern"
output:
  pdf_document:
    keep_tex: true
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "glorys-examples-cache/",
  fig.path = "glorys-examples-figs-cache/",
  fig.width = 7,   # Good for sizeSpectraFit hist and spectra plots, prob want to adapt
  fig.height = 9
)
# comment = "#>",

knitr::dep_auto()
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("glorys_examples_for_andy_isobathprocessing-ame.Rmd")

```

Andy adapting from code from Eric and Megan. See comments for more details.

<!-- Start is from Eric, then Megan Feddern added the Section 'Post processing 
summarising into time series' onwards. 
Andy editing here, and making .Rmd instead of .qmd. Also doing the CopernicusMarine 
approach, so skipping the first two options here-->

```{r, libraries}
library(dplyr)
library(tidyr)
library(tidync)
library(lubridate)
library(tidync)
library(tidyverse)   # might not be fully needed?
library(data.table)
library(CopernicusMarine)
library(abind)
library(stars)
# library(fuzzyjoin)

```

## Downloading – straight Python

This is the approach I'm using the most right now – querying temp file downloads straight from Python. You have to install via <https://pypi.org/project/copernicusmarine/> – I like this because I think it's made by the Copernicus development team and seems very stable.

```{python echo=FALSE, eval = FALSE}
# example of 
copernicusmarine.subset(
  dataset_id="cmems_mod_glo_phy_my_0.083deg_P1M-m",
  variables=["thetao"],  # temperature
  minimum_longitude=-126,
  maximum_longitude=-115,
  minimum_latitude=32,
  maximum_latitude=50,
  start_datetime="1994-01-01T00:00:00",
  end_datetime="2024-08-20T23:59:59",
  output_directory="./copernicus-data",  # Directory to save 
  output_filename="glorys_data_01.nc"
)

```

## Downloading – reticulate

As an alternative to pure Python, you can use the reticulate package in R.

```{r, eval = FALSE}
library(reticulate)
#install_python() 
virtualenv_create(envname = "thiscopernicus")
#virtualenv_install("thiscopernicus", packages = c("copernicusmarine"))
use_virtualenv("thiscopernicus", required = TRUE)
#py_install("copernicusmarine")

cm <- import("copernicusmarine")

cm$subset(
    dataset_id = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    start_datetime=paste0(1994,"-01-01T00:00:00"),
    end_datetime=paste0(2024,"-12-31T23:59:59"),
    variables = list("thetao"),
    minimum_longitude = -126,
    maximum_longitude = -115,
    minimum_latitude = 32,
    maximum_latitude = 50,
    minimum_depth = 50,
    maximum_depth = 250,
    output_filename = "CMEMS_tmp.nc"
  )
```

## Downloading – CopernicusMarine

This example works with the R package CopernicusMarine. You need to authenticate credentials the first time you use it, and I've avoided it because it was a little glitchy during the last Copernicus upgrade

Skip to the one lower down; this is an example, but want to duplicate Megan's
MLD analysis below.
```{r, eval = FALSE}
# library(CopernicusMarine)
my_data <- cms_download_subset(
  product = "GLOBAL_MULTIYEAR_PHY_001_030",
  layer = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
  variable = "thetao",
  region = c(-126, 32, -115, 50), # lon_min, lat_min, lon_max, lat_max
  timerange = c("1994-01-01", "2024-08-20"),
  verticalrange = c(-50, -250), # 50m to 250m depth
  progress = TRUE
)
my_data
```

## Processing – netCDF

For most of these, you can use `tidync` in R to process the data

```{r, eval = FALSE}
# Using R approach we have no .nc file saved it seems, but my_data is a start
# object and has what we need I think. Kelsey suggested just save stars object
# as a netcdf so can easily use this code
temp <- tidync::tidync("CMEMS_tmp.nc") |>
  tidync::hyper_tibble( force = TRUE) |>
  drop_na() |>
  group_by(longitude,latitude,time)
```

## post processing – summarising into time series


Example time series: MLDyolk To get the appropriate time series, we need to summarise the dataset. We can easily just download the data at the depths and latitude we need but to get this time series as "between the 100 and 2,000 m isobath" as defined in Vestfals et al. 2023 and to do that we need to combine our .nc tibble with a bathymetric dataset. To approach this I am going to download beyond the longitude I need then combine datasets. You can download the bathymetry data from GLORYS and then use the grid to assign longitude. Here I am just going to use the dataset as a .csv that Nick and I are already using. No need to reinvent the wheel...

First lets download the mixed layer data, I usually use reticulate so I will just use Eric's approach from above:

Andy then doing it using the R package approach.

```{r, eval = FALSE}
#
cm <- import("copernicusmarine")

cm$subset(
    dataset_id = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    start_datetime=paste0(1994,"-01-01T00:00:00"),
    end_datetime=paste0(2024,"-12-31T23:59:59"),
    variables = list("mlotst"),
    minimum_longitude = -130, #probably more longitude than we need
    maximum_longitude = -115,
    minimum_latitude = 31,
    maximum_latitude = 36,#no depth because MLD is not a variable with depths
    output_filename = "CMEMS_mlotst.nc"
  )
```

R package approach, copying Eric's extraction from above and using Megan's options: 
```{r, mixedlayer}
lon_min <- -130
lon_max <- -115
lat_min <-  31
lat_max <- 36

if(FALSE){    # set to TRUE to redownload. Have not rerun since doing lon_min
    # etc. TODO
  mld_via_package <- cms_download_subset(
    product = "GLOBAL_MULTIYEAR_PHY_001_030",
    layer = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    variable = "mlotst",
    region = c(lon_min, lat_min, lon_max, lat_max),
    timerange = c("1994-01-01", "2024-12-31"),
    # verticalrange   no depth since MLD; if missing then gets given NULL
    progress = TRUE
  )
  mld_via_package <- abind::adrop(mld_via_package) # drop elevation as only one value
  saveRDS(mld_via_package,
          "mld_via_package.Rds")
} else
{
  mld_via_package <- readRDS("mld_via_package.Rds")
}


mld_via_package

summary(mld_via_package)
# plot(mld_via_package)  # not doing as plots every month
```
So a stars object.

Summarise just as Eric did (not doing)

```{r, eval = FALSE}
  temp <- tidync::tidync("CMEMS_mlotst.nc") |>
    hyper_tibble( force = TRUE) |>
    drop_na() |>
    group_by(longitude,latitude,time)
```

Copying in what we have been using for this - this is a static variable from
GLORYS and linked to lat/long using the grid cell info. Andy found the lats and
lons to not be exactly the same as those in the downloaded MLD data though (see below).

```{r, bath}
bath_orig <- read.csv('databathymetry-m.csv') %>%
  as_tibble()
bath_orig

bath <- bath_orig %>%
    filter(longitude >= lon_min,
           longitude <= lon_max,
           latitude >= lat_min,
           latitude <= lat_max)
bath
```

Now wrangle the MLD values,  to
hopefully get agreement with Megan's:
```{r, mldanalysis}
dim(mld_via_package)
prod(dim(mld_via_package))

my_tibble_mld <- as_tibble(mld_via_package)    # makes one big tibble, uses that
                                               # class(mld_via_package) is stars
my_tibble_mld

expect_equal(prod(dim(mld_via_package)),
             nrow(my_tibble_mld))   

# Megan's original approach is below, bits here are taken from it, maybe
# different order
# tibble is huge, so filter to just required months for now
my_tibble_mld_red <- my_tibble_mld %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 1:4) # getting the months for the yolk-sac lifestage

# Could do the averaging over the four months first, assuming that the long and
# lat are consistent. No - since need the keep the correct bathymetries first to
# include those, and then do the averaging.

my_tibble_mld_red
summary(my_tibble_mld_red)

# do a small subset first, just 1994

my_tibble_mld_red_1994 <- filter(my_tibble_mld_red,
                                     year == 1994)
my_tibble_mld_red_1994
```
Try rounding lat and long first as `left_join` without doing that had
`latitude.x` and `latitude.y` I think because of slight numerical differences.

```{r, rounding}
range(diff(my_tibble_mld_red_1994$longitude))
range(diff(my_tibble_mld_red_1994$latitude))

# 0.08333 is 1/12 of a degree. So any approximation should be fine to 0.00001
dp <- 4    # decimal places to round to

my_tibble_mld_red_1994_round <- my_tibble_mld_red_1994 %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
my_tibble_mld_red_1994_round

range(diff(my_tibble_mld_red_1994_round$longitude))
range(diff(my_tibble_mld_red_1994_round$latitude))


# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

bath_round <- bath %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
bath_round
range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, skipping as maybe makes .tex too big
```{r, grids}
plot(my_tibble_mld_red_1994_round$longitude, 
     my_tibble_mld_red_1994_round$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(my_tibble_mld_red_1994_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(my_tibble_mld_red_1994_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       cex = 0.5)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combine}
# Original way round that Megan did, then analysing further down 
# Now join all of them, still for just 1994:
combined <- bath_round %>%
  left_join(my_tibble_mld_red_1994_round)
combined[500:520, ]
summary(combined)
# Looks like 225 NA's, which are presumably the bathymetry with no MLD.

# Try the other way round, I feel we only care about values where have MLD
# I guess need to check if every lat-lon for MLD gets a bottom_depth
combined_2 <- my_tibble_mld_red_1994_round %>%
  left_join(bath_round)
combined_2
summary(combined_2)
# lots of NA's
```

Aha, Megan's way does make more sense, as not trying to assign as MLD grid
includes lots of land (since a rectangle), so end up with lots more rows with
NA's in either MLD or bathymetry when doing the second way. Will want the
efficiency when doing all years.

Now need just the required isobaths:
```{r, combine2}
summary(bath_round)
summary(my_tibble_mld_red_1994_round)

# isobath greater than 100m less than 2,000
combined_need <- combined %>%
  filter(bottom_depth >= 100 & bottom_depth <= 2000)
summary(combined_need)

combined_need[is.na(combined_need$mlotst),]
which(is.na(combined_need$mlotst))
# So they are kind of spread out. Only a few values, so just remove those rows.

combined_need_no_na <- combined_need %>%
  filter(!is.na(mlotst))
combined_need_no_na
summary(combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in  MLD column

mld_yolk_tib <- combined_need_no_na %>%
  group_by(year) %>%
  summarise(mld_yolk = as.numeric(mean(mlotst)))  #woo! we have it
            # need as.numeric else class(pull(mld_yolk_tib)) is "units"

mld_yolk_tib
 

# 1994 gives 31.4, but original from last year is 22.65 (from below).
```

HERE, adapting Megan's comparison code from below, need to rerun above on all
years to then use properly. Saving as commit after 9f38fb2c and then doing above
without setting to just 1994.
```{r, compare, eval = FALSE}
res_2025 <- read.csv('DATA_Combined_glorys_hake_UNSTANDARDIZED.csv')%>%
  select(year, MLDyolk)%>%
  rename(MLDyolk_Old = MLDyolk)

comparison_wide <- old %>%
  left_join(mld_yolk_tib)

comparison <- comparison_wide %>%
  pivot_longer(-year)

ggplot(data=comparison, aes(y=value, x=year, col=name))+
  geom_line()+
  theme_bw() #looks great! These minor differences are from the reanalysis

cor(comparison_wide%>%filter(year>1993&year<2024)%>%select(MLDyolk_Old), 
    comparison_wide%>%filter(year>1993&year<2024)%>%select(MLD_yolk_New)) #0.99 love to see it
```



```{r, meganorig, eval = FALSE}
# Megan's orginal:
combined <- bath%>%left_join(temp)%>% #combining data so we have bottom depth to filter over for isobath
  filter(bottom_depth >= 100 & bottom_depth <= 2000)%>% #isobath greater than 100m less than 2,000
  mutate(date = as_date(time/24,origin="1950-01-01 00:00:00"))%>% #adding in date
  mutate(year=year(date), month = month(date))%>% #getting out year and month from date
  filter(month ==1|month==2|month==3|month==4 )%>% #getting the months for the yolk-sac lifestage
  group_by(year)%>%
  summarise(MLD_yolk_New=mean(mlotst)) #woo! we have it



```

Megan: I am just going to double check this against the old dataset that I have for hake to make sure we didnt mess anything up...you could swap out the file for whatever you have from last years assessment (I think that is what I have...)

```{r, eval = FALSE}
old <- read.csv('DATA_Combined_glorys_hake_UNSTANDARDIZED.csv')%>%
  select(year, MLDyolk)%>%
  rename(MLDyolk_Old=MLDyolk)

comparison_wide <- old%>%left_join(combined)
comparison<-  comparison_wide%>%pivot_longer(-year)

ggplot(data=comparison, aes(y=value, x=year, col=name))+
  geom_line()+
  theme_bw() #looks great! These minor differences are from the reanalysis

cor(comparison_wide%>%filter(year>1993&year<2024)%>%select(MLDyolk_Old), 
    comparison_wide%>%filter(year>1993&year<2024)%>%select(MLD_yolk_New)) #0.99 love to see it

```

citation - see
https://pepijn-devries.github.io/CopernicusMarine/
