---
title: "Extracting and analysing GLORYS time series for hake assessment"
author: "Andrew Edwards, Eric Ward, and Megan Feddern"
output:
  pdf_document:
    keep_tex: true
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "glorys-examples-cache/",
  fig.path = "glorys-examples-figs-cache/",
  fig.width = 7,
  fig.height = 5
)
# comment = "#>",

knitr::dep_auto()
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("glorys_2026.Rmd")
```

```{r, libraries}
library(dplyr)
library(tidyr)
library(tidync)
library(lubridate)
library(tidync)
library(tidyverse)   # might not be fully needed?
library(data.table)
library(CopernicusMarine)
library(abind)
library(stars)
ifelse(Sys.info()["user"] == "edwardsand",
       load_all(),
       library(pacea))   # remotes::install_github("pbs-assess/pacea")
```

TODO SSHjac off California needs correct spatial area

TODO ASTeggs needs transport definition

TODO PUlatelarv neds strength of undercurrent definition

Obtaining time series from GLORYS for the 2026 Pacific Hake assessment
Ecoysystem Report. Getting the GLORYS-related variables as listed in Table H.1
of the 2025 hake assessment.

Simplifying down the needed code from
`glorys_examples_for_andy_isobathprocessing-ame.Rmd`, which was testing the R
workflow and comparing the `mld_yolk` values to what we had in 2025
assessment. That `.Rmd` was adapted from code from Eric and Megan.

Using the bathymetry from Megan that Nick and her had used last year, to save
having to redownload from GLORYS right now. 

See
https://data.marine.copernicus.eu/product/GLOBAL_MULTIYEAR_PHY_001_030/services
for description of the dataset `cmems_mod_glo_phy_my_0.083deg_P1M-m` being downloaded for MLDyolk (and likely
more). It's daily; note there is a monthly one available also (stick with daily
for now as already have the wrangling code done). 

https://documentation.marine.copernicus.eu/PUM/CMEMS-GLO-PUM-001-030.pdf for
description of variables.

For citation - see
https://pepijn-devries.github.io/CopernicusMarine/


\begin{table}[hbp]
\centering
\caption{Variables from Table H.1 of 2025 hake assessment for which GLORYS was
used, and is being used here for updating. 
Shelf break is considered
between 100 and 2,000~m isobaths, defining the longitudinal extent for several
variables. The first subscripts refer to life-history stage
(`pre` is preconditioning, while JA is Jan-Apr, in the egg to late larvae stage).}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{p{3cm}p{6cm}p{6cm}}
\hline % \toprule
Abbreviation~of variable & Definition & Effect on recruitment; expectation for an increase in
variable\\
\hline % \midrule

TEMP$_{\mbox{spawn}}$ & Mean temperature during spawning (shelf break,
Jan-Mar, 130-500~m depth, 31-36°~N) &
At higher temperatures, fish are less likely to spawn but growth rate of larvae
increases;
$\cup$-shaped relationship with recruitment that year\\

AST$_{\mbox{eggs}}$ & Net along-shore transport (shelf break,
Jan-Mar, 40-60~m depth, 31-36°~N)
& Increased northward advection away from juvenile nursery areas decreases
recruitment; $\downarrow$ recruitment that year\\

MLD$_{\mbox{yolk}}$ & Mean mixed layer depth (shelf break
Jan-Apr, 31-36°~N)
& Larvae aggregate at base of mixed layer so mixed layer depth may limit how far
they rise in the water column affecting later transport; $\downarrow$ recruitment that year\\

SSH$_{\mbox{JA.c}}$ & Average sea-surface
height off California as an indicator of basin-scale processes (from coast to
30~km offshore, Jan-Apr, 34.5-42.5°~N) &
Higher sea surface is indicative of higher productivity and better conditions for copepods;
$\uparrow$ recruitment that year\\

MLD$_{\mbox{latelarv}}$ & Mean mixed layer depth (shelf break
Mar-Jun, 31-37°~N)
& Larvae aggregate at base of mixed layer so mixed layer depth may limit how far they rise in the
water column affecting later transport, and possibly less competition and
predation when mixed layer shallower; % Kristin's guess
$\uparrow$ recruitment that year\\

PU$_{\mbox{latelarv}}$ & Strength of poleward undercurrent (from coast to 275~m isobath,
Mar-Jun, 75-275~m depth, 33.5-34.5°~N)
& Increased northward
advection away from juvenile nursery areas decreases recruitment; $\downarrow$ recruitment that year \\

\hline % \bottomrule
\end{tabular}
\end{table}

Looking at the GLORYS .pdf mentioned above for the specific product, plus the above table, seems like we
need:

TEMP$_{\mbox{spawn}}$ -- potential temperature `thetao` (manual also says
`sea_water_potential_temperature` but think just `thetao` is needed).

AST$_{\mbox{eggs}}$ -- haved asked Kristin how she calculated it last
year. Maybe some
rough trigonometry from `uo` and `vo`, respectively the eastward and northward
ocean current velocity. Or just northward, given hypothesis above. Try that first.

MLD$_{\mbox{yolk}}$ and MLD$_{\mbox{latelarv}}$-- mixed layer depth `mlotst` as
used by Megan

SSH$_{\mbox{JA.c}}$ -- sea-surface height `zos` 

PU$_{\mbox{latelarv}}$ -- given description also talks about northward
advection, just do `vo`.

So I think we can download some of these together. Presumably do the 2-d ones in
one go since do not have to specify a depth range. Then the ones for which need
a depth range (`verticalrange`); have to see whether the results come averaged
(i.e. a 2-d field plus time) or still 3-d. I'm guessing the former, which means
will have to do each one spearaetely probably as the depths are different.

1. No depth `verticalrange` needed:

MLD$_{\mbox{yolk}}$, MLD$_{\mbox{latelarv}}$ need `mlotst`, and
SSH$_{\mbox{JA.c}}$ needs `zos`. 
Lat: 31-42.5

MLD wants shelf break (100-2000m). SSH is coast to 30 km offshore, for which
think I can use Wei's masking provided it covers the same latitude. 

2. With depth values, do each separately.


Have also asked Kristin:

2. 0-30km offshore calculations, and 

3. Looks like Wei got the EKE values from Copernicus for me. Wondering why you
did not get those last year (no idea if they were available then)? You had
mentioned that the values can be different from different products.


# Download variables with no `verticalrange`

R package approach, copying Eric's extraction from above and using Megan's options: 
```{r, nodepth}
time_range <- c("1993-01-01", "2024-12-31")   # Though it gives 2025-01-01
# values also

lon_min <- -130
lon_max <- -115
lat_min <-  31
lat_max <- 42.5

if(FALSE){    # set to TRUE to redownload
  data_no_depth <- cms_download_subset(
    product = "GLOBAL_MULTIYEAR_PHY_001_030",
    layer = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    variable = c("mlotst", "zos"),
    region = c(lon_min, lat_min, lon_max, lat_max),
    timerange = time_range,
    # verticalrange   no depth since MLD; if missing then gets given NULL
    progress = TRUE
  )
  data_no_depth <- abind::adrop(data_no_depth) # drop elevation as only one value
  saveRDS(data_no_depth,
          "data_no_depth.Rds")
} else {
  data_no_depth <- readRDS("data_no_depth.Rds")
}

data_no_depth

summary(data_no_depth)
# plot(data_no_depth)  # not doing as plots every month
```
So a stars object.

Copying in Meghan's bathymetry - ``this is a static variable from
GLORYS and linked to lat/long using the grid cell info''. Think I can easily redownload. Andy found the lats and
lons to not be exactly the same as those in the downloaded MLD data though (see
below), maybe they would be if using the same approach

```{r, bath}
bath_orig <- read.csv('databathymetry-m.csv') %>%
  as_tibble()
bath_orig

bath <- bath_orig %>%
    filter(longitude >= lon_min,
           longitude <= lon_max,
           latitude >= lat_min,
           latitude <= lat_max)
bath
```

Actually, may as well get bathymetry directly, didn't work so not run or show here:
```{r, bathdirect, eval = FALSE, echo = FALSE}
if(FALSE){    # set to TRUE to redownload
  bath_direct <- cms_download_subset(
    product = "GLOBAL_MULTIYEAR_PHY_001_030",
    layer = "cmems_mod_glo_phy_my_0.083deg_static",
    variable = c("mask", "deptho"),
    region = c(lon_min, lat_min, lon_max, lat_max),
    timerange = NULL,
    # verticalrange   no depth since MLD; if missing then gets given NULL
    progress = TRUE
  )
  # bath_direct <- abind::adrop(bath_direct) # drop elevation as only one value
  saveRDS(bath_direct,
          "bath_direct.Rds")
} else {
  bath_direct <- readRDS("bath_direct.Rds")
}

# That does not work, trying to figure out:
tt <- cms_products_list()
tttt <- filter(tt, product_id == "GLOBAL_MULTIYEAR_PHY_001_030")  # tbl_df but I
# think is more like a list, some elements have >1 value maybe
 <- as.vector(tttt)$mainVariables   # Does not have bathymetry, even though pdf help
# says batho and mask.
```

Now wrangle the MLD values, actually this contains `zos` also, so can use for SSH.
```{r, mldanalysis}
dim(data_no_depth)
prod(dim(data_no_depth))

mld_tib <- as_tibble(data_no_depth)    # makes one big tibble, uses that
                                       # class(data_no_depth) is stars
mld_tib

expect_equal(prod(dim(data_no_depth)),
             nrow(mld_tib))   
```

## MLDyolk

Do each covariate in turn, MLDyolk first as already have code for it. Could
maybe avoid repeating some things (e.g. join bathymetry to mld earlier than we
do), but may well take longer as not restricting datasets to what we actually
need each time.

```{r, mldyolk}
mld_yolk_data <- mld_tib %>%
  select(-c("zos")) %>%  
  filter(latitude >= 31,
         latitude <= 36) %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 1:4) 
mld_yolk_data

summary(mld_yolk_data)

summary(as.factor(mld_yolk_data$month))

# Could do the averaging over the four months first, assuming that the long and
# lat are consistent. No - since need the keep the correct bathymetries first to
# include those, and then do the averaging.
```

Need to round before joining with bathymetry, since slight numerical
differences. See previous .Rmd for details.

```{r, roundingmldyolk}
dp <- 4    # decimal places to round to

mld_yolk_data_round <- mld_yolk_data %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
mld_yolk_data_round

range(diff(mld_yolk_data_round$longitude))
range(diff(mld_yolk_data_round$latitude))

# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

bath_round <- bath %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
bath_round
range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, though just doing 1994 as presume the locations are the same
for each year:
```{r, gridsmldyolk}
plot(filter(mld_yolk_data_round, year == 1993, month == 1)$longitude, 
     filter(mld_yolk_data_round, year == 1993, month == 1)$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(mld_yolk_data_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(mld_yolk_data_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       pch = 20,
       cex = 0.75)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combinemldyolk}
mld_yolk_combined <- bath_round %>%
  left_join(mld_yolk_data_round)
mld_yolk_combined
mld_yolk_combined[500:520, ]
summary(mld_yolk_combined)
# See previous .Rmd for details on NAs;  mlotst ones looked to around the Channel
# Islands, around -120, 34.
# Megan did the left_join that way. I also tried the other way round, but end up
# with way more NAs; see previous .Rmd.
```

Now need just the required isobaths:
```{r, combine2mldyolk}
# isobath greater than 100m less than 2,000 for MLDyolk
mld_yolk_combined_need <- mld_yolk_combined %>%
  filter(bottom_depth >= 100 & bottom_depth <= 2000)
mld_yolk_combined_need
summary(mld_yolk_combined_need)

# mld_yolk_combined_need[is.na(mld_yolk_combined_need$mlotst),]
```

Plot the locations of NAs for MLD:
```{r, plotmissingmldyolk}
plot(mld_yolk_combined_need$longitude[is.na(mld_yolk_combined_need$mlotst)],
     mld_yolk_combined_need$latitude[is.na(mld_yolk_combined_need$mlotst)])
```
as expected.

Okay, remove those rows:

```{r, removeNAsmldyolk}
mld_yolk_combined_need_no_na <- mld_yolk_combined_need %>%
  filter(!is.na(mlotst))
mld_yolk_combined_need_no_na
summary(mld_yolk_combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in  MLD column
max(mld_yolk_combined_need_no_na$year)

filter(mld_yolk_combined_need_no_na, year ==
        max(mld_yolk_combined_need_no_na$year)) %>%
  summary()
```

That shows only values for January in the final year, so remove it.

```{r, removefinalyearmldyolk}
mld_yolk_tib <- mld_yolk_combined_need_no_na %>%
  filter(year < max(mld_yolk_combined_need_no_na$year)) %>%
  group_by(year) %>%
  summarise(mld_yolk = as.numeric(mean(mlotst)))  #woo! we have it
            # need as.numeric else class(pull(mld_yolk_tib)) is "units"

mld_yolk_tib
mld_yolk_tib %>% tail()
```

Compare to last year's values that went into the assessment (did Megan's check
on MLDyolk in previous .Rmd). Copy some code over from
`ecosystem-summary-hake-glorys.R`, and standardise like we did (though I think
Kristin supplied values already standardised):
```{r, redo2025}
glorys_2025 <- read.csv("../../ecosystem-summary-hake/DATA_Combined_glorys_hake_UW_for_Andy.csv") %>%
    tibble::as_tibble()

summary(glorys_2025)
```

Want each covariate as it's own index object for plotting. 

Note there will be an extra year, so
standardising with that can result in slight differences. Will check, and also 
check that the new figures look similar enough. Maybe just doing the latter for brevity.

```{r, mldyolkindex}
mld_yolk_index_2025 <- dplyr::select(glorys_2025,
                                     year,
                                     value = "MLDyolk") %>%
  dplyr::mutate(anomaly = standardise(value))
class(mld_yolk_index_2025) <- class(oni)

mld_yolk_index_2025   # As used in 2025 assessment
summary(mld_yolk_index_2025)

# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years) 
mld_yolk_index_2026_restr <- mld_yolk_tib %>%
 dplyr::filter(year <= max(mld_yolk_index_2025$year)) %>%
 dplyr::mutate(anomaly = standardise(mld_yolk))
class(mld_yolk_index_2026_restr) <- class(oni)

mld_yolk_index_2026_restr   # As calculated here
summary(mld_yolk_index_2026_restr)

# And calculate the new one for all available years
mld_yolk_index_2026 <- mld_yolk_tib %>%
 dplyr::mutate(anomaly = standardise(mld_yolk))
class(mld_yolk_index_2026) <- class(oni)

mld_yolk_index_2026   # As calculated here. 
summary(mld_yolk_index_2026)
```

Compare the updated and 2025-assessment values:
```{r, compareplot3mldyolk}
plot(mld_yolk_index_2025$year,
     mld_yolk_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(mld_yolk_index_2026_restr$year,
     mld_yolk_index_2026_restr$anomaly,
     type = "o",
     col = "red")
```

Now as a correlation plot
```{r, compareplot4mldyolk}
plot(mld_yolk_index_2025$anomaly,
     mld_yolk_index_2026_restr$anomaly)
abline(a = 0, b = 1)

cor(mld_yolk_index_2025$anomaly,
    mld_yolk_index_2026_restr$anomaly)   # 0.997 like Megan found

cor(mld_yolk_index_2025$value,
    mld_yolk_index_2026_restr$mld_yolk)   # 0.997 like Megan found, cor is the
# same as for anomaly, which is probably obvious 
```

New updated index is
```{r, mldyolkindex2}
plot(mld_yolk_index_2026)
```

Will make the full figure later.


## MLDlatelarv

MLD$_{\mbox{latelarv}}$, using values from the above table.

```{r, mldlatelarv}
mld_late_larv_data <- mld_tib %>%
  filter(latitude >= 31,
         latitude <= 37) %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 3:6) 
mld_late_larv_data

summary(mld_late_larv_data)

summary(as.factor(mld_late_larv_data$month)) # Surely there should be less April
# and June than the others? Slightly confusing. TODO think more.
```

Need to round before joining with bathymetry, since slight numerical
differences. See previous .Rmd for details.

```{r, roundingmldlatelarv}
mld_late_larv_data_round <- mld_late_larv_data %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
mld_late_larv_data_round

range(diff(mld_late_larv_data_round$longitude))
range(diff(mld_late_larv_data_round$latitude))

# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, though just doing 1993 as presume the locations are the same
for each year, shows up well on full screen (not in pdf):
```{r, gridsmldlatelarv}
plot(filter(mld_late_larv_data_round, year == 1993, month == 4)$longitude, 
     filter(mld_late_larv_data_round, year == 1993, month == 4)$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(mld_late_larv_data_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(mld_late_larv_data_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       pch = 20,
       cex = 0.75)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combinemldlatelarv}
mld_late_larv_combined <- bath_round %>%
  left_join(mld_late_larv_data_round)
mld_late_larv_combined
mld_late_larv_combined[500:520, ]
summary(mld_late_larv_combined)
# See previous .Rmd for details on NAs;  mlotst ones looked to around the Channel
# Islands, around -120, 34.
# Megan did the left_join that way. I also tried the other way round, but end up
# with way more NAs; see previous .Rmd.
```

Now need just the required isobaths:
```{r, combine2mldlatelarv}
# isobath greater than 100m less than 2,000 for MLD_latelarv
mld_late_larv_combined_need <- mld_late_larv_combined %>%
  filter(bottom_depth >= 100 & bottom_depth <= 2000)
mld_late_larv_combined_need
summary(mld_late_larv_combined_need)

# mld_late_larv_combined_need[is.na(mld_late_larv_combined_need$mlotst),]
```

Plot the locations of NAs for MLD:
```{r, plotmissingmldlatelarv}
plot(mld_late_larv_combined_need$longitude[is.na(mld_late_larv_combined_need$mlotst)],
     mld_late_larv_combined_need$latitude[is.na(mld_late_larv_combined_need$mlotst)])
```
as expected.

Okay, remove those rows:

```{r, removeNAsmldlatelarv}
mld_late_larv_combined_need_no_na <- mld_late_larv_combined_need %>%
  filter(!is.na(mlotst))
mld_late_larv_combined_need_no_na
summary(mld_late_larv_combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in  MLD column
max(mld_late_larv_combined_need_no_na$year)

filter(mld_late_larv_combined_need_no_na, year ==
        max(mld_late_larv_combined_need_no_na$year)) %>%
  summary()
```

No values here for final year, so second line next not needed here:

```{r, removefinalyearmldlatelarv}
mld_late_larv_tib <- mld_late_larv_combined_need_no_na %>%
  # filter(year < max(mld_late_larv_combined_need_no_na$year)) %>%
  group_by(year) %>%
  summarise(mld_late_larv = as.numeric(mean(mlotst)))  #woo! we have it
            # need as.numeric else class(pull(mld_late_larv_tib)) is "units"

mld_late_larv_tib
mld_late_larv_tib %>% tail()
```

Compare to last year's values that went into the assessment (did Megan's check
on MLDyolk in previous .Rmd).

Note there will be an extra year, so
standardising with that can result in slight differences. Will check, and also 
check that the new figures look similar enough. Maybe just doing the latter for brevity.

```{r, mldlatelarvindex}
mld_late_larv_index_2025 <- dplyr::select(glorys_2025,
                                          year,
                                          value = "MLDlate") %>%
    dplyr::mutate(anomaly = standardise(value))
class(mld_late_larv_index_2025) <- class(oni)

mld_late_larv_index_2025   # As used in 2025 assessment
summary(mld_late_larv_index_2025)

# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years) 
mld_late_larv_index_2026_restr <- mld_late_larv_tib %>%
 dplyr::filter(year <= max(mld_late_larv_index_2025$year)) %>%
 dplyr::mutate(anomaly = standardise(mld_late_larv))
class(mld_late_larv_index_2026_restr) <- class(oni)

mld_late_larv_index_2026_restr   # As calculated here
summary(mld_late_larv_index_2026_restr)

# And calculate the new one for all available years
mld_late_larv_index_2026 <- mld_late_larv_tib %>%
 dplyr::mutate(anomaly = standardise(mld_late_larv))
class(mld_late_larv_index_2026) <- class(oni)

mld_late_larv_index_2026   # As calculated here. 
summary(mld_late_larv_index_2026)
```

Compare the updated and 2025-assessment values:
```{r, compareplot3mldlatelarv}
plot(mld_late_larv_index_2025$year,
     mld_late_larv_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(mld_late_larv_index_2026_restr$year,
     mld_late_larv_index_2026_restr$anomaly,
     type = "o",
     col = "red")
```

Now as a correlation plot
```{r, compareplot4mldlatelarv}
plot(mld_late_larv_index_2025$anomaly,
     mld_late_larv_index_2026_restr$anomaly)
abline(a = 0, b = 1)

cor(mld_late_larv_index_2025$anomaly,
    mld_late_larv_index_2026_restr$anomaly)
```
Check above value is high.

New updated index is
```{r, mldlatelarvindex2}
plot(mld_late_larv_index_2026)
```

Will make the full figure later.

## SSHJAc

SSH$_{\mbox{JA.c}}$ needs `zos`, sea-surface height. Also need differenct latitude and spatial domain.

TODO: need to figure out 0-30km part. THESE RESULTS WILL CURRENTLY ASSUME THE
SAME spatial area as for the two MLD ones above. `mld_tib` still has `zos` in.

```{r, sshjac}
ssh_jac_data <- mld_tib %>%
  select(-c("mlotst")) %>%      
  filter(latitude >= 34.5,
         latitude <= 42.5) %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 1:4) 
ssh_jac_data

summary(ssh_jac_data)
tail(ssh_jac_data)
summary(as.factor(ssh_jac_data$month)) 
```

Need to round before joining with bathymetry, since slight numerical
differences. See previous .Rmd for details.

```{r, roundingsshjac}
ssh_jac_data_round <- ssh_jac_data %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
ssh_jac_data_round

range(diff(ssh_jac_data_round$longitude))
range(diff(ssh_jac_data_round$latitude))

# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, though just doing 1993 as presume the locations are the same
for each year, shows up well on full screen (not in pdf):
```{r, gridssshjac}
plot(filter(ssh_jac_data_round, year == 1993, month == 4)$longitude, 
     filter(ssh_jac_data_round, year == 1993, month == 4)$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(ssh_jac_data_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(ssh_jac_data_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       pch = 20,
       cex = 0.75)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combinesshjac}
ssh_jac_combined <- bath_round %>%
  left_join(ssh_jac_data_round)
ssh_jac_combined
ssh_jac_combined[500:520, ]
summary(ssh_jac_combined)
# See previous .Rmd for details on NAs;  mlotst ones looked to around the Channel
# Islands, around -120, 34.
# Megan did the left_join that way. I also tried the other way round, but end up
# with way more NAs; see previous .Rmd.
```

Now need just the required isobaths:  TODO need something else here for SSHjac, so
commenting out
```{r, combine2sshjac}
# isobath greater than 100m less than 2,000 for MLD_latelarv
ssh_jac_combined_need <- ssh_jac_combined

# %>%
#  filter(bottom_depth >= 100 & bottom_depth <= 2000)
#ssh_jac_combined_need
#summary(ssh_jac_combined_need)

# ssh_jac_combined_need[is.na(ssh_jac_combined_need$zos),]
```

Plot the locations of NAs for zos:
```{r, plotmissingsshjac}
plot(ssh_jac_combined_need$longitude[is.na(ssh_jac_combined_need$zos)],
     ssh_jac_combined_need$latitude[is.na(ssh_jac_combined_need$zos)])
```
as expected.

Okay, remove those rows:

```{r, removeNAsshjac}
ssh_jac_combined_need_no_na <- ssh_jac_combined_need %>%
  filter(!is.na(zos))
ssh_jac_combined_need_no_na
summary(ssh_jac_combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in  MLD column
max(ssh_jac_combined_need_no_na$year)

filter(ssh_jac_combined_need_no_na, year ==
        max(ssh_jac_combined_need_no_na$year)) %>%
  summary()
```

No non-January values for final year, so remove those:

```{r, removefinalyearsshjac}
ssh_jac_tib <- ssh_jac_combined_need_no_na %>%
  filter(year < max(ssh_jac_combined_need_no_na$year)) %>%
  group_by(year) %>%
  summarise(ssh_jac = as.numeric(mean(zos)))  #woo! we have it
            # need as.numeric else class(pull(ssh_jac_tib)) is "units"

ssh_jac_tib
ssh_jac_tib %>% tail()
```

Compare to last year's values that went into the assessment (did Megan's check
on MLDyolk in previous .Rmd).

Note there will be an extra year, so
standardising with that can result in slight differences. Will check, and also 
check that the new figures look similar enough. Maybe just doing the latter for brevity.

```{r, sshjacindex}
ssh_jac_index_2025 <- dplyr::select(glorys_2025,
                                          year,
                                          value = "SSHjac") %>%
    dplyr::mutate(anomaly = standardise(value))
class(ssh_jac_index_2025) <- class(oni)

ssh_jac_index_2025   # As used in 2025 assessment
summary(ssh_jac_index_2025)

# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years) 
ssh_jac_index_2026_restr <- ssh_jac_tib %>%
 dplyr::filter(year <= max(ssh_jac_index_2025$year)) %>%
 dplyr::mutate(anomaly = standardise(ssh_jac))
class(ssh_jac_index_2026_restr) <- class(oni)

ssh_jac_index_2026_restr   # As calculated here
summary(ssh_jac_index_2026_restr)

# And calculate the new one for all available years
ssh_jac_index_2026 <- ssh_jac_tib %>%
 dplyr::mutate(anomaly = standardise(ssh_jac))
class(ssh_jac_index_2026) <- class(oni)

ssh_jac_index_2026   # As calculated here. 
summary(ssh_jac_index_2026)
```

TODO These correctly do not look good yet as have not done the correct spatial area.

Compare the updated and 2025-assessment values:
```{r, compareplot3sshjac}
plot(ssh_jac_index_2025$year,
     ssh_jac_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(ssh_jac_index_2026_restr$year,
     ssh_jac_index_2026_restr$anomaly,
     type = "o",
     col = "red")
```

Now as a correlation plot
```{r, compareplot4sshjac}
plot(ssh_jac_index_2025$anomaly,
     ssh_jac_index_2026_restr$anomaly)
abline(a = 0, b = 1)

cor(ssh_jac_index_2025$anomaly,
    ssh_jac_index_2026_restr$anomaly)
```
Check above value is high.

New updated index is
```{r, sshjacindex2}
plot(ssh_jac_index_2026)
```

Will make the full figure later.


# Download variables with `verticalrange`

Going to do each separately (to keep filesizes down).

## TEMPspawn

TEMP$_{\mbox{spawn}}$ -- potential temperature `thetao`

So you restrict it to the area based on shelf break (100-2,000 m) but then still
average over the depths 130-500 m. And 31-36degN - do manually here then no need
to restrict later (for MLD had different ranges for the same download).

Adapting code from above:
```{r, tempspawn}
if(FALSE){    # set to TRUE to redownload
  data_temp_spawn_depth <- cms_download_subset(
    product = "GLOBAL_MULTIYEAR_PHY_001_030",
    layer = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    variable = "thetao",
    region = c(lon_min, 31, lon_max, 36),
    timerange = time_range,
    verticalrange = c(-130, -500), # 130m to 500m depth
    progress = TRUE
  )
  
  # data_temp_spawn_depth <- abind::adrop(data_temp_spawn_depth) # drop elevation as only
  # one value (when we were doing MLD)
  saveRDS(data_temp_spawn_depth,
          "data_temp_spawn_depth.Rds")
} else {
  data_temp_spawn_depth <- readRDS("data_temp_spawn_depth.Rds")
}

data_temp_spawn_depth
# Aha column values shows, I think, the range for elevation.
# Tease the values out:

st_dimensions(data_temp_spawn_depth)$elevation$values
# would be a bit easier to see as intervals

summary(data_temp_spawn_depth)
```
So a stars object. Note that the shallowest elevation interval is -130.6 to
-109.7, so including more than we would really want if we're saying 130m. 
Also does not include 500 m, so it's just the intervals that include the minimum
but not the maximum. 

Had said could tweak the values if what we do here does not match last year's,
but they match exactly. TODO mention in write up.

If just average each cell over elevation, will end up with something of
similar structure to what we had above, so can then just use that code. So do
the averaging first. 

TODO note somewhere that mean temperature (and current velocities) over unevenly
spaced depth layers is not a true mean, but method is consistent
over the years.

So, average over depth:
```r, tempspawnmean}
# Calculate the mean over the 'elevation' dimension, should keep the other
# dimensions and look like the MLD objects from earlier
data_temp_spawn <- st_apply(
  data_temp_spawn_depth,
  MARGIN = c("longitude", "latitude", "time"),  # dimensions to keep
  FUN = mean,
  na.rm = TRUE,
  .fname = "temp"   # not tested yet but should work
)

data_temp_spawn
summary(data_temp_spawn)
```

Now wrangle the temp values, actually this contains `zos` also, so can use for SSH.
```{r, tempspawnanalysis}
dim(data_temp_spawn)
prod(dim(data_temp_spawn))

temp_spawn_tib <- as_tibble(data_temp_spawn)    # makes one big tibble, uses that
                                       # class(data_temp_spawn) is stars
temp_spawn_tib
names(temp_spawn_tib)[4] <- "temp"   # can remove when .fname above is tested
temp_spawn_tib

expect_equal(prod(dim(data_temp_spawn)),
             nrow(temp_spawn_tib))   
```

Now can copy and adapt the SSHjac code from above. 

```{r, tempspawndata}
temp_spawn_data <- temp_spawn_tib %>%
  # select(-c("mlotst")) %>%      
  filter(latitude >= 31,
         latitude <= 36) %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 1:3) 
temp_spawn_data

summary(temp_spawn_data)
tail(temp_spawn_data)
summary(as.factor(temp_spawn_data$month)) 
```

Need to round before joining with bathymetry, since slight numerical
differences. See previous .Rmd for details.

```{r, roundingtempspawn}
temp_spawn_data_round <- temp_spawn_data %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
temp_spawn_data_round

range(diff(temp_spawn_data_round$longitude))
range(diff(temp_spawn_data_round$latitude))

# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, though just doing 1993 as presume the locations are the same
for each year, shows up well on full screen (not in pdf):
```{r, gridstempspawn}
plot(filter(temp_spawn_data_round, year == 1993, month == 1)$longitude, 
     filter(temp_spawn_data_round, year == 1993, month == 1)$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(temp_spawn_data_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(temp_spawn_data_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       pch = 20,
       cex = 0.75)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combinetempspawn}
temp_spawn_combined <- bath_round %>%
  left_join(temp_spawn_data_round)
temp_spawn_combined
temp_spawn_combined[500:520, ]
summary(temp_spawn_combined)
# See previous .Rmd for details on NAs;  mlotst ones looked to around the Channel
# Islands, around -120, 34.
# Megan did the left_join that way. I also tried the other way round, but end up
# with way more NAs; see previous .Rmd.
```

Now need just the required isobaths:
```{r, combine2tempspawn}
# isobath greater than 100m less than 2,000 for temp_spawn
temp_spawn_combined_need <- temp_spawn_combined %>%
  filter(bottom_depth >= 100 & bottom_depth <= 2000)
temp_spawn_combined_need
summary(temp_spawn_combined_need)

# temp_spawn_combined_need[is.na(temp_spawn_combined_need$zos),]
```

Plot the locations of NAs for temp:
```{r, plotmissingtempspawn}
plot(temp_spawn_combined_need$longitude[is.na(temp_spawn_combined_need$temp)],
     temp_spawn_combined_need$latitude[is.na(temp_spawn_combined_need$temp)])
```
as expected.

Okay, remove those rows:

```{r, removeNAtempspawn}
temp_spawn_combined_need_no_na <- temp_spawn_combined_need %>%
  filter(!is.na(temp))
temp_spawn_combined_need_no_na
summary(temp_spawn_combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in  MLD column
max(temp_spawn_combined_need_no_na$year)

filter(temp_spawn_combined_need_no_na, year ==
        max(temp_spawn_combined_need_no_na$year)) %>%
  summary()
```

No non-January values for final year, so remove those:

```{r, removefinalyeartempspawn}
temp_spawn_tib <- temp_spawn_combined_need_no_na %>%
  filter(year < max(temp_spawn_combined_need_no_na$year)) %>%
  group_by(year) %>%
  summarise(temp_spawn = as.numeric(mean(temp)))  #woo! we have it
            # need as.numeric else class(pull(temp_spawn_tib)) is "units"

temp_spawn_tib
temp_spawn_tib %>% tail()
```

Compare to last year's values that went into the assessment (did Megan's check
on MLDyolk in previous .Rmd).

Note there will be an extra year, so
standardising with that can result in slight differences. Will check, and also 
check that the new figures look similar enough. Maybe just doing the latter for brevity.

```{r, tempspawnindex}
temp_spawn_index_2025 <- dplyr::select(glorys_2025,
                                          year,
                                          value = "Tspawn") %>%
    dplyr::mutate(anomaly = standardise(value))
class(temp_spawn_index_2025) <- class(oni)

temp_spawn_index_2025   # As used in 2025 assessment
summary(temp_spawn_index_2025)

# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years) 
temp_spawn_index_2026_restr <- temp_spawn_tib %>%
 dplyr::filter(year <= max(temp_spawn_index_2025$year)) %>%
 dplyr::mutate(anomaly = standardise(temp_spawn))
class(temp_spawn_index_2026_restr) <- class(oni)

temp_spawn_index_2026_restr   # As calculated here
summary(temp_spawn_index_2026_restr)

# And calculate the new one for all available years
temp_spawn_index_2026 <- temp_spawn_tib %>%
 dplyr::mutate(anomaly = standardise(temp_spawn))
class(temp_spawn_index_2026) <- class(oni)

temp_spawn_index_2026   # As calculated here. 
summary(temp_spawn_index_2026)
```


Compare the updated and 2025-assessment values:
```{r, compareplot3tempspawn}
plot(temp_spawn_index_2025$year,
     temp_spawn_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(temp_spawn_index_2026_restr$year,
     temp_spawn_index_2026_restr$anomaly,
     type = "o",
     col = "red")
```
Wow, awesome. 

Now as a correlation plot
```{r, compareplot4tempspawn}
plot(temp_spawn_index_2025$anomaly,
     temp_spawn_index_2026_restr$anomaly)
abline(a = 0, b = 1)

cor(temp_spawn_index_2025$anomaly,
    temp_spawn_index_2026_restr$anomaly)
```
Check above value is high.

New updated index is
```{r, tempspawnindex2}
plot(temp_spawn_index_2026)
```

Will make the full figure later.



## ASTeggs

AST$_{\mbox{eggs}}$ -- alongshore transport of eggs. Try just northward first.

So you restrict it to the area based on shelf break (100-2,000 m) but then still
average over the depths 40-60 m. And 31-36degN. Maybe some
rough trigonometry from `uo` and `vo`, respectively the eastward and northward
ocean current velocity. Or just northward, given hypothesis above. Try that first.

Transport is volume $\times$ speed ish, so may need to do something more complicated.

Adapting code from TEMPspawn:
```{r, asteggs}
if(FALSE){    # set to TRUE to redownload
  data_ast_eggs_depth <- cms_download_subset(
    product = "GLOBAL_MULTIYEAR_PHY_001_030",
    layer = "cmems_mod_glo_phy_my_0.083deg_P1M-m",
    variable = c("uo", "vo"),
    region = c(lon_min, 31, lon_max, 36),
    timerange = time_range,
    verticalrange = c(-40, -60), # 130m to 500m depth
    progress = TRUE
  )
  
  # data_ast_eggs_depth <- abind::adrop(data_ast_eggs_depth) # drop elevation as only
  # one value (when we were doing MLD)
  saveRDS(data_ast_eggs_depth,
          "data_ast_eggs_depth.Rds")
} else {
  data_ast_eggs_depth <- readRDS("data_ast_eggs_depth.Rds")
}

data_ast_eggs_depth
# Aha column values shows, I think, the range for elevation.
# Tease the values out:

st_dimensions(data_ast_eggs_depth)$elevation$values
# would be a bit easier to see as intervals

summary(data_ast_eggs_depth)
```
For TEMPspawn we had: Note that the shallowest elevation interval is -130.6 to
-109.7, so including more than we would really want if we're saying 130m. TODO
could tweak the values if what we do here does not match last year's.

But here we asked for 40-60 and got interval that contains 40 but not 60. So
it's always on the shallow side. 

So, average over depth - or do we want the sum northward to work out advection,
or what??? TODO wait to hear from others. Just try northward for now.
```r, asteggsmean}
# Calculate the mean over the 'elevation' dimension, should keep the other
# dimensions and look like the MLD objects from earlier
data_ast_eggs <- st_apply(
  data_ast_eggs_depth,
  MARGIN = c("longitude", "latitude", "time"),  # dimensions to keep
  FUN = mean,
  na.rm = TRUE
  # .fname = "temp"   # not tested yet but should work
)

data_ast_eggs
summary(data_ast_eggs)
```

Surprising if `st_apply()` works with two attributes (help does not say), but it
has produced something of the correct size. Ah, and not changed the attribute
names, which is slightly inconsistent with one-attribute things earlier (which
renamed it `mean` I think, which here would be ambiguous).

Now wrangle the values
```{r, asteggsanalysis}
dim(data_ast_eggs)
prod(dim(data_ast_eggs))

ast_eggs_tib <- as_tibble(data_ast_eggs)    # makes one big tibble, uses that
                                       # class(data_ast_eggs) is stars
ast_eggs_tib    # so this keeps uo and vo, which are the mean velocities over
# the depth  
# names(ast_eggs_tib)[4] <- "temp"   # can remove when .fname above is tested
ast_eggs_tib

expect_equal(prod(dim(data_ast_eggs)),
             nrow(ast_eggs_tib))   
```

Now do latitude (shouldn't change anything as specified in download) and
required months:
```{r, asteggsdata}
ast_eggs_data <- ast_eggs_tib %>%
  # select(-c("mlotst")) %>%      
  filter(latitude >= 31,
         latitude <= 36) %>%
  mutate(year = year(time),
         month = month(time))%>% 
  filter(month %in% 1:3) 
ast_eggs_data

summary(ast_eggs_data)
tail(ast_eggs_data)
summary(as.factor(ast_eggs_data$month)) 
```

Still has `uo` and `vo`.

Need to round before joining with bathymetry, since slight numerical
differences. See previous .Rmd for details.

```{r, roundingasteggs}
ast_eggs_data_round <- ast_eggs_data %>%
  mutate(across(c(longitude, latitude),
                round,
                dp))
ast_eggs_data_round

range(diff(ast_eggs_data_round$longitude))
range(diff(ast_eggs_data_round$latitude))

# Then do the same for the bathymetry data sets
range(diff(bath$longitude))
range(diff(bath$latitude))

range(diff(bath_round$longitude))
range(diff(bath_round$latitude))
```

Check the two grids, though just doing 1993 as presume the locations are the same
for each year, shows up well on full screen (not in pdf):
```{r, gridsasteggs}
plot(filter(ast_eggs_data_round, year == 1993, month == 1)$longitude, 
     filter(ast_eggs_data_round, year == 1993, month == 1)$latitude,
     xlim = range(c(range(bath_round$longitude),
                    range(ast_eggs_data_round$longitude))),
     ylim = range(c(range(bath_round$latitude),
                    range(ast_eggs_data_round$latitude))),
     main = "Red is bathymetry")
points(bath_round$longitude,
       bath_round$latitude,
       col = "red",
       pch = 20,
       cex = 0.75)
```

Makes sense - MLD grid is on a rectangle (but presumably NA's are on land), and
bathymetry grid is already just the ocean.

```{r, combineasteggs}
ast_eggs_combined <- bath_round %>%
  left_join(ast_eggs_data_round)
ast_eggs_combined
ast_eggs_combined[500:520, ]
summary(ast_eggs_combined)
# See previous .Rmd for details on NAs;  mlotst ones looked to around the Channel
# Islands, around -120, 34.
# Megan did the left_join that way. I also tried the other way round, but end up
# with way more NAs; see previous .Rmd.
```

Now need just the required isobaths:  HERE
```{r, combine2asteggs}
# isobath greater than 100m less than 2,000 for ast_eggs
ast_eggs_combined_need <- ast_eggs_combined %>%
  filter(bottom_depth >= 100 & bottom_depth <= 2000)
ast_eggs_combined_need
summary(ast_eggs_combined_need)

# ast_eggs_combined_need[is.na(ast_eggs_combined_need$zos),]
```

Plot the locations of NAs for uo:
```{r, plotmissingasteggs}
plot(ast_eggs_combined_need$longitude[is.na(ast_eggs_combined_need$uo)],
     ast_eggs_combined_need$latitude[is.na(ast_eggs_combined_need$uo)])
```
as expected.

Okay, remove those rows:

```{r, removeNAasteggs}
ast_eggs_combined_need_no_na <- ast_eggs_combined_need %>%
  filter(!is.na(uo))
ast_eggs_combined_need_no_na
summary(ast_eggs_combined_need_no_na)
# No NA's left, even though only removed the rows with NAs in only uo column
max(ast_eggs_combined_need_no_na$year)

filter(ast_eggs_combined_need_no_na, year ==
        max(ast_eggs_combined_need_no_na$year)) %>%
  summary()
```

No non-January values for final year, so remove those; this is a bit different
to earlier indices as we still have uo and vo, but can easily choose one or the other:
```{r, removefinalyearasteggs}
ast_eggs_tib <- ast_eggs_combined_need_no_na %>%
  filter(year < max(ast_eggs_combined_need_no_na$year)) %>%
  group_by(year) %>%
  summarise(ast_eggs_uo = as.numeric(mean(uo)),
            ast_eggs_vo = as.numeric(mean(vo)),)  #woo! we have it
            # need as.numeric else class(pull(ast_eggs_tib)) is "units"

ast_eggs_tib
ast_eggs_tib %>% tail()
```

Compare to last year's values that went into the assessment (did Megan's check
on MLDyolk in previous .Rmd).

Note there will be an extra year, so
standardising with that can result in slight differences. Will check, and also 
check that the new figures look similar enough. Maybe just doing the latter for brevity.

### Eastward flow only

Here we need a uo and vo index calculated this year, to see if either match what
was used last year. Do uo (eastward) first:
```{r, asteggsindexuo}
ast_eggs_index_2025 <- dplyr::select(glorys_2025,
                                          year,
                                          value = "Tspawn") %>%
    dplyr::mutate(anomaly = standardise(value))
class(ast_eggs_index_2025) <- class(oni)

ast_eggs_index_2025   # As used in 2025 assessment
summary(ast_eggs_index_2025)

# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years)
# uo first for asteggs
ast_eggs_index_2026_restr_uo <- ast_eggs_tib %>%
 dplyr::filter(year <= max(ast_eggs_index_2025$year)) %>%
 dplyr::select(c("year", "ast_eggs_uo")) %>%
 dplyr::mutate(anomaly = standardise(ast_eggs_uo))    # with uo now in the
# object name
class(ast_eggs_index_2026_restr_uo) <- class(oni)

ast_eggs_index_2026_restr_uo   # As calculated here
summary(ast_eggs_index_2026_restr_uo)

# And calculate the new one for all available years
ast_eggs_index_2026_uo <- ast_eggs_tib %>%
 dplyr::select(c("year", "ast_eggs_uo")) %>%    
 dplyr::mutate(anomaly = standardise(ast_eggs_uo))
class(ast_eggs_index_2026_uo) <- class(oni)

ast_eggs_index_2026_uo   # As calculated here. 
summary(ast_eggs_index_2026_uo)
```


Compare the updated and 2025-assessment values:
```{r, compareplot3asteggsuo}
plot(ast_eggs_index_2025$year,
     ast_eggs_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(ast_eggs_index_2026_restr_uo$year,
     ast_eggs_index_2026_restr_uo$anomaly,
     type = "o",
     col = "red")
```
So, as expected, last year's was not simply the mean eastward velocity. Now see if
it was just the mean northward velocity.

### Northward flow only

Here we need a uo and vo index calculated this year, to see if either match what
was used last year:
```{r, asteggsindexvo}
# Calculating here but just restricting to the same years as used for 2025
# assessment (need to do for comparison because standardisation depends on the
# years)
# vo first for asteggs
ast_eggs_index_2026_restr_vo <- ast_eggs_tib %>%
 dplyr::filter(year <= max(ast_eggs_index_2025$year)) %>%
 dplyr::select(c("year", "ast_eggs_vo")) %>%
 dplyr::mutate(anomaly = standardise(ast_eggs_vo))    # with vo now in the
# object name
class(ast_eggs_index_2026_restr_vo) <- class(oni)

ast_eggs_index_2026_restr_vo   # As calculated here
summary(ast_eggs_index_2026_restr_vo)

# And calculate the new one for all available years
ast_eggs_index_2026_vo <- ast_eggs_tib %>%
 dplyr::select(c("year", "ast_eggs_vo")) %>%    
 dplyr::mutate(anomaly = standardise(ast_eggs_vo))
class(ast_eggs_index_2026_vo) <- class(oni)

ast_eggs_index_2026_vo   # As calculated here. 
summary(ast_eggs_index_2026_vo)
```

Compare the updated and 2025-assessment values:
```{r, compareplot3asteggsvo}
plot(ast_eggs_index_2025$year,
     ast_eggs_index_2025$anomaly,
     type = "o",
     col = "blue")
lines(ast_eggs_index_2026_restr_vo$year,
     ast_eggs_index_2026_restr_vo$anomaly,
     type = "o",
     col = "red")
```

All over the place, so index is not based on northward flow only. TODO wait to
hear back from NOAA colleagues.

Now as a correlation plot, just for uo
```{r, compareplot4asteggsuo}
plot(ast_eggs_index_2025$anomaly,
     ast_eggs_index_2026_restr_uo$anomaly)
abline(a = 0, b = 1)

cor(ast_eggs_index_2025$anomaly,
    ast_eggs_index_2026_restr_uo$anomaly)
```


New updated index is
```{r, asteggsindex2}
# plot(ast_eggs_index_2026_uo)
```

Will make the full figure later.

HERE not much point in trying PUlatelarv wich is poleward undercurrent, for
which we need a similar calculation probably regarding how its calculated.


```{r, saveindices}
glorys_indices_2026 <- list(
    temp_spawn_index = temp_spawn_index_2026,
    ssh_jac_index = ssh_jac_index_2026,
    # ast_eggs_index = ast_eggs_index_2026,
    # pu_late_larv_index = pu_late_larv_index_2026,
    mld_late_larv_index = mld_late_larv_index_2026,
    mld_yolk_index = mld_yolk_index_2026)
saveRDS(glorys_indices_2026,
        "glorys_indices_2026.Rds")
```    
    


