# Changes from `r last_assess_yr`
## Changes from the `r last_assess_yr` assessment

* Update Stock Synthesis to version `r ss_version`
* Add new catch data for `r last_assess_yr` and update historical catches 
* Add new weight-at-age data for `r last_assess_yr` and update
  weight-at-age data
* Add the `r last_assess_yr` acoustic survey biomass estimate and age data
* Add the `r last_assess_yr` age-1 index estimate
* Add new fishery age composition data for `r last_assess_yr` and historical
  fishery age composition data
* No changes to acoustic survey biomass estimates or age-1 index
  estimates this year
* All model runs in the assessment use MCMC

## Bridging: Model Updates and Additions

```{r bridge-biomass, fig.height = 4, out.height = "80%"}
p <- list()
p[[1]] <-
  plot_rel_biomass(d_obj = d_obj_bridge_rel_biomass[[1]],
                   ylim = c(0, 2.5),
                   x_labs_mod = 10,
                   wrap_y_label = TRUE,
                   leg_pos = "none")
p[[2]] <-
  plot_recdevs(d_obj = d_obj_bridge_recdev[[1]],
               leg_pos = "none",
               x_labs_mod = 10,
               line_width = 0.1)

plot_grid(plotlist = p,
          nrow = 1,
          ncol = 2)
```

## Bridging: Model Updates and Additions (survey)

```{r bridge-survey, fig.height = 4, out.height = "80%"}
p <- list()
p[[1]] <-
  plot_survey_index_fits(d_obj = d_obj_bridge_age2_index[[1]],
                         survey_type = "age2",
                         rev_colors = FALSE,
                         leg_pos = "none")

p[[2]] <-
  plot_survey_index_fits(d_obj = d_obj_bridge_age1_index[[1]],
                         survey_type = "age1",
                         xlim = c(1995, 2022),
                         ylim = c(0, 11),
                         y_breaks = seq(0, 11, by = 2),
                         rev_colors = FALSE,
                         leg_pos = "none")

plot_grid(plotlist = p,
          nrow = 1,
          ncol = 2)
```

## Bridging: Summary

\bsmall

* Updating Stock Synthesis software has no effect on results. Relevant
  changes include:
  - Weight-at-age specified in first year of forecast, rather than
    carried forward
  - Report files for models using empirical weight-at-age for growth do
    not supply output that uses growth parameters
* Updating data previous to `r last_data_yr` had very little effect on the
  stock trajectory
  \pause
* Adding `r last_data_yr` catch and weight-at-age data resulted in a slight
  upward shift in current female spawning biomass, because the mean weights
  in `r last_data_yr` were mostly above the recent 5-year average mean
  weights (assumption before the addition of these data)
* Adding the `r last_data_yr` survey biomass estimate and age compositions
  leads to slightly more optimistic estimates of biomass for the last few
  years and reduces uncertainty
* The addition of the age-1 index resulted in a small increase in $B_0$,
  an increase in relative spawning biomass from 2019--2022, and an
  increase in the estimated size of the 2018 and 2020 cohorts
* Adding `r last_assess_yr` fishery age composition data had little impact
  on historical estimates, shifted recent recruitment estimates
  (esp. 2019--2021), and increased estimates of 2022 (slightly) and 2023
  (considerably) female spawning biomass.
  
\esmall
