# BCCM

## British Columbia Continental Margin (BCCM) model

Physical biogeochemical oceanographic model

- Peña et al. (2019) [Deep-Sea Research II](doi:10.1016/j.dsr2.2019.104637)

- Regional Ocean Modelling Systen (ROMS)

- curvilinear grid at 3 km x 3 km resolution

- we interpolate to regular `r colorize("2 km x 2km grid")`

- `r colorize("161,025 cells", "red")`

- monthly from 1993-2019

- huge filesizes required hosting outputs on Zenodo not GitHub

<!-- - `r colorize("new grid also used for HOTSSea outputs, and has interpolated
  depths", "red")` -->

<!--
\
\footnotesize
Peña, M.A., Fine, I. and Callendar, W. 2019. Interannual variability in primary production and shelf-offshore transport of nutrients along the northeast Pacific Ocean margin. Deep-Sea Research II, doi:10.1016/j.dsr2.2019.104637.
\normalsize
-->

## Available variables

The variables are:

   - dissolved oxygen concentration
   - pH
   - salinity
   - temperature
   - depth-integrated phytoplankton
   - depth-integrated primary production.

For applicable variables these are given for

   - sea surface
   - 0-40 m integration
   - 40-100 m integration
   - 100 m to the sea bottom
   - sea bottom.


## List of available variables (full domain; Canadian EEZ also available)
bccm\_data\_full for available variables
```{r bccm_data_full list}
bccm_data_full
```

## Built in plotting - primary production in April 2018

```{r bccmplotfull}
plot(bccm_primaryproduction_full())
```

## Built in plotting - primary production throughout  2018

```{r bccmplotfullyear, fig.height = 5}
plot(bccm_primaryproduction_full(), months = c(1, 4, 7, 10))
```


<!--
pdf was getting locked, 12 months doesn't fit
-->
