# BCCM

## British Columbia Continental Margin (BCCM) model

Physical biogeochemical oceanographic model

- Regional Ocean Modelling Systen (ROMS)

- Curvilinear grid at 3km x 3km resolution

- We interpolate to regular grids

\
Model output provided by Angelica Peña at the Institute of Ocean Sciences (Fisheries and
Oceans Canada).

\
\footnotesize
Peña, M.A., Fine, I. and Callendar, W. 2019. Interannual variability in primary production and shelf-offshore transport of nutrients along the northeast Pacific Ocean margin. Deep-Sea Research II, doi:10.1016/j.dsr2.2019.104637.
\normalsize

## Mapping to grids

Original mapping:

- clipped to Canada's Pacific Exclusive Economic Zone

- Inshore (2 km x 2 km) and offshore (6 km x 6 km)

- 40,480 cells

- Monthly from 1993-2019

New additional mapping:

- `r colorize("covers full domain (into US waters)", "red")`

- `r colorize("2 km x 2km grid throughout", "red")`

- `r colorize("161,025 cells", "red")`

- Monthly from 1993-2019

- `r colorize("New grid also used for HOTSSea outputs, and has interpolated
  depths", "red")`


## Available variables

The variables are:

   - dissolved oxygen concentration
   - pH
   - salinity
   - temperature
   - depth-integrated phytoplankton
   - depth-integrated primary production.

For applicable variables these are given for

   - sea surface
   - 0-40 m integration
   - 40-100 m integration
   - 100 m to the sea bottom
   - sea bottom.


## List of available variables (original just Canada)
bccm_data for available variables
```{r bccm_data list}
bccm_data
```

## List of available variables (full domain)
bccm\_data\_full for available variables
```{r bccm_data_full list}
bccm_data_full
```

## Built in plotting - primary production in Canadian waters

```{r bccmplot}
plot(bccm_primaryproduction())
```

## Built in plotting - primary production in Canadian waters

```{r bccmplotfull}
plot(bccm_primaryproduction_full())
```

## Built in plotting - monthly primary production through 2018

```{r bccmplot2}
plot(bccm_primaryproduction_full(), months = 1:12)
```
