# BCCM

## British Columbia Continental Margin (BCCM) model
\pause
Physical biogeochemical oceanographic model 

- Regional Ocean Modelling Systen (ROMS)

- Curvilinear grid at 3km x 3km resolution

- Interpolated to regular grid

- Clipped to Canada's Pacific EEZ

\
Data provided by Angelica Peña at the Institute of Ocean Sciences (Fisheries and
Oceans Canada).

\
\footnotesize
Peña, M.A., Fine, I. and Callendar, W. 2019. Interannual variability in primary production and shelf-offshore transport of nutrients along the northeast Pacific Ocean margin. Deep-Sea Research II, doi:10.1016/j.dsr2.2019.104637.
\normalsize


## Nearest neighbour linear interpolation

Curvilinear --> regular grid

- Inshore (2 x 2km) and offshore (6 x 6km)

- Nearest neighbour interpolation 

  - Performace = RMSE error 
  
- 40,480 cells 

- Monthly from 1993-2019


## List of available layers
bccm_data for available variables
```{r bccm_data list}
bccm_data
```

## Downloading data locally
The data are saved as a `sf` (simple feature) in a separate GitHub repository `pacea-data` (https://github.com/pbs-assess/pacea-data).

\
There are built in functions to download specific variables:
```{r bccmdataDLprompt, eval=FALSE}
bccm_avg100mtoBot_ph()

pacea would like to download and store these data in the directory:
C:\Users\TAIT\AppData\Local/pacea/Cache
Is that okay?
1: Yes
2: No

Selection: 
```

```{r bccm data head, eval=FALSE}
Simple feature collection with 33642 features and 324 fields
Geometry type: POLYGON
Dimension:     XY
Bounding box:  xmin: 164671.6 ymin: 163875 xmax: 1097612 ymax: 1103653
Projected CRS: NAD83 / BC Albers
# A tibble: 33,642 × 325
  `1993_1` `1993_2` `1993_3` `1993_4` ...                       geometry
     <dbl>    <dbl>    <dbl>    <dbl> ...                  <POLYGON [m]>
1     7.83     7.87     7.88     7.90 ... ((575612.5 1103653, 577612.5 ~
2     7.83     7.87     7.88     7.90 ... ((577612.5 1103653, 579612.5 ~
3     7.83     7.87     7.88     7.90 ... ((579612.5 1103653, 581612.5 ~
4     7.83     7.87     7.88     7.89 ... ((581612.5 1103653, 583612.5 ~
5     7.79     7.82     7.84     7.82 ... ((619612.5 1103653, 621612.5 ~
6     7.79     7.82     7.83     7.82 ... ((621612.5 1103653, 623612.5 ~
# i 33,636 more rows
# i 320 more variables: ...
```

## Downloading data locally
\
Or all variables (will take several minutes):
```{r bccm_all, eval=FALSE}
bccm_all_variables()

Downloading all BCCM data may take many minutes. Files will be downloaded to pacea_cache directory:
C:\Users\TAIT\AppData\Local/pacea/Cache
Would you like to continue?
1: Yes
2: No

Selection: 
```


## Help file
```{r bccmhelp, eval = FALSE}
?bccm_surface_oxygen
```

```{r bccmhelp2, eval = FALSE}
BCCM environmental variable data

Description
  Loading and/or downloading British Columbia continental margin (BCCM) model data for the BC Pacific region to local drive.

  To download all BCCM data, use the function 'bccm_all_variables()'. See help page for details ('?bccm_all_variables').

Usage
  bccm_bottom_oxygen(update = FALSE, ask = interactive(), force = FALSE)

  bccm_bottom_ph(update = FALSE, ask = interactive(), force = FALSE)

  bccm_bottom_salinity(update = FALSE, ask = interactive(), force = FALSE)
  ...
  
Arguments
  update	  Logical. Would you like to check for a newer version of the layer?

  ask	      Logical. Should the user be asked before downloading the data to local cache? Defaults to the value of interactive().

  force	    Logical. Should download of data be forced? Overrides 'ask' argument if TRUE.

Format
  A simple features dataframe.
...
```

## Built in plotting

```{r bccmplot}
plot(bccm_surface_temperature(), months.plot = c("April", "july"), years.plot = c(2013, 2016, 2019))
```

## Built in plotting

```{r bccmplot2}
plot(bccm_surface_ph())
```

## Extras -- climatology function

Built in functions to calculate 30-year climatology

```{r calc_climhelp, eval=FALSE}
?calc_clim
```

```{r calc_climhelp2, eval=FALSE}
Description
  Function for calculating climatology of a pacea data object.

Usage
calc_clim(
  data,
  clim_years = c(1991:2020),
  clim_time = "month",
  time_period_return = "all"
)

Arguments
  data	               pacea data object: BCCM, OISST, buoy_sst data only

  clim_years	         climatology period years

  clim_time	           time units (e.g. month) to summarize climatologies

  time_period_return	 vector of value(s) for the specific time units to estimate climatologies (e.g. '4' for week 4 or April). Set to equal 'all' for all time units.
...
```

## Extras -- climatology function

Output with each climatology time unit (e.g. week, month)
```{r calc_clim, eval=FALSE}
calc_clim(bccm_surface_temperature())
```

```{r calc_clim2, eval=FALSE}
Simple feature collection with 486960 features and 4 fields
Geometry type: POLYGON
Dimension:     XY
Bounding box:  xmin: 164671.6 ymin: 163875 xmax: 1097612 ymax: 1103653
Projected CRS: NAD83 / BC Albers
# A tibble: 486,960 x 5
  month clim_value clim_sd clim_n                  geometry
  <dbl>      <dbl>   <dbl>  <int>             <POLYGON [m]>
1     1       5.80   0.721     27 ((575612.5 1103653, 5776~
2     2       5.47   0.610     27 ((575612.5 1103653, 5776~
3     3       5.64   0.627     27 ((575612.5 1103653, 5776~
4     4       6.61   0.623     27 ((575612.5 1103653, 5776~
5     5       8.49   0.732     27 ((575612.5 1103653, 5776~
6     6      10.5    0.662     27 ((575612.5 1103653, 5776~
# i 486,954 more rows
# i Use `print(n = ...)` to see more rows
```
\
And we can also use this with the OISST and buoy data
```{r calc_climother, eval=FALSE}
calc_clim(oisst_7day)
calc_clim(buoy_sst)
```

## Extras -- anomaly function

Built in functions to calculate anomaly to the 30 year climatology

```{r calc_anomhelp, eval=FALSE}
?calc_anom
```

```{r calc_anomhelp2, eval=FALSE}
Description
  Function for calculating climatology of a pacea data object.

Usage
calc_anom(
  data,
  clim_years = c(1991:2020),
  clim_time = "month",
  time_period_return = "all",
  years_return
)

Arguments
  data	              pacea data object: BCCM, OISST, buoy_sst data only
  
  clim_year	          climatology period years

  clim_time	          time units (e.g. month) to summarize climatologies

  time_period_return	vector of value(s) for the specific time units to estimate climatologies (e.g. '4' for week 4 or April). Set to equal 'all' for all time units.

  years_return	      vector of value(s) to return the years of interest. Defaults to all years in input data
...
```



## Extras -- anomaly function

Specify year(s) and time period (e.g. week, month) to return - otherwise all.
```{r calc_anom, eval=FALSE}
(anom.dat <- calc_anom(bccm_surface_temperature(), time_period_return = "April", years_return = c(2010, 2015)))
```

```{r calc_anom2, eval=FALSE}
Simple feature collection with 40580 features and 2 fields
Geometry type: POLYGON
Dimension:     XY
Bounding box:  xmin: 164671.6 ymin: 163875 xmax: 1097612 ymax: 1103653
Projected CRS: NAD83 / BC Albers
# A tibble: 40,580 x 3
  `2010_4` `2015_4`                                                                       geometry
     <dbl>    <dbl>                                                                  <POLYGON [m]>
1    0.114     1.09 ((575612.5 1103653, 577612.5 1103653, 577612.5 1101653, 575612.5 1101653, 575~
2    0.124     1.08 ((577612.5 1103653, 579612.5 1103653, 579612.5 1101653, 577612.5 1101653, 577~
3    0.150     1.07 ((579612.5 1103653, 581612.5 1103653, 581612.5 1101653, 579612.5 1101653, 579~
4    0.204     1.08 ((581612.5 1103653, 583612.5 1103653, 583612.5 1101653, 581612.5 1101653, 581~
5    0.204     1.07 ((583612.5 1103653, 585612.5 1103653, 585612.5 1101653, 583612.5 1101653, 583~
6    0.244     1.08 ((585612.5 1103653, 587612.5 1103653, 587612.5 1101653, 585612.5 1101653, 585~
# i 40,574 more rows
# i Use `print(n = ...)` to see more rows
```

## Extras -- anomaly plot

```{r anom_plot, warning=FALSE}
plot(anom.dat, months.plot = "April", years.plot = c(2010, 2015))
```

## Extras -- other tidbits in vignettes

Geospatial functions in vignettes

- `area_mean()` estimates inshore and offshore spatial mean (200m isobath)

- Masking/clipping specific areas (e.g. fishing regions of interest)

- Various other data visualisation examples





