# Buoys

```{asis, echo = long_talk}
## Sea surface temperature

Andrea Hilborn (IOS) maintains
https://github.com/IOS-OSD-DPG/Pacific_SST_Monitoring
which is updated weekly to show recent SST from various sources for the Northeast Pacific (and anomalies
and marine heatwave information).

  \
```
```{r andreapage, echo=FALSE, out.width = "80%", eval = long_talk}
knitr::include_graphics(paste0(here::here(),
                               "/inst/andrea-github.png"))
```


## Sea surface temperature data from buoys

Data from `r nrow(buoy_metadata)` buoys maintained by DFO/ECCC (Environment and
Climate Change Canada), hosted through CIOOS (Canadian Integrated Ocean
Observing System).

```{r buoysmap, echo=FALSE, out.width = "50%"}
knitr::include_graphics(paste0(here::here(),
                               "/man/figures/buoys_map.png"))
```

## Sea surface temperature data from buoys

The data are saved as a `tibble` in `pacea`:
```{r buoysdata}
buoy_sst
```
\pause
```{r buoysdata2}
tail(buoy_sst)
```

## Sea surface temperature data from buoys

Plot data from a single buoy for all years, the default is for buoy C46205:
```{r buoysplot}
plot(buoy_sst)
```
