# Buoys

## Sea surface temperature

Andrea Hilborn (IOS) maintains
https://github.com/IOS-OSD-DPG/Pacific_SST_Monitoring
which is updated weekly to show recent SST from various sources for the Northeast Pacific (and anomalies
and marine heatwave information).

```{r andreapage, echo=FALSE, out.width = "80%"}
knitr::include_graphics(paste0(here::here(),
                               "/inst/andrea-github.png"))
```

## Sea surface temperature data from buoys

Data from `r nrow(buoy_metadata)` buoys maintained by DFO/ECCC (Environment and
Climate Change Canada).


```{r buoysmap, echo=FALSE, out.width = "50%"}
knitr::include_graphics(paste0(here::here(),
                               "/man/figures/buoys_map.png"))
```

## Sea surface temperature data from buoys

Extensive data processing to get daily mean SST:

- Resolution is every 10 minutes or

\pause

- TODO update when finalise processing


## Sea surface temperature data from buoys

The data are saved as a `tibble` in `pacea`:
```{r buoysdata}
buoy_sst

tail(buoy_sst)
```

## Help file
```{r buoyshelp, eval=FALSE}
?oisst_month
TODO paste help details here

```

## New slide

```{r buoysplot}
plot(hake_biomass)
```
