# Buoys

## Sea surface temperature

Andrea Hilborn (IOS) maintains
https://github.com/IOS-OSD-DPG/Pacific_SST_Monitoring
which is updated weekly to show recent SST from various sources for the Northeast Pacific (and anomalies
and marine heatwave information).

  \

```{r andreapage, echo=FALSE, out.width = "80%"}
knitr::include_graphics(paste0(here::here(),
                               "/inst/andrea-github.png"))
```

## Sea surface temperature data from buoys

Data from `r nrow(buoy_metadata)` buoys maintained by DFO/ECCC (Environment and
Climate Change Canada), hosted through CIOOS (Canadian Integrated Ocean
Observing System).


```{r buoysmap, echo=FALSE, out.width = "50%"}
knitr::include_graphics(paste0(here::here(),
                               "/man/figures/buoys_map.png"))
```

## Sea surface temperature data from buoys

Extensive data processing to calculate daily mean SST:

- Resolution is every 10 minutes or every hour

- Come with quality control flags (mostly)

- We decided that for a daily mean we require:
  - at least one measurement every 2 hours
  - $\geq 10$ 2-hour measurements every day

- Switch from DFO data stream to ECCC data stream when appropriate

- Still working on fixing a few outliers in the high-resolution data (hard to automate)


## Sea surface temperature data from buoys

The data are saved as a `tibble` in `pacea`:
```{r buoysdata}
buoy_sst
```
\pause
```{r buoysdata2}
tail(buoy_sst)
```

## Sea surface temperature data from buoys

Plot data from a single buoy for all years, the default is for buoy C46205:
```{r buoysplot}
plot(buoy_sst)
```

## Sea surface temperature data from buoys

Easy to look at values for any buoy
```{r buoysplot2}
plot(buoy_sst,
     stn_id = "C46185")
```

## Help file for every object
```{r buoyshelp, eval=FALSE}
?buoy_sst

Daily average sea surface temperatures for 19 buoys (yielding over
200,000 values) calculated with data from Environment and Climate
Change Canada (ECCC) and DFO.

Description:

     A tibble of daily average calculations of sea surface temperature
     in coastal Canadian Pacific waters. The earliest data are from
     September 1987, and 14 buoys were still providing data as of May
     2023. See the example code below to see the start and end dates
     for each buoy.

Usage:

     buoy_sst

Format:

     A tibble with columns:

     date: date (of class `Date`) used to calculate the `sst`, based on
          UTC -8 hours (i.e. Pacific Standard Time, not changing due to
          daylight savings)

    ...
```

## Vignettes for each type of data

We have written vignettes for each type of data. Viewable directly on GitHub
site (see the README).

```{r buoysvignette, echo=FALSE, out.width = "60%"}
knitr::include_graphics(paste0(here::here(),
                               "/inst/buoys-vignette-screenshot.png"))
```
