# HOTSSea

- Hindcast of the Salish Sea (HOTSSea)
- physical ocean model from 1980 to 2018 [Oldford et al., (*Geoscientific Model
Development*, in press)](https://doi.org/10.5194/gmd-2024-58).
- pacea includes temperature and salinty
- mapped to same 2 km x 2km grid as `bccm_full` (slight overlap in domain)
- 6,165 spatial cells
- every month from January 1980 to December 2018
- various depth ranges
- monthly statistics (`min`, `mean`, `max`, `std`)
- saved in the same format as BCCM and OISST output (so same plotting
  etc. functions can be used)

## Available variables
HOTSSea objects in pacea are (40 in total):
```{r hotssea}
hotssea_data
```

## Example plot

```{r hotsseaplot, fig.width = 12}
plot(hotssea_avg30to150m_temperature_max())
```

## File storage solutions

- original BCCM results saved in separate GitHub repo: `pacea-data`
- full BCCM and HOTSSea results are much larger, so saved on Zenodo
- get automatically downloaded and locally cached when users first need them

\
\footnotesize
Oldford, G.L., Jarníková, T., Christensen, V., and Dunphy, M. (in press).
HOTSSea v1: a NEMO-based physical Hindcast of the Salish Sea (1980–2018)
supporting ecosystem model
development. *Geoscientific Model
Development*. https://doi.org/10.5194/gmd-2024-58 .
\normalsize
