# Indices

## Climatic and oceanographic indices

Various climate and oceanographic indices are currently included in pacea:

```{r indicestable, echo=FALSE}

kableExtra::kable(pacea_indices, "latex", booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"))
```

## Climatic and oceanographic indices

For example, to see a plot of the Oceanic Niño Index (ONI) anomaly, simply type

```{r oniplot}
plot(oni)
```

This shows the onset of El Niño (positive index) conditions in April 2023.

## Climatic and oceanographic indices

The values are readily available for doing your own analyses:
```{r oni}
oni
tail(oni)
```

```{asis, echo = long_talk}
## Climatic and oceanographic indices

Each climatic and oceanographic index is saved as a data
object, and properly documented, described, and referenced in its help file, for
example:
```
```{r onihelp, eval = FALSE, echo = long_talk}
?oni

ONI - Oceanographic Niño Index

Description:

     The Oceanic Niño Index is a monthly index which is one measure of
     the El Niño-Southern Oscillation.

Usage:

     oni

Format:

     A tibble also of class `pacea_index` with columns:

     year: year of value

     month: month (1 to 12) of value

     val: absolute values of three-month averages (preceding, current,
          and next month), deg C; note that recent values may change in
          subsequent updates - see details

     anom: anomalies based on 30-year base periods that are updated
          every 5 years, deg C; note that recent values may change in
          subsequent updates - see details
```
```{asis, echo = long_talk}
## Climatic and oceanographic indices

Continuing the help...
```
```{r onihelp2, eval = FALSE, echo = long_talk}
     The Oceanic Niño Index (ONI) is a 3-month running mean of sea
     surface temperature (SST) anomalies in the Niño 3.4 region (5 deg
     N to 5 deg S, 120 deg W to 170 deg W) plotted on the center month.
     The SST anomalies are calculated based on 30-year base periods
     that are updated every 5 years, which accounts for global warming
     and some of the decadal-scale SST variability (as seen in the
     Pacific Decadal Oscillation index). The ONI is provided by the
     NOAA’s National Weather Service National Centers for Environmental
     Prediction CPC:
     http://www.cpc.ncep.noaa.gov/products/analysis_monitoring/ensostuff/ensoyears.shtml

     ...

         Because of the high frequency filter applied to the ERSSTv5 data,
     `ONI values may change up to two months after the initial real
     time value is posted`. Therefore, the most recent ONI values
     should be considered an estimate.
```

```{asis, echo = more_details}
## Climatic and oceanographic indices

So for some analyses you may want to restrict values to a specified time period
(for which values should not change in future `pacea` updates), for example

  \
```
```{r onilocal, echo = more_details, eval = more_details}
oni_fix <- dplyr::filter(oni,
                         year < 2023 | year == 2023 & month < 3)
tail(oni_fix)       # Will always end with February 2023
```

```{asis, echo = more_details}
## Further plotting options and styles

Each index has a default plotting style (which you can override). For `oni` it
shows the monthly anomalies as colour-code bars:
```

```{r oniplot2, echo = more_details, eval = more_details}
plot(oni)
```

```{asis, echo = more_details}
## Further plotting options and styles

Another option is as a black line with filled-in colouring:
```

```{r oniplot3, echo = more_details, eval = more_details}
plot(oni,
     style = "red_blue")
```

```{asis, echo = more_details}
## Further plotting options and styles

To see `oni` as an annual (not monthly index):
```

```{r oniannual, echo = more_details, eval = more_details}
plot(oni,
     smooth_over_year = TRUE,
     lwd = 6)
```

## Further plotting options and styles

Say you want to see if specific events coincide with El Niño (based on a
true story, see help for details):
```{r onievent}
plot(oni,
     event_years = c(1996, 2003, 2004, 2006, 2007, 2008, 2009, 2010, 2014, 2016, 2017, 2019),
     xlim = c(lubridate::dmy(01011995),
              lubridate::dmy(01012024)),
     lwd = 2)
```

```{asis, echo = more_details}
## Technical aside regarding `plot()`

The object `oni` has the class `pacea_index`:  \
```

```{asis, echo = more_details, eval = long_talk}
class(oni)
```

```{asis, echo = more_details}
  \
which ensures that `plot(oni)`
automatically uses our specialised function `plot.pacea_index(oni)`, giving
the red and blue colours, tickmarks, axis labelling, etc.

  \

So see `?plot.pacea_index` for further options. Similarly, check the `class` of
other objects saved in `pacea`.
```
