# Indices

## Climatic and oceanographic indices

Various climate and oceanographic indices are currently included in pacea:

```{r indicestable, echo=FALSE, eval=FALSE}
# Used to create tex that is copied below and red added to; just manually check
#  pacea_indices to update
kableExtra::kable(pacea_indices, "latex", booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "striped"))
```

\begin{table}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}{lllrr}
\toprule
Object & Description & Resolution & Start year & End year\\
\midrule
\cellcolor{gray!10}{pdo} & \cellcolor{gray!10}{Pacific Decadal Oscillation} & \cellcolor{gray!10}{monthly} & \cellcolor{gray!10}{1854} & \cellcolor{gray!10}{\textcolor{red}{2025}}\\
npi\_monthly & North Pacific Index (monthly) & monthly & 1899 & \textcolor{red}{2024}\\
\cellcolor{gray!10}{npi\_annual} & \cellcolor{gray!10}{North Pacific Index (annual)} & \cellcolor{gray!10}{annual} & \cellcolor{gray!10}{1899} & \cellcolor{gray!10}{\textcolor{red}{2024}}\\
alpi & Aleutian Low Pressure Index & annual & 1900 & \textcolor{red}{2022}\\
\cellcolor{gray!10}{oni} & \cellcolor{gray!10}{Oceanic Niño Index} & \cellcolor{gray!10}{monthly} & \cellcolor{gray!10}{1950} & \cellcolor{gray!10}{\textcolor{red}{2024}}\\
\addlinespace
npgo & North Pacific Gyre Oscillation & monthly & 1950 & \textcolor{red}{2025}\\
\cellcolor{gray!10}{ao} & \cellcolor{gray!10}{Arctic Oscillation} & \cellcolor{gray!10}{monthly} & \cellcolor{gray!10}{1950} & \cellcolor{gray!10}{\textcolor{red}{2025}}\\
soi & Southern Oscillation Index & monthly & 1951 & \textcolor{red}{2025}\\
\cellcolor{gray!10}{\textcolor{red}{bi}} & \cellcolor{gray!10}{\textcolor{red}{North Pacific Bifurcation Index}} & \cellcolor{gray!10}{\textcolor{red}{annual}} & \cellcolor{gray!10}{\textcolor{red}{1967}} & \cellcolor{gray!10}{\textcolor{red}{2024}}\\
mei & Multivariate El Niño Southern Oscillation Index & monthly & 1979 & \textcolor{red}{2025}\\
\bottomrule
\end{tabular}}
\end{table}

<!--
## New index

Now includes Mike Malick's North Pacific Current Bifurcation Index:

```{r biplot}
plot(bi, lwd = 6)
```
-->
