# Introduction

## Acknowledgments

- `r colorize("Co-authors: Joe Watson, Angelica Peña, Andrea Hilborn, Charles
  Hannah, and Chris Rooper.", "blue")`

- Kelsey Flynn, Jessica Nephin, Lindsay Davidson, Strahan Tucker,
  Brianna Wright, Patrick Thompson, Matt Grinnell, Sean Anderson, Philina
  English, Chris Grandin, Jennifer Boldt, and others.

- `r colorize("Carley Colclough for designing the logo.", "blue")`

- DFO's Competitive Science Research Fund for funding (project 21-FS-03-13).

```{asis, echo = explain_assessment}
## What is a stock assessment?

- Collect data from surveys and commercial fisheries, including ages of fish
- Data inform statistical population models to estimate historical biomass
- Models provide advice to fisheries managers

Pacific Hake assessment from 2023, estimate the historical population size:
```

```{r hakeintro, echo = FALSE, eval = explain_assessment}
plot(hake_biomass,
     y_max = 3.5)
```

```{asis, echo = explain_assessment}
## What is a stock assessment?

Advice often in the form of 'decision tables' that indicate probabilities of future events given
different catches:

| Catch (t) in 2023 | Probability spawning biomass declines from 2023 to 2024 | Probability spawning biomass in 2024 falls below 40% of unfished biomass |
| ----------------: | ------------------------------------------------------: | -----------------------------------------------------------------------: |
| 0 | 50% | 2% |
| 180,000 | 72% | 3% |
| 225,000 | 75% | 3% |
| 320,000 | 78% | 3% |
| 430,000 | 85% | 5% |
```

## Motivation

- Revised Fisheries Act: “... the Minister shall take into account the environmental conditions affecting a fish stock."

- `r colorize("Yet <50\\% of DFO’s stock assessments currently use environmental data.", "blue")`

- Only 28% of assessments in Pacific Region use environmental data.

- `r colorize("Leading cause of not using environmental data is availability of the data.", "red")`

  \

\footnotesize

Kulka DW, Thompson S, Cogliati K, Olmstead M, Austin D, Pepin D. (2022). An Accounting of Integration of Environmental Variables in Fishery Stock Assessments in Canada. Can. Tech. Rep. Fish. Aquat. Sci. 3473: viii + 79 p.
https://publications.gc.ca/collections/collection_2022/mpo-dfo/Fs97-6-3473-eng.pdf

\normalsize

##
```{r motivation1, echo=FALSE, out.width = "100%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-1.png"))
```

##
```{r motivation2, echo=FALSE, out.width = "100%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-2.png"))
```

## Motivation (based on a true story)

A search for sea surface temperature yields an overwhelming number (341) of choices:

```{r motivation4, echo=FALSE, out.width = "90%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation3.png"))
```

Likely requires extensive data wrangling to be usable, which usually takes way,
way, longer than anticipated.

So the SST analysis did not happen.

## Motivation

- Motivated by Dan Duplisea's `gslea` package for the Gulf of St. Lawrence.

- Outputs from a BC physical biogeochemical model have been shared by Angelica Peña.

- But the netCDF format may be unfamiliar to non-oceanographers --
  we take care of the wrangling into R

- Open Data is great, but can be hard to
  `r colorize("convert raw data into usable information", "red")`.

- Primary audience is DFO stock assessment scientists, but usable by anyone
  (with a minimal working knowledge of R).

  \

\footnotesize

Duplisea et al. (2020). gslea: the Gulf of St Lawrence ecosystem
approach data matrix R-package. R package version 0.1 https://github.com/duplisea/gslea

\normalsize

## Why an R package?

- R is the computer programming language most widely used by stock assessment scientists.

- An R package is the standard way to share code (and data).

- Easy installation:

\textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}


- Ensures proper documentation of data objects and functions (helps users).

- Include vignettes that walk users through various features.

- All data (except one type) is saved within the package, no further downloading
  required, so not relying on external websites being functional.

## Why an R package on GitHub?

- Can host freely on GitHub: https://github.com/pbs-assess/pacea

- Standard way to collaborate on code.

- Code is completely open source: transparent, traceable, and transferable.

```{r whyR, echo=FALSE, out.width="40%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/github-screenshot-small.png"))
```

- 'Passing' badge: ensures package can build in R.

- 'codecov' is the amount of the code that is covered by unit tests, which help weed
  out bugs.

## Spatial domain

Started thinking about ecosystem assessment regions like in gslea:

```{r gslea, echo=FALSE, out.width="50%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/gslmap.plain-1.png"))
```

But it became non-trivial to decide on regions, and we decided to explicitly include
spatial data.

## `pacea` objects

Currently, pacea contains:

- `r format(nrow(buoy_sst), big.mark = ",")` calculations of daily sea surface
   temperature based on data from `r nrow(buoy_metadata)` buoys

- outputs from the spatial British Columbia continental margin (BCCM) model

- NOAA's spatial Optimum Interpolation Sea Surface Temperature (OISST) record

- 9 climatic and oceanographic indices, such as the Pacific Decadal Oscillation
   and those related to El Niño

- estimates of abundances for Harbour Seals and Pacific Hake

- zooplankton biomass anomalies for the Strait of Georgia
