# Motivation

##

```{r motivation1, echo=FALSE, out.width = "100%", eval = long_motivation}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-1.png"))
```

##

```{r motivation2, echo=FALSE, out.width = "100%", eval = long_motivation}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-2.png"))
```

## Motivation (based on a true story)

A search for sea surface temperature yields an overwhelming number (341) of choices:

```{r motivation4, echo=FALSE, out.width = "90%", eval = long_motivation}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation3.png"))
```

\pause

Likely requires extensive data wrangling to be usable, which usually takes way,
way, longer than anticipated.

So the SST analysis did not happen.

## Motivation

- Revised Fisheries Act: “... the Minister shall take into account the environmental conditions affecting a fish stock."

- `r colorize("Yet <50\\% of DFO’s stock assessments currently use environmental data.", "blue")`

- Only 28% of assessments in Pacific Region.

- `r colorize("Leading cause of not using environmental data is availability of the data.", "red")`

\pause

- Open Data is great, but can be hard to
  `r colorize("convert raw data into usable information", "red")`.

- Primary audience is DFO stock assessment scientists, but usable by anyone
  (with a minimal working knowledge of R).

<!--
\

\footnotesize

Kulka DW, Thompson S, Cogliati K, Olmstead M, Austin D, Pepin D. (2022). An Accounting of Integration of Environmental Variables in Fishery Stock Assessments in Canada. Can. Tech. Rep. Fish. Aquat. Sci. 3473: viii + 79 p.
https://publications.gc.ca/collections/collection_2022/mpo-dfo/Fs97-6-3473-eng.pdf

\normalsize
-->


<!--
## Motivation

- Initial motivation was Dan Duplisea's `gslea` package for the Gulf of St. Lawrence.

- Outputs from a BC physical biogeochemical model have been shared by Angelica Peña.

- But the netCDF format may be unfamiliar to non-oceanographers --
  we take care of the wrangling into R.

- Open Data is great, but can be hard to
  `r colorize("convert raw data into usable information", "red")`.

- Primary audience is DFO stock assessment scientists, but usable by anyone
  (with a minimal working knowledge of R).

  \

\footnotesize

Duplisea et al. (2020). gslea: the Gulf of St Lawrence ecosystem
approach data matrix R-package. R package version 0.1 https://github.com/duplisea/gslea

\normalsize
-->

<!-- just manually comment out if don't want, trying to automate was getting a -->
<!-- bit too clever, especially with cacheing -->

<!--
## Motivation

- push to include ecosystem information in advice to fisheries managers

- availability of data and model output is the leading impediment
  - where to find it, wrangling it (e.g netCDF into R), understanding it
  - Open Data is great, but have to convert raw data into usable information

- Primary audience is DFO stock assessment scientists, but usable by anyone
  (with a minimal working knowledge of R).

## What is a stock assessment?

- We collect data from surveys and commercial fisheries, including ages of fish
- Data inform statistical population models to estimate historical biomass
\pause
- Models provide advice to fisheries managers

Pacific Hake assessment from 2025, estimate the historical population size:

```{r hakeintro, echo = FALSE, eval = explain_assessment}
plot(hake_biomass,
     y_max = 3.5)
```

## What is a stock assessment?

Advice often in the form of 'decision tables' that indicate probabilities of future events given
different catches:

| Catch (t) in 2025 | Probability spawning biomass declines from 2025 to 2026 | Probability spawning biomass in 2026 falls below 40% of unfished biomass |
| ----------------: | ------------------------------------------------------: | -----------------------------------------------------------------------: |
| 0 | 89% | 11% |
| 150,000 | 100% | 17% |
| 300,000 | 100% | 23% |
| 450,000 | 100% | 30% |

-->

## Hence... pacea

- Released in November 2023.

- Maintained and extended since then.

- Include vignettes that walk users through various features.

- All data (except oceanographic model output) is saved within the package
  - no further downloading required
  - not relying on external websites being functional.

- Code is completely open source: transparent, traceable, and transferable.

## pacea palooza 2025

- Given quite a few talks, but not done a detailed how-to presentation

- Gave short talk to Molecular Genetics Section

- Claire thought of hosting a workshop

**Day 1**

- motivation and overview (this talk)

- overview of GitHub site

- we will walk through example code of each type of data and model output

  - based on the vignettes that are rendered on GitHub 
  - so you do **not** need to frantically copy everything down that we type
  
- options:
  - people play with relevant outputs on their own?
  - look at people's questions? work through one together?
  - troubleshoot any installation issues

## pacea palooza 2025

**Between Day 1 and Day 2**

- carry on digging into the package
- wrangle any of your own data that you want to use with pacea outputs to 
   investigate a problem
- try some analyses

**Day 2**

- people say what they have done
- any problems we can work through together 
- people could plug away individually (or in pairs/groups, particularly if some people
  are less familiar with R) and we help out
- whatever else people want


<!--
## Why an R package?

- R is the computer programming language most widely used by stock assessment scientists.

- An R package is the standard way to share code (and data).

- Easy installation:

\textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}


- Ensures proper documentation of data objects and functions (helps users).

- Include vignettes that walk users through various features.

- All data (except oceanographic model output) is saved within the package, no further downloading
  required, so not relying on external websites being functional.
-->

<!--
## Why an R package on GitHub?

- R is the computer programming language most widely used by stock assessment scientists.

- Can host freely on GitHub: https://github.com/pbs-assess/pacea

- Easy installation:

\textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}

- Code is completely open source: transparent, traceable, and transferable.

```{r whyR, echo=FALSE, out.width="40%", eval = long_motivation}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/github-screenshot-small.png"))
```
-->

```{r, eval = FALSE, echo = FALSE}
- 'Passing' badge: ensures package can build in R.

- 'codecov' is the amount of the code that is covered by unit tests, which help weed
  out bugs.

## Spatial domain

Started thinking about ecosystem assessment regions like in gslea:
```

```{r gslea, echo=FALSE, out.width="50%", eval = FALSE}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/gslmap.plain-1.png"))
```

```{r, eval = FALSE, echo = FALSE}
But it became non-trivial to decide on regions, and we decided to explicitly include
spatial information, especially regarding outputs of the biogeochemical model.
```
<!--

## `pacea` objects (climatic/oceanographic)

Currently, pacea contains (`r colorize("red is updated/new since last year's meeting", "red")`):

- `r colorize(format(nrow(buoy_sst), big.mark = ","), "red")` calculations of daily sea surface
   temperature based on data from `r nrow(buoy_metadata)` buoys

- outputs from the spatial British Columbia continental margin (BCCM) model,
   `r colorize("now expanded from just Canada's EEZ to include the full model domain,
   down to Columbia River", "red")`

- `r colorize("outputs from the spatial Hindcast of the Salish Sea (HOTSSea)
  physical oceanography model for the Salish Sea, 1980-2018", "red")`

- NOAA's spatial Optimum Interpolation Sea Surface Temperature (OISST) record
  `r colorize("updated monthly", "red")`

- $\cancel{9}$ `r colorize(nrow(pacea_indices), "red")` climatic and oceanographic indices, such as the Pacific Decadal Oscillation and those related to El Niño

## `pacea` objects (biological)

- estimates of abundances for Harbour Seals

- estimates of spawning stock biomass and annual recruitments for
 `r colorize("Pacific Hake", "red")` and `r colorize("Pacific Herring", "red")`

- `r colorize("zooplankton biomass anomalies in the Strait of Georgia for 25 species groups", "red")`
-->

