# Introduction

## Backgound and motivation

Copy things from existing powerpoint.

## ideas

describe the data currently contained, demonstrate the plotting functions and how to access the data, and mention caveats, assumptions, and problems encountered (and solved, so you don't have to!) when wrangling certain data sets.



##
```{r motivation1, echo=FALSE, out.width = "100%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-1.png"))
```

##
```{r motivation2, echo=FALSE, out.width = "100%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation-2.png"))
```

## Motivation (based on a true story)

A search for sea surface temperature yields an overwhelming number (341) of choices:

```{r motivation4, echo=FALSE, out.width = "90%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/motivation3.png"))
```

Likely requires extensive data wrangling to be usable, which usually takes way,
way, longer than anticipated.

So the SST analysis did not happen.

## Guidelines for users

Read the README!!

Check the NEWS!!

We have built comprehensive vignettes and help files for each type of data object in `pacea`.

Please go through them as this will be how to really understand the details.

For any simple questions just contact us. For detailed questions or bug reports
please post an `Issue` on GitHub and a minimum working example:

 - often helps you figure out the answer
 - using `Issues` allow you to first check if we are already working on it
 - allows others to see (and maybe help answer) questions being asked

We will describe updates to `pacea` in the NEWS, such as updating certain data
sets and adding new ones.


## `pacea` objects

Currently, pacea contains:

- `r format(nrow(buoy_sst), big.mark = ",")` calculations of daily sea surface
   temperature based on data from `r nrow(buoy_metadata)` buoys

- outputs from the spatial British Columbia continental margin (BCCM) model

- NOAA's spatial Optimum Interpolation Sea Surface Temperature (OISST) record

- 9 climatic and oceanographic indices, such as the Pacific Decadal Oscillation
   and those related to El Niño

- estimates of abundances for Harbour Seals and Pacific Hake.
