# OISST

## NOAA's Optimal Interpolation Sea Surface Temperature

OISST - long term gridded climate data record

- 1/4° longitude x latitude

- Interpolated to fill gaps

- Canada's Pacific EEZ

- `r colorize("925 gridded cells", "red")`

- `r colorize("September, 1981 to (relatively) present", "red")`
  - Updated ~monthly

\
2 data products: weekly and monthly mean SST

- `oisst_7day` and `oisst_month`

\
\footnotesize
R `erddap` - https://www.ncei.noaa.gov/products/optimum-interpolation-sst

Huang, B., C. Liu, V. Banzon, E. Freeman, G. Graham, B. Hankins, T. Smith, and H. Zhang, 2021: Improvements of the Daily Optimum Interpolation Sea Surface Temperature (DOISST) Version 2.1. J. Climate, 34, 2923–2939, https://doi.org/10.1175/JCLI-D-20-0166.1.
\normalsize

## OISST -- monthly mean
The data are saved as a `sf` (simple feature) `tibble` object in `pacea`:

```{r oisstdata, eval=FALSE}
oisst_month
```

```{r oisstdata2, eval=FALSE}
Simple feature collection with 467125 features and 7 fields
Geometry type: POINT
Dimension:     XY
Bounding box:  xmin: -138.625 ymin: 46.625 xmax: -123.125 ymax: 54.625
Geodetic CRS:  WGS 84
# A tibble: 467,125 x 8
   year month   sst sst_sd sst_n start_date end_date
* <dbl> <dbl> <dbl>  <dbl> <int> <date>     <date>
1  1981     9  16.4  0.830    30 1981-09-01 1981-09-30
2  1981    10  14.3  0.838    31 1981-10-01 1981-10-31
3  1981    11  12.7  1.38     30 1981-11-01 1981-11-30
4  1981    12  10.2  0.549    31 1981-12-01 1981-12-31
5  1981     9  16.4  1.05     30 1981-09-01 1981-09-30
6  1981    10  14.2  0.697    31 1981-10-01 1981-10-31
# i 467,119 more rows
# i 1 more variable: geometry <POINT [B0]>
# i Use `print(n = ...)` to see more rows
```

```{asis, echo = more_details}
## Help file
```
```{r oissthelp, eval = FALSE, echo = more_details}
?oisst_month
```

```{r oissthelp2, eval = FALSE, echo = more_details}
Description
  Simple features objects of NOAAs Optimal Interpolation SST

oisst_7day
  7-day mean sea surface temperature (deg C)

oisst_month
  Monthly mean sea surface temperature in (deg C)

Usage
  oisst_7day
  oisst_month

Format
  year    Year

  week    Calendar week

  month   Calendar month

  sst     Mean sea surface temperature, SST (deg C)
...
```

## Built in plots -- default
```{r oisstplot}
plot(oisst_month)
```

## Built in plots -- select years and months
```{r oisstplot custom}
plot(oisst_month, months.plot = c("April", "June"), years.plot = c(2021:2022))
```
