# Populations

## Estimates of fish and seal populations

General intro slide, then show code on next slide. In pacea we include estimates of fish populations, including uncertainty, as calculated from recent
stock assessments (have general slide earlier)..

Currently, we just have estimates of:

Pacific Hake, a major component of the ecosystem, with more
to come. blah blah blah

## Hake biomass

The biomass time series is saved as a `tibble` object in `pacea`:
```{r hakebiomass}
hake_biomass

tail(hake_biomass)
```

## Hake biomass
```{r hakebiomasshelp, eval=FALSE}
?hake_biomass

Pacific Hake annual spawning stock biomass (females only) as estimated
by the 2023 stock assessment.

Description:

     The Pacific Hake stock is managed and assessed through an
     Agreement between Canada and the United States. The spawning stock
     biomass (mature females) estimates come from the most recent joint
     stock assessment, and, importantly, are for the coastwide stock
     from California to British Columbia.

Usage:

     hake_biomass

Format:

     A tibble also of class `pacea_biomass` with columns:

     year: year of the estimate of spawning stock biomass (mature
          females), where the estimate is for the start of the year

     low: low end (2.5th percentile) of the 95% credible interval for
          biomass, units as for `median`

     median: median estimate of biomass, in millions of tonnes of
          female spawning biomass at the start of the year

     ...
```


## Hake biomass

Objects have tailored plotting functions associated with them:
```{r hakebiomassplot}
plot(hake_biomass)
```

## Hake recruitment

```{r hakerecuitmentplot}
plot(hake_recruitment)
```


## blah blah

For example, it is important to
realise that the assessment considers a coastwide stock off the west coast of
Canada and the United States.

To plot the time series of estimated coastwide biomass,
simply do


where the solid circles are the annual medians, with credible intervals given by
the shaded regions (mostly reproducing Figure d in the latest assessment).

Note that the `hake_biomass` object also has class `pacea_biomass`:
```{r hakebiomassclass}
class(hake_biomass)
```
such that the above command `plot(hake_biomass)` first looks to use our tailored function
`plot.pacea_biomass()`, which gives the style of plot shown above. We use this
approach for `hake_recruitment` and various other objects in `pacea`, to enable
easy default plotting.

The `hake_recruitment` object contains annual estimates of recruitment of age-0
fish
```{r hakerec}
hake_recruitment
tail(hake_recruitment)
```
for which values are billions of age-0 fish, and our default plot is
```{r hakerecplot}
plot(hake_recruitment)
```

The dots represent the medians, with blue bars showing the 95% credible intervals.
This reproduces Figure f in the latest stock assessment (without some extra
details).

Note that, as mentioned in `?hake_recruitment` the most recent few years are
very uncertain as they are not informed by data. In particular, the final year
represents recruitment at the start of the year but has no data to inform it,
and so should definitely not be used for any analyses (we include it as it is
shown in the stock assessment; it also essentially represents the prior distribution used for recruitment).

The estimated uncertainty of recruitments is large, and so we also include
scaled values that make it easier to compare years. The 2010 recruitment is the
second largest in the time series (below only 1980's) and as such people have an
intuition that it is large, and we can compare other years by looking at
recruitment scaled by the 2010 recruitment:
```{r hake2010}
hake_recruitment_over_2010
tail(hake_recruitment_over_2010)
plot(hake_recruitment_over_2010)
```

This shows the medians and 95% credible intervals (red lines) of the recruitment
divided by that in 2010, which clearly shows that, for example, the 2014 is
clearly smaller than the 2010 recruitment (highlighted by the green dashed
line), which is not intuitive in the above `plot(hake_recruitment)` plot (see
`?hake_recruitment_over_2010` for more details and motivation).

This approach
somewhat scales out some of the uncertainty in the absolute scale of $R_0$, the
mean unfished equilibrium recruitment. This is similarly seen by instead scaling
recruitments by $R_0$, as
```{r hakeR0}
hake_recruitment_over_R0
tail(hake_recruitment_over_R0)
plot(hake_recruitment_over_R0)
```

Note that the Pacific Hake stock assessment is conducted annually at the start of each year,
and these estimates will be updated in `pacea` probably around March. So
estimates will change (particularly for the most recent years) as more data
inform the stock assessment model.

We anticipate adding assessment output for other species soon,
so stay tuned.
