# Populations

## Estimates of animal populations

In `pacea` we include estimates of populations as
calculated from recent stock assessments or analyses.

Currently, we have:

- Pacific Hake spawning biomass and recruitments for the coastwide stock off Canada and United States

- Pacific Harbour Seals abundance (the most abundant pinniped species in the Northeast Pacific)

- Pacific Herring spawning biomass and recruitments for all five stock
  assessment regions

- Zooplankton biomass anomalies for 24 species groups in the Strait of Georgia


## Hake biomass

The hake biomass time series is saved as a `tibble` object in `pacea`:
```{r hakebiomass}
hake_biomass

tail(hake_biomass)
```
```{asis, echo = long_talk}
## Hake biomass
```
```{r hakebiomasshelp, eval=FALSE, echo=long_talk}
?hake_biomass

Pacific Hake annual spawning stock biomass (females only) as estimated
by the 2024 stock assessment.

Description:

     The Pacific Hake stock is managed and assessed through an
     Agreement between Canada and the United States. The spawning stock
     biomass (mature females) estimates come from the most recent joint
     stock assessment, and, importantly, are for the coastwide stock
     from California to British Columbia.

Usage:

     hake_biomass

Format:

     A tibble also of class `pacea_biomass` with columns:

     year: year of the estimate of spawning stock biomass (mature
          females), where the estimate is for the start of the year

     low: low end (2.5th percentile) of the 95% credible interval for
          biomass, units as for `median`

     median: median estimate of biomass, in millions of tonnes of
          female spawning biomass at the start of the year

     ...
```


## Hake biomass

Time series of spawning biomass:
```{r hakebiomassplot}
plot(hake_biomass)
```

## Hake recruitment

Estimates of annual recruitment (exclude recent years for any analyses since not
greatly informed by data; see `?hake_recruitment`):
```{r hakerecuitmentplot}
plot(hake_recruitment)
```

## Herring biomass

The herring spawning biomass time series is saved similarly to hake, but
contains values for five regions:
```{r herringbiomass}
herring_spawning_biomass

tail(herring_spawning_biomass)
```

## Herring biomass

Estimates of spawning biomass for one region (HG = Haida Gwaii):
```{r herringbiomassplot}
plot(herring_spawning_biomass,
     region = "HG")
```

## Herring recruitment

Estimates of annual age-2 recruitment for one region (HG = Haida Gwaii):
```{r herringrecuitmentplot}
plot(herring_recruitment,
     region = "HG")
```

## Pacific Harbour Seals

Estimated abundances for seven regions (and coastwide) were calculated by DFO
(2022; SAR 2022/034), and included here:
```{r seals}
harbour_seals

tail(harbour_seals)
```

## Pacific Harbour Seals

Estimates for each region and coastwide, reproducing Figure 3 of DFO (2022):
```{r sealplot}
plot(harbour_seals)
```

```{asis, echo = long_talk}
## Pacific Harbour Seals

Or just plot estimates for a single region:
```
```{r sealplot2, echo = long_talk, eval = long_talk}
plot(harbour_seals,
     region = "SOG")
```

## Zooplankton in the Strait of Georgia

Biomass anomalies of zooplankton biomass from 1996 onwards from

Perry et al. (2021), as extended by Kelly Young each year for DFO's

State of the Pacific Ocean Report.
```{r zooplankton}
zooplankton_sog
```

## Zooplankton in the Strait of Georgia

Default plot of anomalies of total biomass:
```{r zooplot}
plot(zooplankton_sog)
```

```{asis, echo = long_talk}
## Zooplankton in the Strait of Georgia

Or plot anomalies for a specific species group:
```
```{r zooplot2, echo = long_talk, eval = long_talk}
plot(zooplankton_sog,
     species_group = "cladocera")
```
