# Populations

## Estimates of animal populations

In `pacea` we include estimates of populations, with measures of uncertainty, as
calculated from recent stock assessments.

Currently, we have estimates of:

- Pacific Hake (the coastwide stock off Canada and United States)

- Pacific Harbour Seals (most abundant pinniped species in the Northeast Pacific)

- Pacific Herring (coming soon)

## Hake biomass

The hake biomass time series is saved as a `tibble` object in `pacea`:
```{r hakebiomass}
hake_biomass

tail(hake_biomass)
```

## Hake biomass
```{r hakebiomasshelp, eval=FALSE}
?hake_biomass

Pacific Hake annual spawning stock biomass (females only) as estimated
by the 2023 stock assessment.

Description:

     The Pacific Hake stock is managed and assessed through an
     Agreement between Canada and the United States. The spawning stock
     biomass (mature females) estimates come from the most recent joint
     stock assessment, and, importantly, are for the coastwide stock
     from California to British Columbia.

Usage:

     hake_biomass

Format:

     A tibble also of class `pacea_biomass` with columns:

     year: year of the estimate of spawning stock biomass (mature
          females), where the estimate is for the start of the year

     low: low end (2.5th percentile) of the 95% credible interval for
          biomass, units as for `median`

     median: median estimate of biomass, in millions of tonnes of
          female spawning biomass at the start of the year

     ...
```


## Hake biomass

Time series of spawning biomass:
```{r hakebiomassplot}
plot(hake_biomass)
```

## Hake recruitment

Estimates of annual recruitment (exclude recent years for any analyses since not
greatly informed by data; see `?hake_recruitment`):
```{r hakerecuitmentplot}
plot(hake_recruitment)
```

## Pacific Harbour Seals

Estimated abundances for seven regions (and coastwide) were calculated by DFO
(2022; SAR 2022/034), and included here:
```{r seals}
harbour_seals

tail(harbour_seals)
```

## Pacific Harbour Seals

Estimates for each region and coastwide, reproducing Figure 3 of DFO (2022):
```{r sealplot}
plot(harbour_seals)
```

## Pacific Harbour Seals

Or just plot estimates for a single region:
```{r sealplot2}
plot(harbour_seals,
     region = "SOG")
```
