<!--
## Other updates

- $\cancel{204,373}$ `r colorize(format(nrow(buoy_sst), big.mark = ","), "red")` calculations of daily sea surface
   temperature based on data from `r nrow(buoy_metadata)` buoys

- NOAA's spatial Optimum Interpolation Sea Surface Temperature (OISST) record
  `r colorize("updated monthly", "red")`

\pause

- `r colorize("expanded", "red")`  British Columbia continental margin (BCCM) model outputs
   `r colorize("to include the full model domain and 2x2 km resolution", "red")`


- `r colorize("bathymetry data", "red")`

- estimates of abundances for Harbour Seals (will be updated from next assessment)

- code for `r colorize("3d printing", "red")` of biomass-at-age
-->

# Uses of pacea

## Ecosystem summaries

Stock-specific ecosystem summary -- [Edwards et al. (2026)](https://www.fisheries.noaa.gov/s3/2026-02/hake-assessment-2026.pdf)
```{r, ecosumplot, echo = FALSE}
source(paste0(here::here(),
              "/report/ecosystem-summary-hake/ecosystem-summary-hake.R"))
# png(paste0(here::here(), "/talks/talks-manual-figures/ecosystem-summary-hake.png"),
#    width = 400,   # 670 px matches size from running from a chunk
#    height = 360)  # 480

#ecosystem_summary_hake()
#dev.off()
```

```{r, ecosumfunc, echo = TRUE, eval = FALSE}
ecosystem_summary_hake()
```
```{r, ecosum, echo = FALSE, fig.height = 4.2}
res <- ecosystem_summary_hake(short_talk_version = TRUE)
```

## Ongoing uses

Stock assessments:

- Petrale Sole
- Dover Sole
- Yelloweye Rockfish

Others:

- body conditions of 35 fish species (English et al., 2026)
- spatial densities of humpback whales
- fluctuations of infectious hematopoietic necrosis virus (IHPV) in Sockeye Salmon
- Salmon Prize prediction project

# New products

## Coming soon

- BCCM model updated to 2025

- Functionality to extract GLORYS (Global Ocean Physics Reanalysis) outputs, such
  as mixed-layer depths and eddy kinetic energy 
  
- Future projections downscaled from IPCC forecasts using Angelica's BCCM model
  - Surface and bottom temperatures
  - Three future socio-economic pathways
<!-- - Done by Trond Kristiansen et al., contract through Robyn Forrest -->

\pause

<!-- Three future socio-economic pathways:

- SSP1-RCP2.6 -- `r colorize("sustainable development", "red")`
- SSP2-RCP4.5 -- `r colorize("middle-of-the-road development", "red")`
- SSP5-RCP8.5 -- `r colorize("fossil fueled development", "red")`

\pause

Think we will have monthly averages, corresponding to time periods:

- 2021-2040
- 2041-2060
- 2080-2090
- or maybe just 2050 and 2090

-->

## Potentially coming soon

- Pan-basin Pattern (due to breakdown of PDO-ecosystem correlations)
- Relative Oceanic Niño Index (RONI)
- Lightstation data (temperature and salinity)
- Functionality to extract the oceanographic indices (Python or just .csv)
- Oxygen time series in BC waters

<!-- 
```{r otherinfo, echo=FALSE, out.width = "100%"}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/pacea-montage.png"))
```


# Installation

 - https://github.com/pbs-assess/pacea

 - can read vignettes online

 <!-- - installation:
   \textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}  -->

\medskip

\medskip

\begin{center}
\includegraphics[width = 2cm]{"../man/figures/logo.png"}\\
\emph{We wrangle the data so you don't have to}
\end{center}

<!-- 
```{r citation, eval=FALSE}
citation("pacea")

Edwards AM, Tai TC, Watson J, Peña MA, Hilborn A, Hannah CG, Rooper CN, Flynn
  KL, Oldford GL (2024). "pacea: An R package of Pacific ecosystem
  information to help facilitate an ecosystem approach to fisheries
  management."
  https://github.com/pbs-assess/pacea.
```
-->
