# Summary

## Commitments from package authors

We will:

- \tiny (try to) \normalsize not break the package, by working on branches on
  GitHub before merging.

- ensure any improvements are back compatible.

- update the NEWS as we add/update data and make improvements.

- ideally update the following `r colorize("on the 20th of every month (or soon after)", "red")` to get the latest raw data:
  - climatic and oceanographic indices
  - buoy sst
  - OISST

## Commitments from package authors

- update other data as appropriate:
  - hake assessment output: February each year
  - harbour seals: next time assessment is updated
  - herring assessment output: when SAR published each year
  - BCCM model output: the model may be run into 2020 (boundary
    conditions come from a global model currently only available to Jan 2021).
    - note that BCCM is a research model (not an operational model) so updates
      depend on projects and access to High Performance Computing.

## Recommended guidelines for users

Please:

 - read the README!!

 - check the help files and vignettes

 - check the NEWS (for updates to data, useful additions, new data, etc.)

 - understand the limitations of the data

 - cite `pacea` and the original sources for any data you use

```{asis, echo = more_details}
## Recommended guidelines for users

To check when you last installed `pacea`
```

```{r installed, echo = more_details, eval = more_details}
pacea_installed()

You last locally installed pacea on 2023-11-01 which is 8 days ago.

Compare that with the updates in the NEWS file (and you can also check dates of the latest commits at
https://github.com/pbs-assess/pacea/commits/main)
```

```{r installedscreenshot, echo=FALSE, out.width="30%", eval = more_details}
knitr::include_graphics(paste0(here::here(),
                               "/talks/talks-manual-figures/github-commits.png"))
```

```{asis, echo = more_details}
## Recommended guidelines for users

 - any problems, make a minimum working example (usually makes you figure it out yourself)
   - check the `Issues` on GitHub in case someone else has asked or we are
     already working on it

   - simple quick question just ask us

   - more detailed questions or bug reports post
     a new `Issue` (allows others to see, and maybe answer, your question)
```

## Current uses of pacea (that we know about)

 - Spiny Dogfish: understanding declines over the past 20 years

 - shrimp: impact of including environmental variables on predicting distributions

 - humpback whales and porpoises: environmental covariates in species distribution models

 - Pacific Saury assessment: finding links between two basin-scale indices and process errors, directly supporting Canada's North Pacific Fisheries
   Commission commitments

## Lessons learned (applicable to other regions)

 - is possible to amaglamate different ecosystem information into one place

 - not an off-the-side-of-your-desk project

 - is being used by DFO scientists (and others)

 - infrastructure of the package could be utilised in other regions if
   interested
   - e.g. my monthly download of buoy data is for all Canadian waters,
   the data processing, storing, and plotting functions could be adopted
   - similarly for other types of data/information

## Summary

<!-- - let us know of any applications (so we understand the uptake and how much
   effort to devote to maintaining it)

   - you could be out on a survey and wondering how warm the water is
        compared to previous years

    - Chris Rooper: looking at Pacific Saury process errors from a stock assessment
 using the indices (PDO etc.). ``It made my life easier today."
-->
 - https://github.com/pbs-assess/pacea

 - installation:
   \textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}

\

\begin{center}
\includegraphics[width = 2cm]{"../man/figures/logo.png"}\\
\emph{We wrangle the data so you don't have to}
\end{center}

\

```{r citation, eval=FALSE}
citation("pacea")

Edwards AM, Tai TC, Watson J, Peña MA, Hilborn A, Hannah CG, Rooper CN
(2023). "pacea: An R package of Pacific ecosystem information to help facilitate an ecosystem approach to fisheries management."
  https://github.com/pbs-assess/pacea.
```
