# Summary

## Other updates

- $\cancel{204,373}$ `r colorize(format(nrow(buoy_sst), big.mark = ","), "red")` calculations of daily sea surface
   temperature based on data from `r nrow(buoy_metadata)` buoys

\pause

- outputs from the spatial British Columbia continental margin (BCCM) model,
   `r colorize("now expanded from just Canada's EEZ to include the full model domain,
   extending into US waters", "red")`

- NOAA's spatial Optimum Interpolation Sea Surface Temperature (OISST) record
  `r colorize("updated monthly", "red")`

- estimates of abundances for Harbour Seals (will be updated from next assessment)


## Ecosystem summaries

Idea is to head towards stock-specific functions like (Johnson et al., 2025):

```{r, ecosumplot, echo = FALSE}
# Before moving into R/
source(paste0(here::here(),
              "/R-temp/ecosystem-summary-hake.R"))
# png(paste0(here::here(), "/talks/talks-manual-figures/ecosystem-summary-hake.png"),
#    width = 400,   # 670 px matches size from running from a chunk
#    height = 360)  # 480

#ecosystem_summary_hake()
#dev.off()
```

```{r, ecosum, fig.height = 4.2}
ecosystem_summary_hake()
```

## Installation

 - https://github.com/pbs-assess/pacea

 - can read vignettes online

 <!-- - installation:
   \textcolor{red}{\texttt{remotes::install\_github("pbs-assess/pacea")}}  -->

\begin{center}
\includegraphics[width = 2cm]{"../man/figures/logo.png"}\\
\emph{We wrangle the data so you don't have to}
\end{center}

```{r citation, eval=FALSE}
citation("pacea")

Edwards AM, Tai TC, Watson J, Peña MA, Hilborn A, Hannah CG, Rooper CN, Flynn
  KL, Oldford GL (2024). "pacea: An R package of Pacific ecosystem
  information to help facilitate an ecosystem approach to fisheries
  management."
  https://github.com/pbs-assess/pacea.
```
