---
title: "Groundfish_IPHC_Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtaining Groundfish IPHC Relative Abundance Indices

This markdown document shows how longline catch data on numerous groundfish species, collected by IPHC, are compiled into PACea. Spatiotempoal covariates are made, in the form of a spatial raster of density for each year from 1995 onwards.

First, we download the groundfish IPHC data using the gfiphc and hookCompetition packages 

```{r}
library(gfiphc)
library(dplyr)

dataset_commonname <-
  c("Yelloweye_Rel_Density",
    "Arrowtooth_Rel_Density",
    "Lingcod_Rel_Density",
    "Dogfish_Rel_Density",
    "Cod_Rel_Density",
    "Halibut_Rel_Density",
    "Redbanded_Rel_Density",
    "Sablefish_Rel_Density",
    "Walleye_Rel_Density")

fit_mod <- F

if(fit_mod)
{
  gf_data <-
  hookCompetition::read_Data_hookcomp(
    species_vec = c("yelloweye rockfish",
            "arrowtooth flounder",
            "lingcod",
            "north pacific spiny dogfish",
            "pacific cod",
            "pacific halibut",
            "redbanded rockfish",
            "sablefish",
            "walleye pollock"),
    at_PBS = T, 
    simple_species_vec = c("yelloweye",
            "arrowtooth",
            "lingcod",
            "dogfish",
            "cod",
            "halibut",
            "redbanded",
            "sablefish",
            "walleye"),
    min_dist = 1e6, 
    min_dist_to_boundary = 1e6
  )
}


```
Next, we fit spatio-temporal censored Poisson models to the Pacific Herring predators to create density layers for relative abundance for each year. We extract the density layer for each species and add it to a list object

```{r}
if(fit_mod)
{
  pacific_herring_predators_ind <- 
  vector('list', length = length(c("yelloweye",
            "arrowtooth",
            "lingcod",
            "dogfish",
            "cod",
            "halibut",
            "redbanded",
            "sablefish",
            "walleye")))
  
  # Make spatial objects required for the fit
  spatial_obj <- 
    hookCompetition::make_spatial_objects(
      gf_data$Reduced_data_sp,
      
    )
  
  names(pacific_herring_predators_ind) <-
  c("yelloweye rockfish",
    "arrowtooth flounder",
    "lingcod",
    "north pacific spiny dogfish",
    "pacific cod",
    "pacific halibut",
    "redbanded rockfish",
    "sablefish",
    "walleye pollock")

count <- 1
for(i in names(pacific_herring_predators_ind))
{
  pacific_herring_predators_ind[[count]] <-
    hookCompetition::spatiotemp_censored_index_fun (
      data=gf_data$Reduced_data_sp[
        gf_data$Reduced_data_sp$year>=1997 &
          gf_data$Reduced_data_sp$station %in% 
          names(table(gf_data$Reduced_data_sp$station))[table(gf_data$Reduced_data_sp$station)>2],], 
      species=i,
      return=T, ICR_adjust=F, cprop=0.95, 
      nthreads=1, keep=F, use_upper_bound=TRUE, 
      upper_bound_quantile=0.85, plot=T, 
      allyears=F, station_effects=T, 
      prior_event=HT.prior(), prior_station=HT.prior(), 
      n_knots=8, seed=0, verbose=F, n_trajectories=1,
      preserve_inter_regional_differences = T,
      spatiotemporal=F, mesh=spatial_obj$mesh, 
      spde=spatial_obj$spde_mod, 
      pixels=spatial_obj$predict_pixels
)$pred_df_plot
  count <- count + 1
}

# save the indices for future use
usethis::use_data_raw(pacific_herring_predators_ind)

}

if(!fit_mod)
{
  pacific_herring_predators_ind <- readRDS('pacific_herring_predators_ind.rds')
  
  names(pacific_herring_predators_ind) <-
  c("yelloweye",
    "arrowtooth",
    "lingcod",
    "dogfish",
    "cod",
    "halibut",
    "redbanded",
    "sablefish",
    "walleye")
}


```

Next, we map the indices to the partition regions

```{r}
# read in BC Grid
BC_Partition_Objects <- PACea::BC_Partition_Objects

# compute the mean values per grid cell - keep only 'BC Grid'
Partition_Polys <- BC_Partition_Objects$BC_Partition[
  BC_Partition_Objects$BC_Partition$Regions_Name=='BC Grid',]

# Convert to wide format
# Then read in as a rasterBrick with repeated locations to save memory
predators_raster <- 
  vector('list', length(pacific_herring_predators_ind))

names(predators_raster) <-
  c("Yelloweye",
    "Arrowtooth",
    "Lingcod",
    "Dogfish",
    "Cod",
    "Halibut",
    "Redbanded",
    "Sablefish",
    "Walleye")
for( i in 1:length(pacific_herring_predators_ind) )
{
  predators_raster[[i]] <- 
    cbind(sf::st_drop_geometry(pacific_herring_predators_ind[[i]]),
          sf::st_coordinates(pacific_herring_predators_ind[[i]])) %>%
    dplyr::filter(!is.na(region_INLA)) %>%
    tidyr::pivot_wider(id_cols = c(X, Y),
                names_from = year, values_from = median)
  
  predators_raster[[i]] <-
    raster::brick(sp::SpatialPixelsDataFrame(
      points=cbind(predators_raster[[i]]$X, predators_raster[[i]]$Y),
      data = predators_raster[[i]][,-c(1,2)],
      tolerance = 0.0025,
      proj4string = sp::CRS(sf::st_crs(Partition_Polys)$proj4string)
    ))
  
}

# remove the original data.frame
rm(pacific_herring_predators_ind)

predator_years <- as.numeric(substr(names(predators_raster[[1]]),2,10))

# Create a data.frame object for storing the aggregated predator abundance values
nyear=length(min(predator_years):max(predator_years))
npoly=length(Partition_Polys$Regions_Name)
nspecies=length(predators_raster)

Partition_df <-
  data.frame(Poly_ID=rep(1:npoly, times=nyear*nspecies),
             Species=rep(rep(names(predators_raster), each=npoly), times=nyear),
             Year=rep(min(predator_years):max(predator_years), each=npoly*nspecies),
             Mean_Rel_Abund=NA,
             Fraction_NA=NA
  )

# Loop through the years and months and form the mappings - save memory this way
for(i in min(predator_years):max(predator_years))
{
  print(paste0('Currently processing year ',i))
   
    # Find the corresponding column in the raster data
    raster_ind <- which(predator_years==i)

    
    for(j in 1:length(predators_raster))
    {
      
      Partition_df <- 
        Partition_df %>%
        dplyr::filter() %>%
        dplyr::mutate(Mean_Rel_Abund = 
                 ifelse(Species==names(predators_raster)[j] & Year==i,
                        exactextractr::exact_extract(
                          predators_raster[[j]][[raster_ind]],
                          Partition_Polys,
                          fun='mean'
                        ), Mean_Rel_Abund),
                       Fraction_NA =
                   ifelse(Species==names(predators_raster)[j] & Year==i,
                     exactextractr::exact_extract(
                        predators_raster[[j]][[raster_ind]],
                        Partition_Polys,
                        fun=function(value,cov_frac){
                          1-sum(cov_frac*!is.na(value))/sum(cov_frac)
                          }
                        ), Fraction_NA),
                 Fraction_NA = ifelse(is.na(Fraction_NA),1,Fraction_NA)
        )
      
    }
    print('successfully mapped to polygons')
    
}


# Convert to wide format with a column per species
Groundfish_IPHC_DF <-
  Partition_df %>%
  tidyr::pivot_wider(
    id_cols = c(Year, Poly_ID),
    names_from = c(Species),
    values_from = c(Mean_Rel_Abund, Fraction_NA)
  )
  
usethis::use_data(Groundfish_IPHC_DF, overwrite = T)

# save the data key, pointing the common names of the datasets to the correct DF
# data_key <- 
#   data.frame(
#     Common_Name = dataset_commonname,
#     DF_Name = 'Groundfish_IPHC_DF',
#     Time_Resolution = 'Annual',
#     Data_Type = 'Model-based Raster',
#     Units = 'Relative Density (Proportional to #Individuals per Unit Area)',
#     Author = 'Joe Watson',
#     Citation = NA,
#     Comments = ''
#   )

# append to the master Data_Key and remove duplicates
# Data_Key <-
#   rbind(PACea::Data_Key,
#         data_key)
# 
# Data_Key <-
#   Data_Key[!duplicated(Data_Key),]
# 
# # Update the master key
# usethis::use_data(Data_Key, overwrite=T)


```

Now make a quick plot of yelloweye rockfish density in 1999 as a sanity check

```{r}
 ggplot2::ggplot(cbind(Partition_Polys, 
              Groundfish_IPHC_DF %>% 
                dplyr::filter(Year==1999) %>% 
                dplyr::select(Mean_Rel_Abund_Yelloweye)
              ), 
        ggplot2::aes(fill=Mean_Rel_Abund_Yelloweye)) + 
  ggplot2::geom_sf()

# Extract only the indices that correspond to the West Coast Haida Gwaii
BC_Partition_Objects$index_vectors$BC_Major_Area_Boundaries[3]
# Index 58 is the correct row from the BC_Partition_Objects$Mapping_Matrix

# Compute the mean using the Mapping_Matrix
WCHG_Mean_1999 <-
t(BC_Partition_Objects$Mapping_Matrix[,
                                    BC_Partition_Objects$index_vectors$BC_Major_Area_Boundaries[3]
                          ]) %*% 
  as.matrix(
  Groundfish_IPHC_DF %>% 
       dplyr::filter(Year==1999) %>% 
       mutate(Mean_Rel_Abund_Yelloweye=
                ifelse(is.na(Mean_Rel_Abund_Yelloweye),
                       0,
                       Mean_Rel_Abund_Yelloweye)) %>%
       dplyr::pull(Mean_Rel_Abund_Yelloweye)
  )


ggplot2::ggplot(BC_Partition_Objects$BC_Partition[
  BC_Partition_Objects$BC_Partition$Poly_Name=='WCHG',
], 
        ggplot2::aes(fill=WCHG_Mean_1999)) + 
  ggplot2::geom_sf()

# Plot the fraction of the polygon which has missing data
 ggplot2::ggplot(cbind(Partition_Polys, 
              Groundfish_IPHC_DF %>% 
                dplyr::filter(Year==1999) %>% 
                dplyr::select(Fraction_NA_Yelloweye)
              ), 
        ggplot2::aes(fill=Fraction_NA_Yelloweye)) + 
  ggplot2::geom_sf()
```

