---
title: "Analysing recruitment indices from 2026 hake assessment"
author: "Andrew Edwards"
output: pdf_document
fontsize: 12pt
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  cache = TRUE,
  cache_path = "hake-pbsedm-2026-cache/",
  fig.path = "hake-pbsedm-2026-figs-cache/",
  fig.width = 7,
  fig.height = 6
)

knitr::dep_prev()
```

```{r, build, echo = FALSE, eval = FALSE}
# To build either run this line or click knit button in RStudio:
rmarkdown::render("hake-pbsedm-2026.Rmd")
```

```{r, packages, echo = FALSE, cache = FALSE}
# load_all()  # for pacea, maybe don't need though
# Or:
# remotes::install_github("pbs-assess/pacea")
# library(pacea)
library(dplyr)
library(pbsEDM)
# library(pacea)    # make sure to update
library(gridExtra)
library(ggplot2)

# From Chris's hake
f <- function(x, digits = 0){

  if(is.null(x)){
    return(x)
  }

  format(round(x,
               digits),
         big.mark = ",",
         nsmall = digits)
}

```

This is a quick attempt at using pbsEDM to analyse the indices we are showing
the 2026 Pacific Hake stock assessment, which are based on those determined as
important for recruitment by Vestfals et al.

Using `sockeyPrize/report/fraser/chilko/analysis-2025-4.Rmd` as a
template. Whittling down as necessary.
This should actually be a bit simpler, just need to put all the indices together
(we have some from pacea and some from GLORYS, already both saved). One wrinkle
is that they are already shifted so that they are aligned with the expected
year of recruitment of which the variable is influencing; see hake assessment
ecosystem summary appendix. In theory we would hope that EDM then teases out
lags of 0, as the true lag of 1 is already accounted for.

Ideally will resave the unlagged values, but for now the figure code is
intertwined with the saving code in `ecosystem-summary-hake.R` and
`ecosystem-summary-hake-glorys.R`, so have to undo both to do that. No time
right now.

So lags for all covariates should in theory come out as 0 being best. 

```{r, load}
assess_yr <- 2026
hake_indices_pacea <- readRDS(file = paste0("../hake_indices_pacea_",
                                             assess_yr,
                                             ".rds"))

hake_indices_glorys <- readRDS(file = paste0("../hake_indices_glorys_",
                                             assess_yr,
                                             ".rds"))
summary(hake_indices_pacea)
summary(hake_indices_glorys)
```

So there is the issue that the time series for pacea is longer than that for
glorys. So maybe for now just investigate each one separately.

Want to predict `hake_recruitment_median` for each set of covariates. Could also
do for recruitment deviations at some point, though recruitment will get
normalised here over the relevant time period.


Delete this once decided don't need it for reference:
```{r sumreturns, eval = FALSE}
one_returns <- dplyr::filter(returns_all,
                             River == river)

one_returns_age45 <- rowwise(one_returns) %>%
  mutate(age4 = sum(c_across(any_of(classes_age4)),
                    na.rm = TRUE)) %>%
  mutate(age5 = sum(c_across(any_of(classes_age5)),
                    na.rm = TRUE)) %>%
  ungroup() %>%
  select("System",
         "River",
         "ReturnYear",
         "age4",
         "age5")

one_returns_age45 %>% a()

input_returns <- select(one_returns_age45,      # only for Chilko
                      year = ReturnYear,
                      age4,
                      age5) %>%
  filter(year != 1951)       # since age4 and age5 are 0 for 1951.
which(input_returns == 0)    # 1952 age5 for Chilko, check others manually here
input_returns[input_returns == 0] = NA

input_returns

# in analysis-2025-3.Rmd had added NA for 2025, but then take out later, so just
# not bothering here.

# Now add five earlier years with NA's, since things like PDO and spawners might
#  go back 4-5 years.
yr <- min(input_returns$year)
yr
input_returns <- rbind(c(yr - 5, NA, NA),
                     c(yr - 4, NA, NA),
                     c(yr - 3, NA, NA),
                     c(yr - 2, NA, NA),
                     c(yr - 1, NA, NA),
                     input_returns)
input_returns
input_returns %>% tail()

if(min(input_returns$year != 1947)){
  stop("check covariates and years")
}

```

## Covariates from pacea

Just want the median hake recruitment:
```{r input}
input_pacea <- select(hake_indices_pacea,
                      -c("hake_recruitment_low",
                         "hake_recruitment_high",
                         "cumulative_index"))
input_pacea %>% head(10) %>% as.data.frame()
input_pacea %>% tail(10) %>% as.data.frame()                     ```

```{r, indices}
summary(input_pacea)

T <- nrow(input_pacea)   # T as in index T is final year of data, can include
                        # NA's at start. Not used in calculations, just plotting
                        # and printing results.

T    # do double check that manually, esp for age5 and other rivers


# TODO might need 
#
hake_first_year_index <- min(which(!is.na(input_pacea$hake_recruitment_median)))
#and age5 if needed
#age4_first_year_index

hake_first_year <- input_pacea$year[hake_first_year_index]
hake_first_year
```

Vestfals et al. already did some of the ideas part that we had for the Sockeye analysis.

# EDM with covariates

- Do a time series panel plot first of all time series.

- Then do the correlation plots to take a look.
options though. See multiview embedding also.

Do a big plot of everything:
```{r plotcovariates, fig.height = 9}
to_plot <- tidyr::pivot_longer(input_pacea,
                               cols = -year) %>%
  mutate(name = factor(name))

p <- ggplot(to_plot,
            aes(year, value)) +
  geom_path() +                        # leaves gaps for NA's, but none now
  # facet_wrap(~name, scales = "free_y")
  facet_grid(name ~ ., scales = "free_y")

p

# It reorders them alphabetically; need something like this:
#  mutate(name = factor(name, levels = paste0("X", 1:12))) %>%
```

Not worrying about span idea for now, can select the relevant covariates as lags
in EDM (i.e. just exclude ones that don't have full span).

```{r corrplot}
# pairs(input_pacea,
#      lower.panel = panel.cor)
GGally::ggpairs(input_pacea)
```

Useful, but of course doesn't highlight correlation of lags.
```{r corrplot22, fig.width = 9}
corrplot::corrplot(cor(input_pacea,
                       use = "pairwise.complete.obs"),
  method = "number",
  type = "upper" # show only upper side
  )
```

In Haigh et al. (2019) we used a correlation coefficient between covaraties of
0.3 to exclude related covariates. Let's do that here.

Just looking manually at the figure, we can see that should not have, in the
same analysis (because correlation has magnitude $>0.3$):


- `pdo_pre_index` and `bi_index` -0.45. Be nice to test both separately.

Just do one run with PDO and one with BI.

Doing everything first as a baseline, then doing those two options.

## Only covariates that span the full time series - see analysis-2025-3.Rmd

# MVE analysis

### Including everything

Notes from sockeye:

These results do match those from `full8` in `analysis-2025-3.Rmd` from which
this is adapted.

Timings. Each extra lag should be $2^XX$ longer.

9 lags - 3.8 seconds

12 lags - 21 seconds (8 times longer, makes sense.

14 lags - 90 seconds (makes sense, four times longer, ish)

15 lags - 200 secs, about three minutes (makes sense).

16 lags - should be about 320 seconds, 5 minutes.

17 lags - took 720 seconds, 12 minutes before. Then 1587 secs, 26 minutes at
home.

18 lags - probably about an hour, so a sensible limit.

Change eval once run once. Just try 0 and 1 lags for now (0 should get teased
out if results are consistent with Vestfals et al. as we've already shifted the
years)
```{r, full9, eval = TRUE}
lags_full <- list(hake_recruitment_median = c(0, 1),
                  herring_index = c(0, 1),
                  bi_index = c(0,1),
                  pdo_pre_index = c(0,1),
                  npgo_pre_index = c(0,1),
                  hake_log_age_1_index = c(0,1))  

tictoc::tic()

HERE - next one gives an error in untranform_predictions, so has done the EDM
part very quickly
mve_full <- multiview_embedding(data = input_pacea,   # not span
                                response = "hake_recruitment_median",
                                lags = lags_full)
# Centre and scaling and first differencing get done automatically, in
#  state_space_reconstruction_for_sve()
tictoc::toc()

saveRDS(mve_full9, file = "mve_full9.rds")
```

```{r, full9load}
mve_full9 <- readRDS("mve_full9.rds")

mve_full9$rho_prediction_from_mve
mve_full9$rho_each_top_subset

# So there's lots, so need to just show a few of these:
mve_full9$lags_of_top_subsets[1:20]

# Look at the top few predictions:

mve_full9[["response_each_subset"]][mve_full9$top_subsets_for_rho_index[1:5]]

mve_full9$response_predicted_from_mve   # Overall mve prediction, including for T+1.
```

```{r full9fig}
plot(input_pacea$age4,
     mve_full9$response_predicted_from_mve[-(T+1)],
     xlab = "Age4 returns - observed",
     ylab = "Age4 returns - predicted")

legend("topleft", legend = c(paste("rho is",
                                    round(mve_full9$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_full9$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_full9$response_predicted_from_mve[T+1], col = "red")
```

## Remove correlated covariates

As noted above, removing three correlated covariates. Then running one analysis
with PDO and one with BI. Won't take too long to run.

```{r, reduced}
input_pacea_reduced <- select(input_pacea,
                              -c("fraser_peak_spring",
                                 "total",
                                 "total_spawners"))
```

### With PDO not BI

First, lags with PDO (not BI). Do not eval this chunk after doing it once if it
takes a while and previous chunks have changed.

```{r, red1}
lags_reduced_pdo <- list(age4 = c(1, 2, 3, 4),
                         effective_spawners = c(4),
                         pdo_winter_mean = c(1, 2),
                         fraser_mean_spring = c(1, 2),
                         pink = c(1, 2),
                         sst_spring = c(1, 2),
                         zoo_total_biomass = c(1, 2))
                         # bi_anomaly = c(1, 2))
tictoc::tic()
mve_reduced_pdo <- multiview_embedding(data = input_pacea_reduced,
                                       response = "age4",
                                       lags = lags_reduced_pdo)
# Centre and scaling and first differencing get done automatically, in
#  state_space_reconstruction_for_sve()
tictoc::toc()

saveRDS(mve_reduced_pdo, file = "mve_reduced_pdo.rds")
```

Reload and then see results.

```{r, red1load}
mve_reduced_pdo <- readRDS("mve_reduced_pdo.rds")

mve_reduced_pdo$rho_prediction_from_mve
mve_reduced_pdo$rho_each_top_subset

# So there's lots, so need to just show a few of these:
mve_reduced_pdo$lags_of_top_subsets[1:20]

# Look at the top few predictions:

mve_reduced_pdo[["response_each_subset"]][mve_reduced_pdo$top_subsets_for_rho_index[1:5]]

mve_reduced_pdo$response_predicted_from_mve   # Overall mve prediction, including for T+1.
```

```{r red1fig}
plot(input_pacea$age4,
     mve_reduced_pdo$response_predicted_from_mve[-(T+1)],
     xlab = "Age4 returns - observed",
     ylab = "Age4 returns - predicted")

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_pdo$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_pdo$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_pdo$response_predicted_from_mve[T+1], col = "red")
```

### With BI not PDO

Do same for BI. Pick the best, to be the prediction. Then repeat
for age5, and add together.

```{r, red2, eval = FALSE}
lags_reduced_bi <- list(age4 = c(1, 2, 3, 4),
                        effective_spawners = c(4),
                        bi_anomaly = c(1, 2),
                        fraser_mean_spring = c(1, 2),
                        pink = c(1, 2),
                        sst_spring = c(1, 2),
                        zoo_total_biomass = c(1, 2))

tictoc::tic()
mve_reduced_bi <- multiview_embedding(data = input_pacea_reduced,
                                      response = "age4",
                                      lags = lags_reduced_bi)
# Centre and scaling and first differencing get done automatically, in
#  state_space_reconstruction_for_sve()
tictoc::toc()

saveRDS(mve_reduced_bi, file = "mve_reduced_bi.rds")
```

Reload and then see results.

```{r, red2load}
mve_reduced_bi <- readRDS("mve_reduced_bi.rds")

mve_reduced_bi$rho_prediction_from_mve
mve_reduced_bi$rho_each_top_subset

# So there's lots, so need to just show a few of these:
mve_reduced_bi$lags_of_top_subsets[1:20]

# Look at the top few predictions:

mve_reduced_bi[["response_each_subset"]][mve_reduced_bi$top_subsets_for_rho_index[1:5]]

mve_reduced_bi$response_predicted_from_mve   # Overall mve prediction, including
# for T+1.

lm_reduced_bi <- lm(mve_reduced_bi$response_predicted_from_mve[-(T+1)] ~
                      input_pacea$age4)

```

```{r red2fig}
plot(input_pacea$age4,
     mve_reduced_bi$response_predicted_from_mve[-(T+1)],
     xlab = "Age4 returns - observed",
     ylab = "Age4 returns - predicted")

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_bi$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_bi$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_bi$response_predicted_from_mve[T+1], col = "red")
```

## Age 5

### With PDO not BI

If we use the same as age4 plus the extra lags (see notes above) it will take
about 6 hours, so was going to look through results above of first 20 sets of lags and not
use any here that didn't seem important. But can get a bit confusing with
age4/age5 difference, and so might not be obvious what to remove. But take a
look. So here are the lags, then copy the code that defined them, then remove
ones that show up in top 20 sets, to leave ones not in top 20.

First set of age4 lags.
```{r, lagscheck}
lags_reduced_pdo
lags_reduced_pdo_not_in_top_20 <- list(
                                       fraser_mean_spring = c(1),
                                       zoo_total_biomass = c(1, 2))
```

So may as well not use those for age5, since need to narrow down. Zooplankton
could be because of the short time series, haven't dug into it.

So results in 18 lags below, which is about an hour.

First, lags with PDO (not BI). Do not eval this chunk after doing it once if it
takes a while and previous chunks have changed.

```{r, red3}
lags_reduced_pdo_age5 <- list(age4 = c(1),
                              age5 = c(1, 2, 3, 4, 5),
                              effective_spawners = c(5),
                              pdo_winter_mean = c(1, 2, 3),
                              fraser_mean_spring = c(2, 3),
                              pink = c(1, 2, 3),
                              sst_spring = c(1, 2, 3))
                              # zoo_total_biomass = c(1, 2, 3))
                              # bi_anomaly = c(1, 2))
tictoc::tic()
mve_reduced_pdo_age5 <- multiview_embedding(data = input_pacea_reduced,
                                            response = "age5",
                                            lags = lags_reduced_pdo_age5)
# Centre and scaling and first differencing get done automatically, in
#  state_space_reconstruction_for_sve()
tictoc::toc()

saveRDS(mve_reduced_pdo_age5, file = "mve_reduced_pdo_age5.rds")
```

Reload and then see results.

```{r, red3load}
mve_reduced_pdo_age5 <- readRDS("mve_reduced_pdo_age5.rds")

mve_reduced_pdo_age5$rho_prediction_from_mve
mve_reduced_pdo_age5$rho_each_top_subset

# So there's lots, so need to just show a few of these:
mve_reduced_pdo_age5$lags_of_top_subsets[1:20]

# Look at the top few predictions:

mve_reduced_pdo_age5[["response_each_subset"]][mve_reduced_pdo_age5$top_subsets_for_rho_index[1:5]]

mve_reduced_pdo_age5$response_predicted_from_mve   # Overall mve prediction, including for T+1.
```

```{r red3fig}
plot(input_pacea$age5,
     mve_reduced_pdo_age5$response_predicted_from_mve[-(T+1)],
     xlab = "Age5 returns - observed",
     ylab = "Age5 returns - predicted")

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_pdo_age5$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_pdo_age5$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_pdo_age5$response_predicted_from_mve[T+1], col = "red")
```

### With BI not PDO

Do same for BI. First, manually check lags from age4, see what was not in the
top 20. For age4:

```{r, lagscheckbi}
lags_reduced_bi


lags_reduced_bi_not_in_top_20 <- list(
                                      zoo_total_biomass = c(1, 2))
```
So `fraser_mean_spring` lag 1 was useful when BI included, so include here. Will
take about two
hours then with 19 lags.

```{r, red4}
lags_reduced_bi_age5 <- list(age4 = c(1),
                             age5 = c(1, 2, 3, 4, 5),
                             effective_spawners = c(5),
                             # pdo_winter_mean = c(1, 2, 3),
                             fraser_mean_spring = c(1, 2, 3),
                             pink = c(1, 2, 3),
                             sst_spring = c(1, 2, 3),
                             # zoo_total_biomass = c(1, 2, 3))
                             bi_anomaly = c(1, 2, 3))
tictoc::tic()
mve_reduced_bi_age5 <- multiview_embedding(data = input_pacea_reduced,
                                           response = "age5",
                                           lags = lags_reduced_bi_age5)
# Centre and scaling and first differencing get done automatically, in
#  state_space_reconstruction_for_sve()
tictoc::toc()

saveRDS(mve_reduced_bi_age5, file = "mve_reduced_bi_age5.rds")
```

Reload and then see results.

```{r, red4load}
mve_reduced_bi_age5 <- readRDS("mve_reduced_bi_age5.rds")

mve_reduced_bi_age5$rho_prediction_from_mve
mve_reduced_bi_age5$rho_each_top_subset

# So there's lots, so need to just show a few of these:
mve_reduced_bi_age5$lags_of_top_subsets[1:20]

# Look at the top few predictions:

mve_reduced_bi_age5[["response_each_subset"]][mve_reduced_bi_age5$top_subsets_for_rho_index[1:5]]

mve_reduced_bi_age5$response_predicted_from_mve   # Overall mve prediction, including
# for T+1.

lm_reduced_bi_age5 <- lm(mve_reduced_bi_age5$response_predicted_from_mve[-(T+1)] ~
                      input_pacea$age5)

```

```{r red4fig}
plot(input_pacea$age5,
     mve_reduced_bi_age5$response_predicted_from_mve[-(T+1)],
     xlab = "Age5 returns - observed",
     ylab = "Age5 returns - predicted")

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_bi_age5$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_bi_age5$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_bi_age5$response_predicted_from_mve[T+1], col = "red")
```

## Summarise results

Summarise results of the four runs, then pick the best for age4 and age5 and add
them together. Plot the four figures together, with the same xlim and ylim.

```{r, summarylim}
lim_common_max <- max(c(input_pacea$age4,
                        mve_reduced_pdo$response_predicted_from_mve,
                        mve_reduced_bi$response_predicted_from_mve,
                        input_pacea$age5,
                        mve_reduced_pdo_age5$response_predicted_from_mve,
                        mve_reduced_bi_age5$response_predicted_from_mve),
                      na.rm = TRUE)
lim_common_max
lim_common = c(0, lim_common_max)
```

Panel plot (should write a function, but this is quick enough):
```{r, summaryplot}
par(mfrow = c(2,2))

plot(input_pacea$age4,
     mve_reduced_pdo$response_predicted_from_mve[-(T+1)],
     xlab = "Age4 returns - observed",
     ylab = "Age4 returns - predicted",
     main = "PDO age4",
     xlim = lim_common,
     ylim = lim_common)

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_pdo$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_pdo$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_pdo$response_predicted_from_mve[T+1], col = "red")


plot(input_pacea$age4,
     mve_reduced_bi$response_predicted_from_mve[-(T+1)],
     xlab = "Age4 returns - observed",
     ylab = "Age4 returns - predicted",
     main = "BI age4",
     xlim = lim_common,
     ylim = lim_common)

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_bi$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_bi$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_bi$response_predicted_from_mve[T+1], col = "red")


plot(input_pacea$age5,
     mve_reduced_pdo_age5$response_predicted_from_mve[-(T+1)],
     xlab = "Age5 returns - observed",
     ylab = "Age5 returns - predicted",
     main = "PDO age5",
     xlim = lim_common,
     ylim = lim_common)

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_pdo_age5$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_pdo_age5$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_pdo_age5$response_predicted_from_mve[T+1], col = "red")


plot(input_pacea$age5,
     mve_reduced_bi_age5$response_predicted_from_mve[-(T+1)],
     xlab = "Age5 returns - observed",
     ylab = "Age5 returns - predicted",
     main = "BI age5",
     xlim = lim_common,
     ylim = lim_common)

legend("topleft", legend = c(paste("rho is",
                                    round(mve_reduced_bi_age5$rho_prediction_from_mve, 2)),
                              paste("prediction is",
                                    f(mve_reduced_bi_age5$response_predicted_from_mve[T+1]))))
abline(a = 0, b = 1)
abline(h = mve_reduced_bi_age5$response_predicted_from_mve[T+1], col = "red")
```


```{r, summarise}
summary <- tibble(age = "age4",
                  run = "PDO",
                  rho = mve_reduced_pdo$rho_prediction_from_mve ,
                  prediction = mve_reduced_pdo$response_predicted_from_mve[T+1])
summary <- rbind(summary,
                 c("age4",
                   "BI",
                   mve_reduced_bi$rho_prediction_from_mve,
                   mve_reduced_bi$response_predicted_from_mve[T+1]),
                 c("age5",
                   "PDO",
                   mve_reduced_pdo_age5$rho_prediction_from_mve,
                   mve_reduced_pdo_age5$response_predicted_from_mve[T+1]),
                 c("age5",
                   "BI",
                   mve_reduced_bi_age5$rho_prediction_from_mve,
                   mve_reduced_bi_age5$response_predicted_from_mve[T+1])) %>%
  dplyr::mutate_at(vars(age,
                        run),
                   factor) %>%
  dplyr::mutate_at(vars(rho,
                        prediction),
                   as.numeric)

summary

knitr::kable(summary,
             digits = c(NA, NA, 2, 6))
```

Now pick the best age4 and best age5, and add them together for our prediction
for `r river`:
```{r, combine}
age4_best_index <- which.max(filter(summary,
                                    age == "age4")$rho)
age5_best_index <- which.max(filter(summary,
                                    age == "age5")$rho) +
  sum(summary$age == "age4")        # Assumes always do age4 then age5

summary_best <- summary[c(age4_best_index,
                          age5_best_index), ]
expect_equal(nrow(summary_best), 2)        # just check there's no duplicates

knitr::kable(summary_best,
             digits = c(NA, NA, 2, 7))

total_returns_prediction <- sum(summary_best$prediction)
```

The prediction for combined age4 and age5 returns `r river` is ... `r f(total_returns_prediction)`.

### More notes

Can also see other methods in pbsEDM/R.

And `analysis-2025-3.Rmd` for looking at continuity of some of the time series,
and doing a shorter timespan.
