---
title: "British Columbia Coastline, EEZ and ROMS grid generation"
author: "Travis Tai"
date: "2023-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# set to github/pacea folder
setwd("C:/github/pacea")

library(devtools)
library(ncdf4)
library(ggplot2)
library(sf)
library(dplyr)
library(sp)
library(terra)
load_all()
```

## Generating polygons

Boundary polygons were created for geographic layers and to mask interpolated data. 

### BC coastline

First we tested BC coastline files from three different packages: `bcmaps`, `PBSdata`, and `rnaturalearth`.

Looking a the detail, native file format and metadata, we determined that the 'rnaturalearth' was best suited to produce a coastline 'sf' object for the pacea package. See 'data-raw/coastline/coastline-testing.R' script for testing different mapping options.

To generate the `bc_coast` object for pacea, we used 'rnaturalearth'; code can be found in 'data-raw/coastline/coastline-eez.R'. 

### BC exclusive economic zone

The BC exclusive economic zone object was taken from the 'PBSdata' package. However, the class of the file is a 'polyset', and the object had to be transformed into a shapefile 'SpatialPolygons' object. The 'maptools' package that contains the function to convert a 'polyset' object is being deprecated, so these operations may not work.

See 'data-raw/coastline/coastline-eez.R' for generation of the 'sf' object.


```{r}
head(bc_coast)
head(bc_eez)

plot(bc_coast)
plot(bc_eez)
```

### Create grid

Below is testing of a mixed grid. 

```{r}
# Spatial grid creation

library(sf)
library(devtools)
library(ggplot2)
library(terra)
library(ncdf4)
load_all()

sf_use_s2(FALSE)  # remove spherical geometry (s2) for sf operations

#####
# load data to environment

# GIS hub Pacific Marine Habitat Classes
#  obtained from: https://www.gis-hub.ca/dataset/marine-habitat-classes
#  requires access to gis-hub
pmhc_dir <- "./data-raw/pacific_marine_habitat_classes/Pacific_Marine_Habitat_Classes.gdb"
(pmhc_layers <- st_layers(pmhc_dir))

# use benthic habitat (bh) layer
pmhc <- st_read(pmhc_dir, layer = pmhc_layers$name[3])
str(pmhc)
unique(pmhc$Habitat)

# pacea - rnaturalearth bc coast
data(bc_coast)

# pacea - PBSdata eez
data(bc_eez)


#####
# make sf grid 
grid2 <- st_make_grid(romsbuff_poly, cellsize = 2000)  # 2x2km
grid6 <- st_make_grid(romsbuff_poly, cellsize = 6000)  # 6x6km

# crop grid of buffer area and remove land area cells
grid2c <- grid2[romsbuff_poly]
grid2c <- grid2c[inshore_poly] %>% st_as_sf()
grid6c <- grid6[romsbuff_poly] 
grid6c <- grid6c[offshore_poly] 
grid6c <- st_difference(grid6c, st_union(grid2c)) %>% st_as_sf()

# combine grids
grid26 <- grid6c %>% rbind(grid2c) 

grid26_data <- grid26 %>%
  mutate(dummyvar1 = rnorm(nrow(grid26)))

dum_mat <- matrix(rnorm(nrow(grid26)*360), ncol = 360, nrow = nrow(grid26))

grid26_databig <- grid26 %>% 
  cbind(dum_mat)

grid26_nogeom <- grid26_databig %>% 
  st_drop_geometry()


# visualization
ggplot() + 
  geom_sf(data = grid26_sf, fill = NA) +
  geom_sf(data = tbc) + 
  geom_sf(data = inshore_poly, fill = NA) +
  geom_sf(data = offshore_poly, fill = NA) 

# size of sf grid object, with and without one column of data
# usethis::use_data(grid26)  # without data:                   140 kb
# usethis::use_data(grid26_data)  # with data:                 741 kb
# usethis::use_data(grid26_databig, overwrite=T)  # with 360 column data:     210,578 kb
# usethis::use_data(grid26_nogeom, overwrite=T)   # with 360 column, no geom: 210,431 kb

##  removing the geometry from sf object with 77360 cells, and 360 layers saves about 147 kb, vs 140kb 


#####
## read in and create sst data

nc_dat <- nc_open("data-raw/roms/bcc42_era5glo12r4_mon1993to2019_surTSOpH.nc")
nc_lon <- as.vector(ncvar_get(nc_dat, "lon_rho"))
nc_lat <- as.vector(ncvar_get(nc_dat, "lat_rho"))

nc_var <- ncvar_get(nc_dat, "temp")
nc_varmat <- apply(nc_var, 3, c)

dat <- data.frame(x = nc_lon, y = nc_lat) %>% cbind(nc_varmat)
dat_sf <- st_as_sf(dat, coords = c("x", "y"), crs = "EPSG:4326")

# subset only one layer
tsst_sf1 <- st_transform(dat_sf, crs = "EPSG: 3005")[, 1] %>%
  rename(sst = '1')

#####
# make raster grid
rast2 <- rast(ext(bccm_eez_poly), res = c(2000, 2000), crs = st_crs(bccm_eez_poly)$wkt)
rast3 <- rast(ext(bccm_eez_poly), res = c(3000, 3000), crs = st_crs(bccm_eez_poly)$wkt)
rast6 <- rast(ext(bccm_eez_poly), res = c(6000, 6000), crs = st_crs(bccm_eez_poly)$wkt)


#####
# force data point values into grid and use 'fun' to summarize point data into raster cells
##  I don't think you can include a predict function for interpolation
summary(tsst_sf1)

# create raster to assign values to
r <- terra::rast(tsst_sf1)

# vectorize values for terra package
v <- terra::vect(tsst_sf1)

# rasterize sf point data to grid:
##  this uses a simple 'function' to assign a value to gridded cell based on points that fall within cell
##  can end up with empty cells (at high resolution)

# 2km grid
res(r) <- 2000
sst_rast2 <- terra::rasterize(v, r, field = "sst")
sst_rast2
plot(sst_rast2,range = c(3.4, 10.2))

# 3km grid
res(r) <- 3000
sst_rast3 <- terra::rasterize(v, r, field = "sst")
sst_rast3
plot(sst_rast3,range = c(3.4, 10.2))

# 4km grid
res(r) <- 4000
sst_rast4 <- terra::rasterize(v, r, field = "sst")
sst_rast4
plot(sst_rast4,range = c(3.4, 10.2))

# 6km grid
res(r) <- 6000
sst_rast6 <- terra::rasterize(v, r, field = "sst")
sst_rast6
plot(sst_rast6,range = c(3.4, 10.2))

# resample 6km grid to 2km grid
rr2 <- resample(sst_rast6, rast2, method = "bilinear")

# doesn't work
#rr2 <- resample(v, rast2, method = "bilinear")

summary(rr2)

plot(rr2,range = c(3.4, 10.2))
plot(bccm_eez_poly, add = T)

## NOTE: these methods are not the best for interpolation to a grid, especially at a higher resolution




```

Creating the grid for use an object in pacea was done using code in the file: 
'./data-raw/grids/create-grid.R'

```{r}
head(grid26)

plot(grid26)
```






